### 2010-06-29 paper.tex
Initial import.

\documentclass[preprint]{aastex}

\usepackage{hyperref}
\usepackage{amsmath}

\newcommand{\fixme}[1]{[FIXME: #1]}
\newcommand{\Msun}{M_\odot}
\newcommand{\Mmin}{M_{\textnormal{min}}}
\newcommand{\Mmax}{M_{\textnormal{max}}}
\newcommand{\Nbin}{N_{\textnormal{bin}}}

\bibliographystyle{hapj}

\begin{document}

\title{The Mass Distribution of Stellar-Mass Black Holes}

\author{Will M. Farr \and Niharika Sravan} \affil{Northwestern
  University Center for Interdisciplinary Exploration and Research in
  Astrophysics\\2145 Sheridan Rd., Evanston, IL 60208}
\email{w-farr@northwestern.edu, niharika.sravan@gmail.com}

\author{Andrew Cantrell \and Laura Kreidberg \and Rachel Macdonald
  \and Charles Bailyn}
\affil{Yale University Department of Astrophysics\\P.O. Box
  208101, New Haven, CT 06520}
\email{andrew.cantrell@yale.edu, laura.kreidberg@yale.edu,
  rachel.macdonald@yale.edu, charles.bailyn@yale.edu}

\author{Vicky Kalogera} \affil{Northwestern University Center for
  Interdisciplinary Exploration and Research in Astrophysics\\2145
  Sheridan Rd., Evanston, IL 60208} \email{vicky@northwestern.edu}

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes based on the observed masses of the compact
  object in 17 low mass X-ray binaries.  Modeling the mass
  distribution both parametrically---as a power law, exponential,
  gaussian, or combination of two gaussians---and
  non-parametrically---as histograms with varying numbers of
  bins---we explore the distribution in the context of Markov Chain
  Monte Carlo calculations of model parameters.  We give confidence
  bounds on the shape of the mass distribution in the context of each
  model and compare the models with each other by calculating their
  Bayesian evidence.  We address the existence of a ``gap'' between
  the most massive neutron stars and least massive black holes,
  finding that the best model (the power law), the second-best model
  (a one-bin histogram, or flat distribution), and the third-best
  model (a two-bin histogram) have minimum black hole masses above
  5.00, 3.23, and 3.82 solar masses (95\% confidence), respectively.
  We therefore conclude that our sample of black hole masses provides
  strong evidence of a gap between the maximum neutron star mass and
  minimum black hole mass.
\end{abstract}

\maketitle

\section{Introduction}
\label{sec:intro}

\section{Systems}
\label{sec:systems}

The 17 low mass X-ray binary systems on which this study is based are
listed in Table \ref{tab:sources}.  In each of these systems,
spectroscopic measurements of the secondary star provide an orbital
period for the system and a half-amplitude for the secondary's
velocity curve.  These measurements can be combined into the mass
function,
\begin{equation}
  \label{eq:mass-function}
  f(M) = \frac{P K^3}{2\pi G} = \frac{M \sin^3 i}{\left( 1 + q \right)^2},
\end{equation}
where $P$ is the orbital period, $K$ is the secondary's velocity
semi-amplitude, $M$ is the black hole mass, $i$ is the inclination of
the system, and $q \equiv M_2 / M$ is the mass ratio of the system.

The mass function defines a lower limit on the mass: $f(M) < M$.  To
accurately determine the mass of the black hole, the inclination $i$
and mass ratio $q$ must be measured.  Ideally, this can be
accomplished by fitting ellipsoidal light curves from these systems
but even in the ideal case (see, e.g., \citet{Cantrell2010} on A0620)
this procedure can be complicated by possible contributions from an
accretion disk and hot spots in the disk.  For some systems (e.g.\ GS
1354 \citep{Casares2009}) strong variability prevents determination of
the inclination from the lightcurve at all; in these cases an upper
limit on the inclination often comes from the observed lack of
eclipses in the lightcurve.  In general, accurately determining $q$
and $i$ requires a careful system-by-system analysis.

For the purposes of this paper, we adopt the following simplified
approach to the estimation of the black hole mass from the observed
data.  When an observable is well-constrained, we assume that the true
value is normally distributed about the measured value with a standard
deviation equal to the quoted observational error.  This is the case
for the mass function in all the systems we use, and for many systems'
mass ratios and inclinations.  When a large range is quoted in the
literature for an observable, we take the true value to be distributed
uniformly (for the mass ratio) or isotropically (for the inclination)
within the quoted range.  Table \ref{tab:sources} gives the assumed
distribution for the observables in the 17 systems we use.  We do not
attempt to deal with the systematic biases in the observational
determination of $f$, $q$, and $i$ in any realistic way, instead
leaving this for future work.

\begin{table}
  \begin{center}
    \begin{tabular}{|l|c|c|c|l|}
      \hline
      Source & $f$ ($\Msun$) & $q$ & $i$ (degrees) & References \\
      \hline \hline
      GRS 1915 & $N(9.5, 3.0)$ & $N(0.0857, 0.0284)$ & $N(70, 2)$ &
      \citet{Greiner2001} \\
      XTE J1118 & $N(6.44, 0.08)$ & $N(0.0264, 0.004)$ & $N(68, 2)$ &
      \citet{Gelino2008} \\ & & & & \citet{Harlaftis2005} \\
      Cyg X1 & $N(0.251, 0.007)$ & $N(2.778, 0.386)$ & $I(23, 38)$ &
      \citet{Gies2003} \\ 
      XTE J1650 & $N(2.73, 0.56)$ & $U(0, 0.5)$ & $I(50, 80)$ &
      \cite{Orosz2004} \\
      GRS 1009 & $N(3.17, 0.12)$ & $N(0.137, 0.015)$ & $I(37, 80)$ &
      \cite{Filippenko1999} \\
      M33 X7 & $N(0.46, 0.08)$ & $N(4.47, 0.61)$ & $N(74.6, 1)$ &
      \citet{Orosz2007} \\
      A0620 & $N(2.76, 0.036)$ & $N(0.06, 0.004)$ & $N(50.98, 0.87)$ &
      \citet{Cantrell2010} \\ & & & & \citet{Neilsen2008} \\
      GRO J0422 & $N(1.13, 0.09)$ & $U(0.076, 0.31)$ & $N(45, 2)$ &
      \citet{Gelino2003} \\
      Nova Mus 1991 & $N(3.01, 0.15)$ & $N(0.128, 0.04)$ & $N(54,1.5)$
      & \cite{Gelino2001} \\
      GRO J1655 & $N(2.73,0.09)$ & $N(0.3663, 0.04025)$ & $N(70.2,
      1.9)$ & \citet{Greene2001} \\
      4U 1543 & $N(0.25, 0.01)$ & $U(0.25, 0.31)$ & $N(20.7,1.5)$ & 
      \citet{Orosz2003} \\
      XTE J1550 & $N(7.73,0.4)$ & $U(0,0.04)$ & $N(74.7, 3.8)$ &
      \citet{Orosz2010} \\
      V4641 Sgr & $N(3.13,0.13)$ & $U(0.42,0.45)$ & $N(75,2)$ &
      \citet{Orosz2003} \\
      GS 2023 & $N(6.08, 0.06)$ & $U(0.056,0.063)$ & $N(55,4)$ &
      \citet{Charles2006} \\
      GS 1354 & $N(5.73, 0.29)$ & $N(0.12,0.04)$ & $I(50, 80)$ & 
      \citet{Casares2009} \\
      Nova Oph 77 & $N(4.86,0.13)$ & $U(0, 0.053)$ & $I(60, 80)$ &
      \citet{Charles2006} \\
      GS 2000 & $N(5.01, 0.12)$ & $U(0.035, 0.053)$ & $I(43, 74)$ &
      \citet{Charles2006} \\
      \hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:sources} The source parameters for the 17 low
    mass X-ray binaries used in this work.  $f$ is the mass function for
    the compact object, $q$ is the mass ratio $M_2/M$, and $i$ is the
    inclination of the system to the line of sight.  We indicate the
    distribution used for the true parameters when computing the
    probability distributions for the masses of these systems:
    $N(\mu,\sigma)$ implies a Gaussian with mean $\mu$ and standard
    deviation $\sigma$, $U(a,b)$ is a uniform distribution between $a$ and
    $b$, and $I(\alpha,\beta)$ is an isotropic distribution between the
    angles $\alpha$ and $\beta$.  \fixme{Orosz for XTE J1550 needs a
      real reference \citep{Orosz2010}.}} 
\end{table}

From these assumptions, we can generate probability distributions for
the true mass of the black hole given the observations and errors via
the Monte Carlo method: drawing samples of $f$, $q$, and $i$ from the
assumed distributions and computing the mass implied by Equation
\eqref{eq:mass-function} gives samples of $M$ from the distribution
induced by the relationship in Equation \eqref{eq:mass-function}.
Mass distributions generated in this way for the 17 systems used in
this work are shown in Figure \ref{fig:all-masses}.  These mass
distributions constitute the ``observational data'' we will use in
Section \ref{sec:models} to explore various underlying mass
distributions.

\begin{figure}
  \begin{center}
    \plotone{plots/all-masses} 
  \end{center}

  \caption{\label{fig:all-masses} The mass distributions implied by
    Equation \eqref{eq:mass-function} and the assumed distributions on
    observational parameters given in Table \ref{tab:sources}.  The
    significant asymmetry and long tails in many of these
    distributions are the result of the non-linear relationship
    (Equation \eqref{eq:mass-function}) between $M$, $f$, $q$, and
    $i$.}
\end{figure}

In Figure \ref{fig:masses} we show the combination of these individual
mass distributions into one single distribution.  (Note that this
combination is not a good estimator of the true mass distribution of
black hole masses---see, e.g., \citet{Mandel2010}.  We include it here
only to give a sense of the range of masses of black holes in our
sample; we will come to estimators for the mass distribution.)

\begin{figure}
  \plotone{plots/masses}

  \caption{\label{fig:masses} The mass distributions in Figure
    \ref{fig:all-masses} combined into a single distribution. Note
    that this combination is not a good estimator of the true mass
    distribution of black hole masses in the galaxy---see, e.g.,
    \citet{Mandel2010}.  We include it here only to give a sense of
    the range and general shape of the mass distribution of black
    holes in our sample.}
\end{figure}

\section{Models}
\label{sec:models}

In this section we apply a Bayesian analysis to various models for the
underlying mass distribution from which the systems in Table
\ref{tab:sources} were drawn.  The end result will be the probability
distribution for the parameters of each model implied by the data from
Section \ref{sec:systems} in combination with our prior assumptions
about the probability distribution for the parameters.  Bayes' rule
relates these quantities.  For a model with parameters $\theta$ in the
presence of data $d$, Bayes' rule states
\begin{equation}
  \label{eq:Bayes-rule}
  p(\theta | d) = \frac{p(d | \theta) p(\theta)}{p(d)}.
\end{equation}
Here, $p(\theta|d)$, called the posterior probability distribution
function, is the probability distribution for the parameters $\theta$
implied by the data $d$; $p(d|\theta)$, called the likelihood, is the
probability of observing data $d$ given that the model parameters are
$\theta$; $p(\theta)$, called the prior, reflects our estimate of the
probability of the various model parameters in the absence of any
data; and $p(d)$, called the evidence, is an overall normalizing
constant ensuring that 
\begin{equation}
  \int d\theta\, p(\theta|d) = 1,
\end{equation}
whence
\begin{equation}
  \label{eq:evidence-def}
  p(d) = \int d\theta\, p(d|\theta) p(\theta).
\end{equation}

In our context, the data are the mass distributions given in Section
\ref{sec:systems}: $d = \{ p_i(m)| i = 1, 2, \ldots, 17 \}$.  We
assume that the measurements in Section \ref{sec:systems} are
independent, so the complete likelihood is given by a product of the
likelihoods for the individual measurements.  For a model with
parameters $\theta$ that predicts a mass distribution $p(m|\theta)$
for black holes, we have
\begin{equation}
  \label{eq:likelihood-def}
  p(d|\theta) = \prod_i \int dm\, p_i(m) p(m|\theta).
\end{equation}
That is, the likelihood of an observation is the average over the
individual mass distribution implied by the observation, $p_i(m)$, of
the probability for a black hole of that mass to exist according to
the model of the mass distribution, $p(m | \theta)$.  We approximate
the integrals as averages of $p(m|\theta)$ over the Monte Carlo mass
samples drawn from the distributions in Table \ref{tab:sources} (also
see Figure \ref{fig:all-masses}):
\begin{equation}
  p(d|\theta) \approx \prod_i \frac{1}{N_i} \sum_{j = 1}^{N_i} p(m_{ij} | \theta),
\end{equation}
where $m_{ij}$ is the $j$th sample (out of a total $N_i$) from the
$i$th individual mass distribution.

Our calculation of the likelihood of each observation does not include
any attempt to account for selection effects in the observations.  We
simply assume (almost certainly incorrectly) that any black hole drawn
from the underlying mass distribution is equally likely to be
observed, leaving the analysis of selection effects to future work.

For a mass distribution with several parameters, $p(\theta | d)$ lives
in a multi-dimensional space.  Exploring the entirety of this space
systematically rapidly becomes prohibitive as the number of parameters
increases.  A more efficient way to explore the distribution $p(\theta
| d)$ is to use a Markov Chain Monte Carlo (MCMC) method.  MCMC
methods produce a chain (sequence) of parameter samples, $\{ \theta_i,
i = 1, \ldots \}$, such that a particular parameter, $\theta$, appears
in the sequence with a frequency equal to its probability according to
$p(\theta|d)$.  In this way, regions of parameter space where
$p(\theta|d)$ is large are sampled densely while regions where
$p(\theta|d)$ is small are effectively ignored.  A Markov chain has
the property that the transition probability from one element to the
next, $p(\theta_i \to \theta_{i+1})$, depends only on the value of
$\theta_i$, not on any previous values in the chain.

One way to produce a sequence of MCMC samples is via the following
algorithm, first proposed by \fixme{Metropolis citation needed}:
\begin{enumerate}
  \item Begin with the current sample, $\theta_i$.
  \item Propose a new sample, $\theta_p$, by drawing randomly from an
    arbitrary ``jump proposal distribution'' with probability
    $Q(\theta_i \to \theta_p)$.
  \item Compute the ``acceptance'' probability,
    \begin{equation}
      \label{eq:paccept}
      p_{\textnormal{accept}} \equiv
      \frac{p(\theta_p|d)}{p(\theta_i|d)} \frac{Q(\theta_p \to
        \theta_i)}{Q(\theta_i \to \theta_p)}
    \end{equation}
  \item With probability $\min(1,p_{\textnormal{accept}})$ set
    $\theta_{i+1} = \theta_p$; otherwise set $\theta_{i+1} =
    \theta_i$.
\end{enumerate}
Note that this algorithm is more likely to accept a proposed jump when
it increases the posterior (the first term in Equation
\eqref{eq:paccept}) and when it is to a location in parameter space
from which it is easy to return (the second term in Equation
\eqref{eq:paccept}).  As $i \to \infty$ the samples $\theta_i$ are
distributed according to $p(\theta|d)$.  In practice the number of
samples required before the chain appropriately samples $p(\theta|d)$
depends strongly on the jump proposal distribution; proposal
distributions that often propose jumps toward or within regions of
large $p(\theta|d)$ can be very efficient, while poor proposal
distributions can require prohibitively large numbers of samples
before convergence.

Once we have a chain of samples from $p(\theta|d)$---a distribution of
distributions---the distribution for any quantity of interest can be
computed.  For example, Figure \ref{fig:dists} shows the median
values, and the 10\%, and 90\% quantiles for value of the BH mass
probability distribution at various masses in each of the models
discussed in the following subsections.  Such a figure can be
generated from the MCMC output by computing the value of the
parameterized BH mass distribution at various masses for each
parameter sample output by the MCMC, and then plotting the quantiles
at each mass.

\begin{figure}
  \begin{center}
    \plottwo{plots/dist-parameteric}{plots/dist-non-parameteric}
  \end{center}
  \caption{\label{fig:dists} The median values of the black hole mass
    distribution, $p(m|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Subsections
    \ref{subsec:parameteric-models} and
    \ref{subsec:non-parameteric-models}.  Error bars span the 10\% to
    90\% range.  Note that these ``distributions of distributions''
    are not necessarily normalized, and need not be ``shaped'' like
    the underlying model distributions.}
\end{figure}

It is also common to look at the one-dimensional distribution for a
single parameter obtained by integrating over all other dimensions in
parameter space; this is approximated by a histogram of the MCMC
sample values for that parameter.  Such a distribution is called the
``marginalized'' distribution.

We defer discussion of the appropriateness of the different models
given the data (i.e.\ the ``goodness of fit'') until Section
\ref{sec:model-selection}.

\subsection{Parameteric Models for the Black Hole Mass Distribution}
\label{subsec:parameteric-models}

In this subsection, we discuss the various parameteric models of the
underlying black hole mass distribution that we have analyzed using
the data from Section \ref{sec:systems}.

\subsubsection{Power-Law Models}
\label{subsubsec:power-law}

Many astrophysical distributions are power-laws.  Let us assume that
the BH mass distribution is given by
\begin{equation}
  \label{eq:power-law-dist}
  p(m|\theta) = p(m|\{\Mmin, \Mmax, \alpha \}) =
  \begin{cases}
    A m^\alpha & \Mmin \leq m \leq \Mmax \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The normalizing constant $A$ is given by 
\begin{equation}
  A = \frac{1+\alpha}{\Mmax^{1+\alpha} - \Mmin^{1+\alpha}}.
\end{equation}
To perform the Bayesian analysis, we need to impose a prior on the
parameters $\Mmin$, $\Mmax$, $\alpha$.  We choose to impose a broad
prior that represents our significant uncertainty in the appropriate
values of these parameters.  We use
\begin{equation}
  p(\theta) = p(\{\Mmin, \Mmax, \alpha\}) = 
  \begin{cases}
    \frac{1}{2} \frac{1}{40^2}\frac{1}{20} & 0 \leq \Mmin \leq \Mmax
    \leq 40, \quad -12 \leq \alpha \leq 8 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where the masses are measured in solar masses.  This prior is uniform
in the masses between 0 and $40 \Msun$, and uniform in power-law slope
between $-12$ and $8$.

Our MCMC analysis output is a list of $\{\Mmin, \Mmax, \alpha\}$
values distributed according to the posterior 
\begin{equation}
  p(\theta|d) = p(\{\Mmin, \Mmax, \alpha\}|d) \propto p(d|\{\Mmin,
  \Mmax, \alpha\}) p(\{\Mmin, \Mmax, \alpha\}),
\end{equation}
with the likelihood $p(d|\{\Mmin, \Mmax, \alpha\})$ defined in
Equation \eqref{eq:likelihood-def}.  Figure \ref{fig:dists} presents
the resulting distribution of mass distributions inferred from the
MCMC distribution of parameters.

In Figure \ref{fig:alpha}, we display a histogram of the resulting
samples in $\alpha$; this represents the one-dimensional
``marginalized'' distribution
\begin{equation}
  \label{eq:alpha-pdf}
  p(\alpha|d) = \int d\Mmin\, d\Mmax\, p(\{\Mmin, \Mmax, \alpha\}|d).
\end{equation}
The resulting distribution is quite broad with 
\begin{equation}
  -7.66 < \alpha < -1.36
\end{equation}
enclosing 90\% of the probability.  The median value is $\alpha =
-4.28$, and we have $\alpha < 0$ (a falling underlying mass
distribution) at 98.7\% confidence.  The distribution of the minimum
mass is also quite interesting, but we defer this discussion until
Section \ref{sec:minimum-mass}, where we discuss the distribution for
the minimum black hole masses in the context of all our models.

\begin{figure}
  \begin{center}
    \plotone{plots/alpha}
  \end{center}
  \caption{\label{fig:alpha} A histogram of the MCMC samples for
    $\alpha$ from the power-law distribution defined in Equation
    \eqref{eq:power-law-dist}.  The histogram is an approximation to
    the one-dimensional marginalized PDF in Equation
    \eqref{eq:alpha-pdf}.  The favored value is near $\alpha = -4$.
    The fraction of the distribution with $\alpha < 0$ is 98.7\%.  The
    5\% quantile is $-7.66$, the median is $-4.28$, and the 95\%
    quantile is $-1.36$.}
\end{figure}

\subsubsection{Decaying Exponential}
\label{subsubsec:exponential}

Theoretical arguments in \citet{Fryer2001} suggest that the black-hole
mass distribution may be well-represented by a decaying exponential
with a minimum mass:
\begin{equation}
  \label{eq:exp-def}
  p(m|\theta) = p(m|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{m}{M_0}
    \right] & M_0 \leq m \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose again a broad, flat prior
\begin{equation}
  p(\{\Mmin, M_0 \}) = 
  \begin{cases}
    \frac{1}{40^2} & 0 \leq \Mmin, M_0 \leq 40 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
with $\Mmin$ and $M_0$ measured in solar masses.

Figure \ref{fig:exp-M0} displays the marginalized posterior
distribution for the scale mass of the exponential, $M_0$.  The median
scale mass is $M_0 = 3.00$, and $1.57 \leq M_0 \leq 5.71$ with 90\%
confidence.  We defer discussion of the distribution of $\Mmin$ until
Section \ref{sec:minimum-mass}.  Refer to Figure \ref{fig:dists}
for the resulting shape of the distribution of distributions.

\begin{figure}
  \begin{center}
    \plotone{plots/exp-M0}
  \end{center}
  \caption{\label{fig:exp-M0} The distribution of scale masses, $M_0$,
    measured in units of a solar mass for the exponential underlying
    mass distribution defined in Equation \eqref{eq:exp-def}.  The
    median scale mass is $M_0 = 3.00$; with 90\% confidence $1.57 \leq
    M_0 \leq 5.71$. }
\end{figure}

\subsubsection{Gaussian and Two-Gaussian Models}
\label{subsubsec:gaussian}

The prototypical single-peaked probability distribution is a
Gaussian:
\begin{equation}
  p(m|\theta) = p(m|\{\mu, \sigma\}) = \frac{1}{\sigma \sqrt{2\pi}}
  \exp\left[ - \left(\frac{m - \mu}{\sqrt{2} \sigma} \right)^2 \right].
\end{equation}
We choose a broad, flat prior 
\begin{equation}
  \label{eq:gaussian-def}
  p(\{\mu,\sigma\}) = 
  \begin{cases}
    \frac{1}{40^2} & 0 \leq \mu, \sigma \leq 40 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where both $\mu$ and $\sigma$ are measured in solar masses.  Figure
\ref{fig:gaussian} shows the resulting marginalized distributions for
the parameters $\mu$ and $\sigma$.  We constrain $\mu$ in $7.02 \leq
\mu \leq 9.23$ with 90\% confidence.  (Refer to Figure \ref{fig:dists}
for the shape of the resulting distribution of distributions.)

\begin{figure}
  \begin{center}
    \plotone{plots/gaussian}
  \end{center}
  \caption{\label{fig:gaussian} Marginalized posterior distributions
    for the mean $\mu$ and standard deviation $\sigma$ (both in solar
    masses) for the Gaussian underlying mass distribution defined in
    Equation \eqref{eq:gaussian-def}.  The peak of the Gaussian,
    $\mu$, is constrained in $7.02 \leq \mu \leq 9.23$ with 90\%
    confidence.}
\end{figure}

To look for a second peak in the black-hole mass distribution, we also
examined a two-Gaussian model:
\begin{multline}
  p(m|\theta) = p(m|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right],
\end{multline}
where we restrict $\mu_1 < \mu_2$ with our broad, flat prior:
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    \frac{1}{2} \frac{1}{40^4} & 0 \leq \mu_1 \leq \mu_2 \leq 40,
    \quad 0 \leq \sigma_1, \sigma_2 \leq 40, \quad 0 \leq \alpha \leq
    1 \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The one-Gaussian model is a subset of the Gaussian model: when
$\alpha$ is small, then the presence of the first Gaussian is
irrelevant, and the second gives the same mass distribution as a
single Gaussian; when $\alpha$ is large, the reverse situation again
reproduces the single-Gaussian model; and when $\alpha$ is in the
middle of its range, the two means and standard deviations can be
nearly equal, reproducing again the single-Gaussian model.  This is
precisely what we observe from our MCMC samples in the two-Gaussian
model.  In Section \ref{sec:model-selection}, we will quantify more
precisely how much redundancy we have in this model relative to the
one-Gaussian model, but we should emphasize here that there is no
evidence of a significant contribution from a second peak in the
two-Gaussian model.

\subsection{Non-Parameteric Models for the Black Hole Mass Distribution}
\label{subsec:non-parameteric-models}

The previous subsection discussed models for the underlying black hole
mass distribution that assumed particular parameterized shapes for the
distribution.  In this subsection, we will discuss models that do not
assume a priori a shape for the black hole mass distribution.  The
fundamental non-parameteric distribution in this section is a
histogram with some number of bins, $\Nbin$.  Such a distribution is
piecewise-constant in $m$.

One choice for representing such a histogram would be to fix the bin
locations, and allow the heights to vary.  With this approach, one
should be careful not to ``split'' features of the mass distribution
across more than one bin in order to avoid diluting the sensitivity to
such features; similarly, one should avoid including more than ``one''
feature in each bin.  The locations of the bins, then, are crucial.
An alternative representation of histogram mass distributions avoids
this difficulty.

We choose to represent a histogram mass distribution with $\Nbin$ bins
by allocating a fixed total probability, $1/\Nbin$, to each bin.  The
lower and upper bounds for each bin are allowed to vary; when these
are close to each other (i.e.\ the bin is narrow), the distribution
will have a large value, and conversely when the bounds are far from
each other.  We assume that the non-zero region of the distribution is
contiguous, so we can represent the boundaries of the bins as a
non-decreasing array of masses, $w_0 \leq w_1 \leq \ldots \leq
w_{\Nbin}$, with $w_0$ the minimum and $w_{\Nbin}$ the maximum mass for
which the distribution has support.  This gives the distribution
\begin{equation}
  \label{eq:hist-def}
  p(m|\theta) = p(m|\{w_0, \ldots, w_{\Nbin}\}) = 
  \begin{cases}
    0 & m < w_0 \textnormal{ or } w_{\Nbin} \leq m \\
    \frac{1}{\Nbin} \frac{1}{w_{i+1} - w_i} & w_i \leq m < w_{i+1}
  \end{cases}.
\end{equation}

For priors on the histogram model with $\Nbin$ bins, we assume that
the bin boundaries are uniformly distributed between 0 and $40 \Msun$
subject only to the constraint that the boundaries are non-decreasing
from $w_0$ to $w_{\Nbin}$:
\begin{equation}
  p(\{w_0, \ldots, w_{\Nbin}\}) = 
  \begin{cases}
    \frac{\left(\Nbin+1\right)!}{40^{\Nbin+1}} & 0 \leq w_0 \leq w_1
    \leq \ldots \leq w_{\Nbin} \leq 40 \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}

The median values of the histogram mass distributions that result from
the MCMC samples of the posterior distribution for the $w_i$
parameters for one-, two-, three-, four-, and five-bin histogram
models are shown in Figure \ref{fig:dists}.  As the number of bins
increases, the models are better able to capture features of the mass
distribution.  (But note that in Section \ref{sec:model-selection}, we
will find that the one-bin histogram---which is a flat distribution
between a maximum and minimum mass---is the best model from this
group.)  We defer discussion of the distribution of the minimum mass
from the histogram distributions until Section \ref{sec:minimum-mass}.

\section{Model Selection}
\label{sec:model-selection}

In Section \ref{sec:models}, we discussed a series of models for the
underlying black hole mass distribution.  For each model, we have
assumed that the underlying mass distribution corresponds to the
model, and we ask what distributions are implied by the data for the
parameters in the model.  We have not yet asked which models are more
likely to correspond to the actual distribution.  That is the topic of
this section.

To compare models in the context of a Bayesian analysis, consider a
set of models, $\{M_i| i = 1, \ldots\}$, each with corresponding
parameters $\theta_i$.  Re-writing Equation \eqref{eq:Bayes-rule} to
be explicit about the assumption of a particular model, we have
\begin{equation}
  \label{eq:bayes-explicit-model}
  p(\theta_i | d, M_i) = \frac{p(d|\theta_i, M_i) p(\theta_i | M_i)}{p(d|M_i)}.
\end{equation}
This gives the posterior probability of the parameters $\theta_i$ in
the context of model $M_i$.  But, the particular model itself can be
regarded as a parameter in a larger ``super-model'' that encompasses
all the $M_i$.  If we let the model become a parameter, Bayes' rule
gives us
\begin{equation}
  p(\theta_i, M_i|d) = \frac{p(d|\theta_i, M_i) p(\theta_i |M_i) p(M_i)}{p(d)},
\end{equation}
where we have introduced the model prior $p(M_i)$, which represents
our estimate on the probability that model $M_i$ is correct in the
absence of the data $d$.  The normalizing evidence is now
\begin{equation}
  p(d) = \sum_i \int d\theta_i\, p(\theta_i, M_i|d).
\end{equation}

To compare the various models $M_i$, we are interested in the
marginalized posterior probability of $M_i$ (also called the evidence
for model $M_i$):
\begin{equation}
  p(M_i|d) \equiv \int d\theta_i\, p(\theta_i, M_i|d).
\end{equation}
This quantity can be re-written in terms of the single-model evidence
(see Equations \eqref{eq:bayes-explicit-model} and
\eqref{eq:evidence-def}):
\begin{equation}
  \label{eq:model-evidence-def}
  p(M_i|d) = \frac{p(M_i)}{p(d)} \int d\theta_i
  p(d|\theta_i,M_i) p(\theta_i|M_i) = \frac{p(d|M_i) p(M_i)}{p(d)}.
\end{equation}

There are two broad ways to compare models in the context of the MCMC
analysis of Section \ref{sec:models}.  The first way, which we discuss
in Subsection \ref{subsec:direct-evidence}, is to use the MCMC samples
from the various models to estimate the value of the integral in
Equation \eqref{eq:model-evidence-def}.  The second way is to treat
the model literally as a parameter in a second MCMC analysis.  One
then proposes jumps (see the discussion in Section \ref{sec:models})
not only for the $\theta_i$ within a particular model, but from
parameters $\theta_i$ in model $M_i$ to parameters $\theta_j$ in
$M_j$.  This technique is called a ``reversible-jump MCMC''
\citep{Green1995}; we discuss it in Subsection
\ref{subsec:reversible-jump-mcmc}.

\subsection{Direct Evidence Calculations}
\label{subsec:direct-evidence}

In this section we discuss techniques for using the MCMC samples from
Section \ref{sec:models} to estimate the integral in Equation
\eqref{eq:model-evidence-def}.  The first technique we use comes from
re-writing Equation \eqref{eq:bayes-explicit-model}:
\begin{equation}
  \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} = \frac{p(\theta_i|M_i)}{p(d|M_i)}.
\end{equation}
Integrating both sides over all $\theta_i$, recalling that the prior
is normalized, we have 
\begin{equation}
  \int d\theta_i\, \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} = \frac{1}{p(d|M_i)}.
\end{equation}
We can estimate the integral using our MCMC samples (which are
distributed with local density $p(\theta_i|d,M_i)d\theta_i$) via
\begin{equation}
  \int d\theta_i\, \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} \approx
  \frac{1}{N_i} \sum_j \frac{1}{p(d|\theta_{ij}, M_i)},
\end{equation}
where the $j$th MCMC parameter sample (from a total $N_i$) is
$\theta_{ij}$.  Thus we have
\begin{equation}
  \frac{1}{p(d|M_i)} \approx \frac{1}{N_i} \sum_j \frac{1}{p(d|\theta_{ij}, M_i)};
\end{equation}
the evidence for model $M_i$ can be estimated from the harmonic mean
of the likelihood over the MCMC samples.

Alternately, we can attempt to perform the integral over $\theta_i$
for the model evidence directly using our MCMC samples.  If we could
assign non-overlapping volumes, $V_j$, to each of the $N_i$ MCMC
samples, $\theta_{ij}$, then we could approximate the integral in
Equation \eqref{eq:model-evidence-def} via
\begin{equation}
  \label{eq:direct-evidence-estimate}
  \int d\theta_i
  p(d|\theta_i,M_i) p(\theta_i|M_i) \approx \sum_j
  p(d|\theta_{ij},M_i) p(\theta_{ij}|M_i) V_j.
\end{equation}

To associate a volume to each MCMC sample, we use a data structure
known as a kD-tree.  (Our algorithm is similar to Algorithm P of
\citet{Weinberg2010}.)  A kD-tree decomposes a volume in
$\mathbb{R}^n$ into non-overlapping boxes, each of which contains some
number of points from a given sample.  The decomposition proceeds
recursively: given a region to subdivide and a sample of points, a
dimension is chosen (we choose the dimension of largest extent of the
sample, but other choices are possible) and the region is split along
this dimension such that the numbers of points to each side of the
partition are equal or differ by one.  Each of the resulting
sub-regions and the sub-samples contained in them are further split in
the same manner, until every region contains fewer than some chosen
number of points.  The sequence of splittings forms a binary tree,
with the initial region and point sample at the root, and the final
boxes and subsamples at the leaves.  From experimentation on
artificial point distributions, we find that terminating the splitting
when each box contains fewer than $\sim 64$ to $\sim 256$ samples is
reasonably effective in estimating evidence integrals in $\sim 5$
dimensions.  (See also \citet{Weinberg2010}.)  To estimate the
integral as in Equation \eqref{eq:direct-evidence-estimate}, we walk
the leaves of the kD-tree; for each leaf, we compute the volume
spanned by the samples in that leaf, and assign an equal fraction of
this volume to each sample (note that the volume spanned by the
samples is \emph{smaller} than the leaf region itself).  The integral
is computed using the assigned volumes as in Equation
\eqref{eq:direct-evidence-estimate}.

The harmonic mean estimate of the evidence is known for having large
statistical uncertainty due to the tendency for a few low-likelihood
points to dominate the sum \citep{Weinberg2010}; the statistics of the
kD-tree estimate are less well-known.  In any case, it is useful to
have multiple independent estimates of the evidence for a given model.
In Figure \ref{fig:direct-evidence} we use both methods to estimate
the evidence for the various models discussed in Section
\ref{sec:models}.  We assume that no model is a priori more likely
than any other, so the model priors $p(M_i)$ are equal.  We also
ignore the model-independent constant $p(d)$ in Equation
\eqref{eq:model-evidence-def}, so only ratios of evidence are
relevant.  Due to the large statistical uncertainty expected in the
harmonic mean estimate of the evidence, we perform bootstrap estimates
of the range of the harmonic mean, reporting the 10\% and 90\% values
in the error bars of Figure \ref{fig:direct-evidence}.

\begin{figure}
  \begin{center}
    \plotone{plots/evidence}
  \end{center}
  \caption{\label{fig:direct-evidence} Direct integration and harmonic
    mean evidence for the various models from Section \ref{sec:models}
    (see Subsection \ref{subsec:direct-evidence}).  In increasing
    order along the $x$-axis, the models are the power-law of
    Subsubsection \ref{subsubsec:power-law} (PL), the decaying
    exponential of Subsubsection \ref{subsubsec:exponential} (E), the
    single Gaussian of Subsection \ref{subsubsec:gaussian} (G), the
    double Gaussian of Subsubsection \ref{subsubsec:gaussian} (TG),
    and the one-, two-, three-, four-, and five-bin histogram models
    of Subsection \ref{subsec:non-parameteric-models} (H1, H2, H3, H4,
    H5, respectively).  The cross symbols give the value of the direct
    integration; the plus symbols give the harmonic mean estimator
    with bootstrap estimates of the 10\% and 90\% bounds on the
    harmonic mean.  It appears that the power law and one-bin
    histogram are the two best models, but the error ranges and
    inconsistency between the two techniques for calculating evidence
    cast some doubt on this conclusion.}
\end{figure}

The results of the direct evidence calculation appear to point to the
power law and one-bin histogram as the two most favored models, but
the error ranges and inconsistency between the two techniques for
calculating evidence cast some doubt on this conclusion.  In the next
section, we will confirm these results with a more accurate method for
estimating the evidence.

\subsection{Reversible-Jump MCMC}
\label{subsec:reversible-jump-mcmc}

The technique of reversible-jump MCMC \citep{Green1995} follows the
``super-model'' concept, in which the choice of model, $M_i$, becomes
a parameter, to its logical conclusion.  One performs an MCMC that
proposes jumps between the parameter spaces of the various models that
are under consideration (i.e.\ from parameters $\theta_i$ in model
$M_i$ to parameters $\theta_j$ in model $M_j$).  For this MCMC to be
efficient, proposed jumps into a model should favor regions with large
posterior; when the posterior is highly-peaked in a small region of
parameter space, proposed jumps outside this region are unlikely to be
accepted, and the reversible-jump MCMC samples will require a very
long chain to properly sample the ``super-model'' posterior.  The
evidence for a particular model is proportional to the number of
reversible-jump MCMC samples that live in that model's parameter
space.

We use the kD-trees from the last subsection to interpolate the
posterior between MCMC samples from single-model MCMCs when proposing
jumps in our reversible-jump MCMC.  Thus, proposed jumps track the
posteriors computed from the single-model MCMCs.  (For an alternative
approach to generating jump proposals for reversible-jump MCMCs, see
\citet{Littenberg2009}.)  The interpolation algorithm works as
follows.  We construct a kD-tree as described in the last subsection,
continuing the subdivision until each region contains exactly one
sample point.  To propose a jump, we first choose a model uniformly
from the set of models (i.e.\ according to the model priors).  Then we
choose a point uniformly from the sample chain (i.e.\ a point is
chosen according to the posterior).  Then we find the leaf region in
the kD-tree that contains this point, and propose a jump uniformly
within this region.  This procedure for proposing a jump admits a
straightforward calculation of the jump proposal distribution.  The
advantage of using a kD-tree instead of the method in
\citet{Littenberg2009} is that the tree is locally adaptive to the
structure of the MCMC samples, and therefore more efficiently tracks
the posterior from each of the single-model MCMCs.

We have performed a reversible-jump MCMC that jumps between all the
models (both parameteric and non-parameteric) in Section
\ref{sec:models}.  The counts for the number of samples in each model
appear in Figure \ref{fig:rj}.  The most favored model is the power
law from Subsubsection \ref{subsubsec:power-law}, followed by the one-
and two-bin histograms from Subsection
\ref{subsec:non-parameteric-models}.  This confirms the results from
the direct integration and harmonic mean evidence estimators in the
last subsection.  Interestingly, the theoretical curve from
\citet{Fryer2001} places fourth in the ranking of evidence.  However,
while the exponential distribution is a factor of 3.9 times less
likely than the best-fit power law model, it is only 1.2 times less
likely than the two best histogram models.  A choice of model prior
that only slightly favored the exponential model over the
non-parameteric histograms would put it in second place.

Assuming that the reversible-jump MCMC has converged (and we have
tested ours for convergence), the standard deviation on the number of
counts in each model is given by
\begin{equation}
  \sigma_{N_i} = \sqrt{N_i \left( 1 - \frac{N_i}{N} \right)},
\end{equation}
where $N_i$ is the number of counts in model $M_i$, and $N$ is the
total number of reversible-jump counts.  This quantity is extremely
small relative to the statistical error in the harmonic mean estimator
or the likely truncation error on the direct integration estimate, so
we consider the reversible-jump evidence results to be definitive.

\begin{figure}
  \begin{center}
    \plotone{plots/rj}
  \end{center}
  \caption{\label{fig:rj} Counts of samples in the parameter space of
    each model from Section \ref{sec:models} in the reversible-jump
    MCMC of Subsection \ref{subsec:reversible-jump-mcmc}.  The models
    are the power law of Subsubsection \ref{subsubsec:power-law} (PL),
    the exponential of Subsubsection \ref{subsubsec:exponential} (E),
    the Gaussian (G) and two Gaussian (TG) models of Subsubsection
    \ref{subsubsec:gaussian}, and the one-, two-, three-, four-, and
    five-bin histograms of Subsection
    \ref{subsec:non-parameteric-models} (H1, H2, H3, H4, H5,
    respectively).  The number of counts in each model is proportional
    to the evidence for that model; therefore, the most favored model
    is the power-law, followed by the one- and two-bin histogram.  }
\end{figure}

\section{The Minimum Mass of the Black Hole Mass Distribution}
\label{sec:minimum-mass}

Using the MCMC samples for the parameters of the distributions in
Section \ref{sec:models}, we can form probability distributions for
the minimum black hole mass implied by the model posteriors.  (For the
Gaussian models without a sharp cutoff we use the mass corresponding
to the 1\% quantile of the distribution as the ``minimum black hole
mass.'')  Figure \ref{fig:min-mass} displays the probability
distributions for the minimum mass in the various models.

\begin{figure}
  \begin{center}
    \plottwo{plots/mmin-parameteric}{plots/mmin-non-parameteric}
  \end{center}
  \caption{\label{fig:min-mass} The distributions for the minimum
    black hole mass calculated from the MCMC samples for the models in
    Section \ref{sec:models}.  For the Gaussian distributions without
    a sharp cutoff, we use the mass corresponding to the 1\% mass
    quantile for the ``minimum black hole mass.''  For the three
    most-favored models (the power law, one-, and two-bin histogram),
    the minimum black hole mass is above 5.00, 3.23, and 3.82 solar
    masses, respectively (at 95\% confidence). }
\end{figure}

For the three most-favored models (the power law, one-, and two-bin
histograms), the minimum black hole mass is above 5.00, 3.23, and 3.82
solar masses, respectively, with 95\% confidence.  While these limits
are lower than those of a previous study of fewer systems
\citep{Bailyn1998}, they are still significantly above the maximum
neutron star mass.  The measured masses of the systems in Section
\ref{sec:systems} show strong evidence of a ``gap'' between the
maximum neutron star and minimum black hole mass.

\section{Conclusion}

\bibliography{paper}

\end{document}

### 2010-06-29 plots/doepstopdf.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+set -e
+
+for file in *.eps; do
+    epstopdf $file
+done

### 2010-06-29 paper.tex
Acknowledgements stub, and updated author list.

@@
 
-\author{Andrew Cantrell \and Laura Kreidberg \and Rachel Macdonald
-  \and Charles Bailyn}
-\affil{Yale University Department of Astrophysics\\P.O. Box
-  208101, New Haven, CT 06520}
+\author{Andrew Cantrell \and Laura Kreidberg \and Charles Bailyn}
+
+\affil{Yale University Department of Astrophysics\\
+  P.O. Box 208101, New Haven, CT 06520}
+
 \email{andrew.cantrell@yale.edu, laura.kreidberg@yale.edu,
-  rachel.macdonald@yale.edu, charles.bailyn@yale.edu}
+  charles.bailyn@yale.edu}
 
@@
 
+\acknowledgements
+
+This work was supported by grant XXX.  WMF would like to thank Ilya
+Mandel for illuminating discussions and encouragement.
+
 \bibliography{paper}

### 2010-06-29 paper.tex
Added table for evidence ratios.

@@
     distribution, $p(m|\theta)$, at various masses implied by the
-    posterior $p(\theta|d)$ for the models discussed in Subsections
+    posterior $p(\theta|d)$ for the models discussed in Sections
     \ref{subsec:parameteric-models} and
@@
 analysis of Section \ref{sec:models}.  The first way, which we discuss
-in Subsection \ref{subsec:direct-evidence}, is to use the MCMC samples
+in Section \ref{subsec:direct-evidence}, is to use the MCMC samples
 from the various models to estimate the value of the integral in
@@
 $M_j$.  This technique is called a ``reversible-jump MCMC''
-\citep{Green1995}; we discuss it in Subsection
+\citep{Green1995}; we discuss it in Section
 \ref{subsec:reversible-jump-mcmc}.
@@
     mean evidence for the various models from Section \ref{sec:models}
-    (see Subsection \ref{subsec:direct-evidence}).  In increasing
+    (see Section \ref{subsec:direct-evidence}).  In increasing
     order along the $x$-axis, the models are the power-law of
-    Subsubsection \ref{subsubsec:power-law} (PL), the decaying
-    exponential of Subsubsection \ref{subsubsec:exponential} (E), the
-    single Gaussian of Subsection \ref{subsubsec:gaussian} (G), the
-    double Gaussian of Subsubsection \ref{subsubsec:gaussian} (TG),
+    Section \ref{subsubsec:power-law} (PL), the decaying
+    exponential of Section \ref{subsubsec:exponential} (E), the
+    single Gaussian of Section \ref{subsubsec:gaussian} (G), the
+    double Gaussian of Section \ref{subsubsec:gaussian} (TG),
     and the one-, two-, three-, four-, and five-bin histogram models
-    of Subsection \ref{subsec:non-parameteric-models} (H1, H2, H3, H4,
+    of Section \ref{subsec:non-parameteric-models} (H1, H2, H3, H4,
     H5, respectively).  The cross symbols give the value of the direct
@@
 \ref{sec:models}.  The counts for the number of samples in each model
-appear in Figure \ref{fig:rj}.  The most favored model is the power
-law from Subsubsection \ref{subsubsec:power-law}, followed by the one-
-and two-bin histograms from Subsection
+appear in Figure \ref{fig:rj}; relative probabilities of the models
+are given in Table \ref{tab:rj}.  The most favored model is the power
+law from Section \ref{subsubsec:power-law}, followed by the one-
+and two-bin histograms from Section
 \ref{subsec:non-parameteric-models}.  This confirms the results from
@@
 non-parameteric histograms would put it in second place.
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{|l|r|}
+      \hline
+      Model & Relative Probability \\
+      \hline \hline
+      Power Law (Section \ref{subsubsec:power-law}) & 0.441660 \\
+      \hline
+      Histogram (1 Bin, Section \ref{subsec:non-parameteric-models}) &
+      0.143435 \\
+      \hline
+      Histogram (2 Bin, Section \ref{subsec:non-parameteric-models}) &
+      0.135854 \\
+      \hline 
+      Exponential (Section \ref{subsubsec:exponential}) & 0.112514 \\
+      \hline
+      Histogram (3 Bin, Section \ref{subsec:non-parameteric-models}) &
+      0.075592 \\
+      \hline
+      Histogram (4 Bin, Section \ref{subsec:non-parameteric-models}) &
+      0.038839 \\
+      \hline 
+      Gaussian (Section \ref{subsubsec:gaussian}) & 0.022999 \\ 
+      \hline
+      Histogram (5 Bin, Section \ref{subsec:non-parameteric-models}) &
+      0.021011 \\
+      \hline      
+      Two Gaussian (Section \ref{subsubsec:gaussian}) & 0.008096 \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:rj} Relative probabilities of the various models
+    from Section \ref{sec:models} computed from reversible-jump
+    MCMC samples.}
+\end{table}
 
@@
     each model from Section \ref{sec:models} in the reversible-jump
-    MCMC of Subsection \ref{subsec:reversible-jump-mcmc}.  The models
-    are the power law of Subsubsection \ref{subsubsec:power-law} (PL),
-    the exponential of Subsubsection \ref{subsubsec:exponential} (E),
-    the Gaussian (G) and two Gaussian (TG) models of Subsubsection
+    MCMC of Section \ref{subsec:reversible-jump-mcmc}.  The models
+    are the power law of Section \ref{subsubsec:power-law} (PL),
+    the exponential of Section \ref{subsubsec:exponential} (E),
+    the Gaussian (G) and two Gaussian (TG) models of Section
     \ref{subsubsec:gaussian}, and the one-, two-, three-, four-, and
-    five-bin histograms of Subsection
+    five-bin histograms of Section
     \ref{subsec:non-parameteric-models} (H1, H2, H3, H4, H5,

### 2010-06-30 paper.tex
Typos in prior for Gaussian, Power-Law priors and Exponential likelihood.
Added Ilya to author list, grants in ack.

@@
 
-\author{Will M. Farr \and Niharika Sravan} \affil{Northwestern
-  University Center for Interdisciplinary Exploration and Research in
-  Astrophysics\\2145 Sheridan Rd., Evanston, IL 60208}
+\author{Will M. Farr \and Niharika Sravan \and Ilya Mandel}
+
+\affil{Northwestern University Center for Interdisciplinary
+  Exploration and Research in Astrophysics\\2145 Sheridan Rd.,
+  Evanston, IL 60208}
+
 \email{w-farr@northwestern.edu, niharika.sravan@gmail.com}
@@
 
-\author{Vicky Kalogera} \affil{Northwestern University Center for
-  Interdisciplinary Exploration and Research in Astrophysics\\2145
-  Sheridan Rd., Evanston, IL 60208} \email{vicky@northwestern.edu}
+\author{Vicky Kalogera}
+
+\affil{Northwestern University Center for Interdisciplinary
+  Exploration and Research in Astrophysics\\2145 Sheridan Rd.,
+  Evanston, IL 60208}
+
+ \email{vicky@northwestern.edu}
 
@@
   \begin{cases}
-    \frac{1}{2} \frac{1}{40^2}\frac{1}{20} & 0 \leq \Mmin \leq \Mmax
+    2 \frac{1}{40^2}\frac{1}{20} & 0 \leq \Mmin \leq \Mmax
     \leq 40, \quad -12 \leq \alpha \leq 8 \\
@@
     \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{m}{M_0}
-    \right] & M_0 \leq m \\
+    \right] & \Mmin \leq m \\
     0 & \textnormal{otherwise}
@@
   \begin{cases}
-    \frac{1}{2} \frac{1}{40^4} & 0 \leq \mu_1 \leq \mu_2 \leq 40,
+    2 \frac{1}{40^4} & 0 \leq \mu_1 \leq \mu_2 \leq 40,
     \quad 0 \leq \sigma_1, \sigma_2 \leq 40, \quad 0 \leq \alpha \leq
@@
 
-This work was supported by grant XXX.  WMF would like to thank Ilya
-Mandel for illuminating discussions and encouragement.
+This work was supported by grant \fixme{XXX}.  Calculations for this
+work were performed on the Northwestern Fugu cluster, which was
+partially funded by NSF MRI grant \fixme{XXX}.
 

### 2010-07-19 paper.tex
Removed paper.pdf, added Jon Gair in ack.
Modified mass of GS2023.
Added new ref for GS2023, commented out high-mass systems in table.
New high-mass data in Table 1 (caption not updated yet).

@@
 
-\author{Will M. Farr \and Niharika Sravan \and Ilya Mandel}
+\author{Will M. Farr \and Niharika Sravan}
 
@@
 
-\author{Vicky Kalogera}
+\author{Ilya Mandel \and Vicky Kalogera}
 
@@
 
- \email{vicky@northwestern.edu}
+ \email{ilyamandel@chgk.info, vicky@northwestern.edu}
 
@@
       \citet{Gelino2008} \\ & & & & \citet{Harlaftis2005} \\
-      Cyg X1 & $N(0.251, 0.007)$ & $N(2.778, 0.386)$ & $I(23, 38)$ &
-      \citet{Gies2003} \\ 
       XTE J1650 & $N(2.73, 0.56)$ & $U(0, 0.5)$ & $I(50, 80)$ &
@@
       \cite{Filippenko1999} \\
-      M33 X7 & $N(0.46, 0.08)$ & $N(4.47, 0.61)$ & $N(74.6, 1)$ &
-      \citet{Orosz2007} \\
       A0620 & $N(2.76, 0.036)$ & $N(0.06, 0.004)$ & $N(50.98, 0.87)$ &
@@
       \citet{Orosz2003} \\
-      GS 2023 & $N(6.08, 0.06)$ & $U(0.056,0.063)$ & $N(55,4)$ &
+      GS 2023 & $N(6.08, 0.06)$ & $U(0.056,0.063)$ & $I(66, 70)$ &
       \citet{Charles2006} \\
+      & & & & \citet{Khargharia2010} \\
       GS 1354 & $N(5.73, 0.29)$ & $N(0.12,0.04)$ & $I(50, 80)$ & 
@@
       \citet{Charles2006} \\
+      \hline \hline
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% High Mass Follows
+      Cyg X1 & $N(0.251, 0.007)$ & $N(2.778, 0.386)$ & $I(23, 38)$ &
+      \citet{Gies2003} \\ 
+      M33 X7 & $N(0.46, 0.08)$ & $N(4.47, 0.61)$ & $N(74.6, 1)$ &
+      \citet{Orosz2007} \\
+      NGC 300 X1 & $N(2.6, 0.3)$ & $U(1.05, 1.65)$ & $I(60, 75)$ &
+      \citet{Crowther2010} \\
+      LMC X1 & $N(0.148, 0.004)$ & $N(2.91, 0.49)$ & $N(36.38, 2.02)$
+      & \citet{Orosz2009} \\
+      IC 10 X1 & $N(7.64, 1.26)$ & $U(0.7, 1.7)$ & $I(75, 90)$ & 
+      \citet{Prestwich2007} \\
+       & & & & \citet{Silverman2008} \\
        \hline
@@
 
-In Figure \ref{fig:alpha}, we display a histogram of the resulting
+In %Figure \ref{fig:alpha}
+, we display a histogram of the resulting
 samples in $\alpha$; this represents the one-dimensional
@@
 
-\begin{figure}
-  \begin{center}
-    \plotone{plots/alpha}
-  \end{center}
-  \caption{\label{fig:alpha} A histogram of the MCMC samples for
-    $\alpha$ from the power-law distribution defined in Equation
-    \eqref{eq:power-law-dist}.  The histogram is an approximation to
-    the one-dimensional marginalized PDF in Equation
-    \eqref{eq:alpha-pdf}.  The favored value is near $\alpha = -4$.
-    The fraction of the distribution with $\alpha < 0$ is 98.7\%.  The
-    5\% quantile is $-7.66$, the median is $-4.28$, and the 95\%
-    quantile is $-1.36$.}
-\end{figure}
+% \begin{figure}
+%   \begin{center}
+%     \plotone{plots/alpha}
+%   \end{center}
+%   \caption{\label{fig:alpha} A histogram of the MCMC samples for
+%     $\alpha$ from the power-law distribution defined in Equation
+%     \eqref{eq:power-law-dist}.  The histogram is an approximation to
+%     the one-dimensional marginalized PDF in Equation
+%     \eqref{eq:alpha-pdf}.  The favored value is near $\alpha = -4$.
+%     The fraction of the distribution with $\alpha < 0$ is 98.7\%.  The
+%     5\% quantile is $-7.66$, the median is $-4.28$, and the 95\%
+%     quantile is $-1.36$.}
+% \end{figure}
 
@@
 
-Figure \ref{fig:exp-M0} displays the marginalized posterior
+%Figure \ref{fig:exp-M0}
+ displays the marginalized posterior
 distribution for the scale mass of the exponential, $M_0$.  The median
@@
 
-\begin{figure}
-  \begin{center}
-    \plotone{plots/exp-M0}
-  \end{center}
-  \caption{\label{fig:exp-M0} The distribution of scale masses, $M_0$,
-    measured in units of a solar mass for the exponential underlying
-    mass distribution defined in Equation \eqref{eq:exp-def}.  The
-    median scale mass is $M_0 = 3.00$; with 90\% confidence $1.57 \leq
-    M_0 \leq 5.71$. }
-\end{figure}
+% \begin{figure}
+%   \begin{center}
+%     \plotone{plots/exp-M0}
+%   \end{center}
+%   \caption{\label{fig:exp-M0} The distribution of scale masses, $M_0$,
+%     measured in units of a solar mass for the exponential underlying
+%     mass distribution defined in Equation \eqref{eq:exp-def}.  The
+%     median scale mass is $M_0 = 3.00$; with 90\% confidence $1.57 \leq
+%     M_0 \leq 5.71$. }
+% \end{figure}
 
@@
 work were performed on the Northwestern Fugu cluster, which was
-partially funded by NSF MRI grant \fixme{XXX}.
+partially funded by NSF MRI grant \fixme{XXX}.  We thank Jonathan Gair
+for helpful discussions.
 

### 2010-08-06 paper.tex
In-progress major edits, new plots Makefile, some new plots

@@
 \usepackage{hyperref}
+\usepackage{amssymb}
 \usepackage{amsmath}
@@
 \newcommand{\Nbin}{N_{\textnormal{bin}}}
+\newcommand{\vtheta}{\vec{\theta}}
 
@@
   We perform a Bayesian analysis of the mass distribution of
-  stellar-mass black holes based on the observed masses of the compact
-  object in 17 low mass X-ray binaries.  Modeling the mass
-  distribution both parametrically---as a power law, exponential,
-  gaussian, or combination of two gaussians---and
-  non-parametrically---as histograms with varying numbers of
-  bins---we explore the distribution in the context of Markov Chain
-  Monte Carlo calculations of model parameters.  We give confidence
-  bounds on the shape of the mass distribution in the context of each
-  model and compare the models with each other by calculating their
-  Bayesian evidence.  We address the existence of a ``gap'' between
+  stellar-mass black holes using the observed masses of 15 low-mass
+  X-ray binary systems and five higher-mass, wind-fed X-ray binary
+  systems.  Modeling the mass distribution both parametrically---as a
+  power law, exponential, gaussian, combination of two gaussians, or
+  log-normal distribution---and non-parametrically---as histograms
+  with varying numbers of bins---we explore the distribution in the
+  context of Markov Chain Monte Carlo calculations of model
+  parameters.  We give confidence bounds on the shape of the mass
+  distribution in the context of each model and compare the models
+  with each other by calculating their Bayesian evidence.  The mass
+  distribution of the low-mass systems is best fit by a power-law,
+  while the distribution of the complete sample is best fit by the
+  exponential model.  We address the existence of a ``gap'' between
   the most massive neutron stars and least massive black holes,
-  finding that the best model (the power law), the second-best model
-  (a one-bin histogram, or flat distribution), and the third-best
-  model (a two-bin histogram) have minimum black hole masses above
-  5.00, 3.23, and 3.82 solar masses (95\% confidence), respectively.
-  We therefore conclude that our sample of black hole masses provides
-  strong evidence of a gap between the maximum neutron star mass and
-  minimum black hole mass.
+  finding that the best model (the power law) fitted to the low-mass
+  systems only gives a minimum black-hole mass of 4.7 $\Msun$ (90\%
+  confidence), while the best model (the exponential) fitted to all 20
+  systems gives a minumum black-hole mass of 4.4 $\Msun$ (90\%
+  confidence).  We conclude that our sample of black hole masses
+  provides strong evidence of a gap between the maximum neutron star
+  mass and minimum black hole mass.
 \end{abstract}
@@
 
-The 17 low mass X-ray binary systems on which this study is based are
-listed in Table \ref{tab:sources}.  In each of these systems,
-spectroscopic measurements of the secondary star provide an orbital
-period for the system and a half-amplitude for the secondary's
-velocity curve.  These measurements can be combined into the mass
-function,
+The 20 X-ray binary systems on which this study is based are listed in
+Table \ref{tab:sources}.  We separate the systems into 15 low-mass
+systems in which the central black hole appears to be fed by
+roche-lobe overflow from the secondary, and 5 higher-mass systems in
+which the black hole is fed via winds (these systems all have a
+secondary that appears more massive than the black hole).  The low-
+and higher-mass systems undoubtedly have different evolutionary
+tracks, and therefore it is reasonable that they would have different
+black-hole mass distributions.  We will first analyze the 15 low-mass
+systems alone (Sections \ref{sec:models} and
+\ref{sec:model-selection}), and then the complete sample of 20 systems
+(Section \ref{sec:higher-mass}).
+
+In each of these systems, spectroscopic measurements of the secondary
+star provide an orbital period for the system and a half-amplitude for
+the secondary's velocity curve.  These measurements can be combined
+into the mass function,
 \begin{equation}
@@
 and mass ratio $q$ must be measured.  Ideally, this can be
-accomplished by fitting ellipsoidal light curves from these systems
-but even in the ideal case (see, e.g., \citet{Cantrell2010} on A0620)
-this procedure can be complicated by possible contributions from an
-accretion disk and hot spots in the disk.  For some systems (e.g.\ GS
-1354 \citep{Casares2009}) strong variability prevents determination of
-the inclination from the lightcurve at all; in these cases an upper
-limit on the inclination often comes from the observed lack of
-eclipses in the lightcurve.  In general, accurately determining $q$
-and $i$ requires a careful system-by-system analysis.
+accomplished by fitting ellipsoidal light curves and study of the
+rotational broadening of spectral lines from the secondary, but even
+in the ideal case (see, e.g., \citet{Cantrell2010} on A0620) this
+procedure is complicated.  In particular, contributions from an
+accretion disk and hot spots in the disk can significantly distort the
+measured inclination and mass ratios.  For some systems (e.g.\ GS 1354
+\citep{Casares2009}) strong variability prevents determination of the
+inclination from the lightcurve at all; in these cases an upper limit
+on the inclination often comes from the observed lack of eclipses in
+the lightcurve.  In general, accurately determining $q$ and $i$
+requires a careful system-by-system analysis.
 
@@
 attempt to deal with the systematic biases in the observational
-determination of $f$, $q$, and $i$ in any realistic way, instead
-leaving this for future work.
+determination of $f$, $q$, and $i$ in any realistic way; we are
+currently investigating more realistic treatments of the errors
+(including observational biases that can shift the peak of the true
+mass distribution away from the ``best-fit'' mass in the
+observations).  This careful treatment will appear in future work.
 
@@
   \end{center}
-  \caption{\label{tab:sources} The source parameters for the 17 low
-    mass X-ray binaries used in this work.  $f$ is the mass function for
+  \caption{\label{tab:sources} The source parameters for the 20 low
+    mass X-ray binaries used in this work.  The first 15 systems have
+    low-mass secondaries that feed the black hole via Roche lobe
+    overflow; the last five systems have high-mass secondaries ($q
+    \gtrsim 1$) that feed the black hole via winds.  In each line, $f$
+    is the mass function for
     the compact object, $q$ is the mass ratio $M_2/M$, and $i$ is the
@@
     $b$, and $I(\alpha,\beta)$ is an isotropic distribution between the
-    angles $\alpha$ and $\beta$.  \fixme{Orosz for XTE J1550 needs a
-      real reference \citep{Orosz2010}.}} 
+    angles $\alpha$ and $\beta$.}
 \end{table}
@@
 induced by the relationship in Equation \eqref{eq:mass-function}.
-Mass distributions generated in this way for the 17 systems used in
-this work are shown in Figure \ref{fig:all-masses}.  These mass
-distributions constitute the ``observational data'' we will use in
-Section \ref{sec:models} to explore various underlying mass
-distributions.
+Mass distributions generated in this way for the systems used in this
+work are shown in Figure \ref{fig:low-masses} and Figure
+\ref{fig:high-masses}.  Note that systems for which $i$ is poorly
+constrained have broad ``tails'' on their mass distributions.  These
+mass distributions constitute the ``observational data'' we will use
+in the remainder of this paper.
 
@@
 
-  \caption{\label{fig:all-masses} The mass distributions implied by
+  \caption{\label{fig:low-masses} The mass distributions implied by
     Equation \eqref{eq:mass-function} and the assumed distributions on
-    observational parameters given in Table \ref{tab:sources}.  The
-    significant asymmetry and long tails in many of these
+    observational parameters given in Table \ref{tab:sources} for the
+    lower-mass sources.  The significant asymmetry and long tails in
+    many of these distributions are the result of the non-linear
+    relationship (Equation \eqref{eq:mass-function}) between $M$, $f$,
+    $q$, and $i$.}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/high-masses}
+  \end{center}
+  \caption{\label{fig:high-masses} Mass distributions for the
+    wind-fed, higher-mass systems computed from the distributions on
+    observed data in Table \ref{tab:sources} using Equation
+    \eqref{eq:mass-function}.  The asymmetry and long tails in these
     distributions are the result of the non-linear relationship
-    (Equation \eqref{eq:mass-function}) between $M$, $f$, $q$, and
-    $i$.}
+    between $M$, $f$, $q$, and $i$.}
 \end{figure}
@@
 In Figure \ref{fig:masses} we show the combination of these individual
-mass distributions into one single distribution.  (Note that this
+mass distributions into one single distribution.  Note that this
 combination is not a good estimator of the true mass distribution of
-black hole masses---see, e.g., \citet{Mandel2010}.  We include it here
-only to give a sense of the range of masses of black holes in our
-sample; we will come to estimators for the mass distribution.)
+black hole masses.  To understand why this is not a good estimator of
+the underlying mass distribution, suppose that the observations
+implied an identical mass distribution for each object.  Then the
+combined distribution is equal to the individual distributions, but in
+fact we can be confident that each of the objects was drawn from a
+distribution that was narrower by a factor of approximately
+$\sqrt{N}$.  \citet{Mandel2010} discusses this issue in detail.  We
+include Figure \ref{fig:masses} here only to give a sense of the range
+of masses of black holes in our sample; we will come to estimators for
+the mass distribution.
 
 \begin{figure}
-  \plotone{plots/masses}
-
-  \caption{\label{fig:masses} The mass distributions in Figure
-    \ref{fig:all-masses} combined into a single distribution. Note
-    that this combination is not a good estimator of the true mass
-    distribution of black hole masses in the galaxy---see, e.g.,
-    \citet{Mandel2010}.  We include it here only to give a sense of
-    the range and general shape of the mass distribution of black
-    holes in our sample.}
+  \plotone{plots/masses-high}
+
+  \caption{\label{fig:masses} The mass distributions in Figures
+    \ref{fig:low-masses} and \ref{fig:high-masses} combined into a
+    single distribution. Note that this combination is not a good
+    estimator of the true mass distribution of black hole masses in
+    the galaxy---see, e.g., \citet{Mandel2010}.  We include it here
+    only to give a sense of the range and general shape of the mass
+    distribution of black holes in our sample.}
 \end{figure}
@@
 In this section we apply a Bayesian analysis to various models for the
-underlying mass distribution from which the systems in Table
-\ref{tab:sources} were drawn.  The end result will be the probability
-distribution for the parameters of each model implied by the data from
-Section \ref{sec:systems} in combination with our prior assumptions
-about the probability distribution for the parameters.  Bayes' rule
-relates these quantities.  For a model with parameters $\theta$ in the
-presence of data $d$, Bayes' rule states
+underlying mass distribution from which the low-mass systems in Table
+\ref{tab:sources} were drawn.  We will treat the combined low- and
+higher-mass systems in Section \ref{sec:higher-mass}.  The end result
+will be the probability distribution for the parameters of each model
+implied by the data from Section \ref{sec:systems} in combination with
+our prior assumptions about the probability distribution for the
+parameters.  Bayes' rule relates these quantities.  For a model with
+parameters $\vtheta$ in the presence of data $d$, Bayes' rule states
 \begin{equation}
   \label{eq:Bayes-rule}
-  p(\theta | d) = \frac{p(d | \theta) p(\theta)}{p(d)}.
-\end{equation}
-Here, $p(\theta|d)$, called the posterior probability distribution
-function, is the probability distribution for the parameters $\theta$
-implied by the data $d$; $p(d|\theta)$, called the likelihood, is the
+  p(\vtheta | d) = \frac{p(d | \vtheta) p(\vtheta)}{p(d)}.
+\end{equation}
+Here, $p(\vtheta|d)$, called the posterior probability distribution
+function, is the probability distribution for the parameters $\vtheta$
+implied by the data $d$; $p(d|\vtheta)$, called the likelihood, is the
 probability of observing data $d$ given that the model parameters are
-$\theta$; $p(\theta)$, called the prior, reflects our estimate of the
+$\vtheta$; $p(\vtheta)$, called the prior, reflects our estimate of the
 probability of the various model parameters in the absence of any
@@
 \begin{equation}
-  \int d\theta\, p(\theta|d) = 1,
+  \int d\theta\, p(\vtheta|d) = 1,
 \end{equation}
@@
   \label{eq:evidence-def}
-  p(d) = \int d\theta\, p(d|\theta) p(\theta).
+  p(d) = \int d\vtheta\, p(d|\vtheta) p(\vtheta).
 \end{equation}
@@
 In our context, the data are the mass distributions given in Section
-\ref{sec:systems}: $d = \{ p_i(m)| i = 1, 2, \ldots, 17 \}$.  We
+\ref{sec:systems}: $d = \{ p_i(M)| i = 1, 2, \ldots, 17 \}$.  We
 assume that the measurements in Section \ref{sec:systems} are
@@
 likelihoods for the individual measurements.  For a model with
-parameters $\theta$ that predicts a mass distribution $p(m|\theta)$
+parameters $\vtheta$ that predicts a mass distribution $p(M|\vtheta)$
 for black holes, we have
@@
   \label{eq:likelihood-def}
-  p(d|\theta) = \prod_i \int dm\, p_i(m) p(m|\theta).
+  p(d|\vtheta) = \prod_i \int dM\, p_i(M) p(M|\vtheta).
 \end{equation}
 That is, the likelihood of an observation is the average over the
-individual mass distribution implied by the observation, $p_i(m)$, of
+individual mass distribution implied by the observation, $p_i(M)$, of
 the probability for a black hole of that mass to exist according to
-the model of the mass distribution, $p(m | \theta)$.  We approximate
-the integrals as averages of $p(m|\theta)$ over the Monte Carlo mass
+the model of the mass distribution, $p(M | \vtheta)$.  We approximate
+the integrals as averages of $p(M|\vtheta)$ over the Monte Carlo mass
 samples drawn from the distributions in Table \ref{tab:sources} (also
-see Figure \ref{fig:all-masses}):
-\begin{equation}
-  p(d|\theta) \approx \prod_i \frac{1}{N_i} \sum_{j = 1}^{N_i} p(m_{ij} | \theta),
-\end{equation}
-where $m_{ij}$ is the $j$th sample (out of a total $N_i$) from the
+see Figures \ref{fig:low-masses} and \ref{fig:high-masses}):
+\begin{equation}
+  p(d|\vtheta) \approx \prod_i \frac{1}{N_i} \sum_{j = 1}^{N_i} p(M_{ij} | \vtheta),
+\end{equation}
+where $M_{ij}$ is the $j$th sample (out of a total $N_i$) from the
 $i$th individual mass distribution.
@@
 observed, leaving the analysis of selection effects to future work.
-
-For a mass distribution with several parameters, $p(\theta | d)$ lives
-in a multi-dimensional space.  Exploring the entirety of this space
-systematically rapidly becomes prohibitive as the number of parameters
-increases.  A more efficient way to explore the distribution $p(\theta
-| d)$ is to use a Markov Chain Monte Carlo (MCMC) method.  MCMC
-methods produce a chain (sequence) of parameter samples, $\{ \theta_i,
-i = 1, \ldots \}$, such that a particular parameter, $\theta$, appears
-in the sequence with a frequency equal to its probability according to
-$p(\theta|d)$.  In this way, regions of parameter space where
-$p(\theta|d)$ is large are sampled densely while regions where
-$p(\theta|d)$ is small are effectively ignored.  A Markov chain has
-the property that the transition probability from one element to the
-next, $p(\theta_i \to \theta_{i+1})$, depends only on the value of
-$\theta_i$, not on any previous values in the chain.
+(During the preparation of this work, we became aware of
+\citet{Ozel2010}, which finds that selection effects do not greatly
+affect an analysis similar to ours.)
+
+For a mass distribution with several parameters, $p(\vtheta | d)$
+lives in a multi-dimensional space.  Exploring the entirety of this
+space systematically rapidly becomes prohibitive as the number of
+parameters increases.  A more efficient way to explore the
+distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
+(MCMC) method.  MCMC methods produce a chain (sequence) of parameter
+samples, $\{ \vtheta_i, i = 1, \ldots \}$, such that a particular
+parameter set, $\vtheta$, appears in the sequence with a frequency
+equal to its probability according to $p(\vtheta|d)$.  In this way,
+regions of parameter space where $p(\vtheta|d)$ is large are sampled
+densely while regions where $p(\vtheta|d)$ is small are effectively
+ignored.  A Markov chain has the property that the transition
+probability from one element to the next, $p(\vtheta_i \to
+\vtheta_{i+1})$, depends only on the value of $\vtheta_i$, not on any
+previous values in the chain.
 
 One way to produce a sequence of MCMC samples is via the following
-algorithm, first proposed by \fixme{Metropolis citation needed}:
+algorithm, first proposed by \citet{Metropolis1953}:
 \begin{enumerate}
-  \item Begin with the current sample, $\theta_i$.
-  \item Propose a new sample, $\theta_p$, by drawing randomly from an
-    arbitrary ``jump proposal distribution'' with probability
-    $Q(\theta_i \to \theta_p)$.
+  \item Begin with the current sample, $\vtheta_i$.
+  \item Propose a new sample, $\vtheta_p$, by drawing randomly from a
+    ``jump proposal distribution'' with probability $Q(\vtheta_i \to
+    \vtheta_p)$.  Note that $Q(\vtheta_i \to \vtheta_p)$ can depend on
+    the current parameters, $\vtheta_i$, and any other ``constant''
+    data, but cannot examine the history of the chain beyond the most
+    recent point.  This is necessary to preserve the Markovian
+    property of the chain.
   \item Compute the ``acceptance'' probability,
@@
       p_{\textnormal{accept}} \equiv
-      \frac{p(\theta_p|d)}{p(\theta_i|d)} \frac{Q(\theta_p \to
-        \theta_i)}{Q(\theta_i \to \theta_p)}
-    \end{equation}
-  \item With probability $\min(1,p_{\textnormal{accept}})$ set
-    $\theta_{i+1} = \theta_p$; otherwise set $\theta_{i+1} =
-    \theta_i$.
+      \frac{p(\vtheta_p|d)}{p(\vtheta_i|d)} \frac{Q(\vtheta_p \to
+        \vtheta_i)}{Q(\vtheta_i \to \vtheta_p)}
+    \end{equation}
+  \item With probability $\min(1,p_{\textnormal{accept}})$ ``accept''
+    the propsed $\vtheta_p$, setting $\vtheta_{i+1} = \vtheta_p$;
+    otherwise set $\vtheta_{i+1} = \vtheta_i$.
 \end{enumerate}
-Note that this algorithm is more likely to accept a proposed jump when
-it increases the posterior (the first term in Equation
+This algorithm is more likely to accept a proposed jump when it
+increases the posterior (the first factor in Equation
 \eqref{eq:paccept}) and when it is to a location in parameter space
-from which it is easy to return (the second term in Equation
-\eqref{eq:paccept}).  As $i \to \infty$ the samples $\theta_i$ are
-distributed according to $p(\theta|d)$.  In practice the number of
-samples required before the chain appropriately samples $p(\theta|d)$
-depends strongly on the jump proposal distribution; proposal
-distributions that often propose jumps toward or within regions of
-large $p(\theta|d)$ can be very efficient, while poor proposal
-distributions can require prohibitively large numbers of samples
-before convergence.
-
-Once we have a chain of samples from $p(\theta|d)$---a distribution of
-distributions---the distribution for any quantity of interest can be
-computed.  For example, Figure \ref{fig:dists} shows the median
-values, and the 10\%, and 90\% quantiles for value of the BH mass
-probability distribution at various masses in each of the models
-discussed in the following subsections.  Such a figure can be
-generated from the MCMC output by computing the value of the
-parameterized BH mass distribution at various masses for each
+from which it is easy to return (the second factor in Equation
+\eqref{eq:paccept}); the combination of these influences in Equation
+\eqref{eq:paccept} ensures that the equilibrium distribution of the
+chain is $p(\vtheta|d)$.  As $i \to \infty$ the samples $\vtheta_i$
+are distributed according to $p(\vtheta|d)$.
+
+In practice the number of samples required before the chain
+appropriately samples $p(\vtheta|d)$ depends strongly on the jump
+proposal distribution; proposal distributions that often propose jumps
+toward or within regions of large $p(\vtheta|d)$ can be very
+efficient, while poor proposal distributions can require prohibitively
+large numbers of samples before convergence.
+
+There is no foolproof test for the convergence of a chain.  In this
+work, we test the convergence of our chains by comparing the
+statistics calculated from the entire chain to statistics calculated
+from only the first half of the chain; when the chain has converged,
+the two calculations agree.  This is a necessary, but not sufficient,
+condition for convergence.
+
+We begin the chain at an arbitrary point in parameter space; this is
+equivalent to taking a finite section of an infinite chain that begins
+with the chosen point.  Since every point in parameter space occurs in
+an infinite chain, and no section of the chain is better than any
+other, our chosen finite chain will have the appropriate distribution.
+However, because consecutive samples in a chain are correlated with
+each other, the beginning of our finite chain has a ``memory'' of the
+starting point; we discard enough points at the beginning of the
+finite chain that we can be confident it does not retain a memory of
+the arbitrary starting point.  The points discarded in this way are
+commonly called ``burn-in'' points.
+
+Once we have a chain of samples from $p(\vtheta|d)$---a distribution
+of distributions---the distribution for any quantity of interest can
+be computed.  For example, Figure \ref{fig:dists} shows the median
+values, and the 10\% and 90\% quantiles for the value of the BH mass
+probability distribution at various masses.  (Recall that we are only
+using the 15 low-mass observations in this Section's analysis.)  Such
+a figure can be generated from the MCMC output by computing the value
+of the parameterized BH mass distribution at various masses for each
 parameter sample output by the MCMC, and then plotting the quantiles
@@
   \caption{\label{fig:dists} The median values of the black hole mass
-    distribution, $p(m|\theta)$, at various masses implied by the
+    distribution, $p(M|\theta)$, at various masses implied by the
     posterior $p(\theta|d)$ for the models discussed in Sections
     \ref{subsec:parameteric-models} and
-    \ref{subsec:non-parameteric-models}.  Error bars span the 10\% to
-    90\% range.  Note that these ``distributions of distributions''
-    are not necessarily normalized, and need not be ``shaped'' like
-    the underlying model distributions.}
+    \ref{subsec:non-parameteric-models}.  These distributions use only
+    the 15 low-mass observations in Table \ref{tab:sources} (the
+    complete sample is analyzed in Section \ref{sec:higher-mass}).
+    Error bars span the 10\% to 90\% range.  Note that these
+    ``distributions of distributions'' are not necessarily normalized,
+    and need not be ``shaped'' like the underlying model
+    distributions.}
 \end{figure}
@@
 
-We defer discussion of the appropriateness of the different models
-given the data (i.e.\ the ``goodness of fit'') until Section
-\ref{sec:model-selection}.
+\subsection{Priors}
+
+An important part of any Bayesian analysis is the priors placed on the
+parameters of the model.  The choice of priors can bias the results of
+the analysis through both the shape and the range of prior support in
+parameter space.
+
+A prior that is independent of parameters, $\vtheta$, in some region,
+called ``flat,'' results in a posterior that is proportional to the
+likelihood (see Equation \eqref{eq:Bayes-rule}).  A flat prior does
+not change the shape of the posterior.  However, the choice of a flat
+prior is parameterization-dependent: a change of parameter from
+$\vtheta$ to $\vtheta' = \vec{f}(\vtheta)$ can change a flat
+distribution into one with non-trivial structure.  In this work, we
+choose priors that are flat when the parameters are measured in
+physical units.  In particular, for the log-normal model (Section
+\ref{sec:log-normal}) the natural parameters for the distribution are
+the mean, $\langle \log M \rangle$, and standard deviation,
+$\sigma_{\log M}$, in $\log M$, but we choose priors that are flat in
+$\langle M \rangle$ and $\sigma_M$.
+
+The range of prior support can also affect the results of a Bayesian
+analysis.  Because priors are normalized, prior support over a larger
+region of parameter space results in a smaller prior probability at
+each point.  Such ``wide'' priors are implicitly claiming that any
+particular sub-region of parameter space is less likely than it would
+be under a prior of the same shape but smaller support volume.  This
+difference is important in model selection (Section
+\ref{sec:model-selection}): when comparing two models with the same
+likelihood, one with wide priors will seem less probable than one with
+narrower priors.  Of course, priors should be wide enough to encompass
+all regions of parameter space that have significant likelihood.  To
+make the model comparison in Section \ref{sec:model-selection} fair,
+we choose prior support in parameter space so that the allowed
+parameter values for each model give distributions for which nearly
+all the probability lies in the range $0 \leq M \leq 40 \Msun$.
 
@@
 underlying black hole mass distribution that we have analyzed using
-the data from Section \ref{sec:systems}.
+the low-mass data from Section \ref{sec:systems}.  (Analysis of these
+models including the higher-mass systems appears in Section
+\ref{sec:higher-mass}.)
 
@@
   \label{eq:power-law-dist}
-  p(m|\theta) = p(m|\{\Mmin, \Mmax, \alpha \}) =
+  p(M|\vtheta) = p(M|\{\Mmin, \Mmax, \alpha \}) =
   \begin{cases}
-    A m^\alpha & \Mmin \leq m \leq \Mmax \\
+    A M^\alpha & \Mmin \leq m \leq \Mmax \\
     0 & \textnormal{otherwise}
@@
 \end{equation}
-To perform the Bayesian analysis, we need to impose a prior on the
-parameters $\Mmin$, $\Mmax$, $\alpha$.  We choose to impose a broad
-prior that represents our significant uncertainty in the appropriate
-values of these parameters.  We use
-\begin{equation}
-  p(\theta) = p(\{\Mmin, \Mmax, \alpha\}) = 
+We choose uniform priors on $\Mmin$ and $\Mmax \geq \Mmin$ between 0 and
+$40 \Msun$, and uniform priors on the exponent $\alpha$ in a broad
+range between $-15$ and $13$:
+\begin{equation}
+  p(\vtheta) = p(\{\Mmin, \Mmax, \alpha\}) = 
   \begin{cases}
-    2 \frac{1}{40^2}\frac{1}{20} & 0 \leq \Mmin \leq \Mmax
-    \leq 40, \quad -12 \leq \alpha \leq 8 \\
+    2 \frac{1}{40^2}\frac{1}{28} & 0 \leq \Mmin \leq \Mmax
+    \leq 40, \quad -15 \leq \alpha \leq 13 \\
     0 & \textnormal{otherwise}
-  \end{cases},
-\end{equation}
-where the masses are measured in solar masses.  This prior is uniform
-in the masses between 0 and $40 \Msun$, and uniform in power-law slope
-between $-12$ and $8$.
+  \end{cases}.
+\end{equation}
 
@@
 \begin{equation}
-  p(\theta|d) = p(\{\Mmin, \Mmax, \alpha\}|d) \propto p(d|\{\Mmin,
+  p(\vtheta|d) = p(\{\Mmin, \Mmax, \alpha\}|d) \propto p(d|\{\Mmin,
   \Mmax, \alpha\}) p(\{\Mmin, \Mmax, \alpha\}),
@@
 the resulting distribution of mass distributions inferred from the
-MCMC distribution of parameters.
-
-In %Figure \ref{fig:alpha}
-, we display a histogram of the resulting
-samples in $\alpha$; this represents the one-dimensional
-``marginalized'' distribution
+parameters output by the MCMC.
+
+In Figure \ref{fig:power-law} , we display a histogram of the resulting
+samples in each of the parameters $\Mmin$, $\Mmax$, and $\alpha$
+individually; this represents the one-dimensional ``marginalized''
+distribution
 \begin{equation}
   \label{eq:alpha-pdf}
-  p(\alpha|d) = \int d\Mmin\, d\Mmax\, p(\{\Mmin, \Mmax, \alpha\}|d).
-\end{equation}
-The resulting distribution is quite broad with 
-\begin{equation}
-  -7.66 < \alpha < -1.36
-\end{equation}
-enclosing 90\% of the probability.  The median value is $\alpha =
--4.28$, and we have $\alpha < 0$ (a falling underlying mass
-distribution) at 98.7\% confidence.  The distribution of the minimum
+  p(\alpha|d) = \int d\Mmin\, d\Mmax\, p(\{\Mmin, \Mmax, \alpha\}|d),
+\end{equation}
+and similarly for $\Mmin$ and $\Mmax$.
+
+The marginalized distribution for $\alpha$ is broad, with
+\begin{equation}
+  -11.8 < \alpha < 6.8
+\end{equation}
+enclosing 90\% of the probability.  We have $p(\alpha < 0) = 0.6$.
+The median value is $\alpha = -3.35$.  The distribution of the minimum
 mass is also quite interesting, but we defer this discussion until
@@
 
-% \begin{figure}
-%   \begin{center}
-%     \plotone{plots/alpha}
-%   \end{center}
-%   \caption{\label{fig:alpha} A histogram of the MCMC samples for
-%     $\alpha$ from the power-law distribution defined in Equation
-%     \eqref{eq:power-law-dist}.  The histogram is an approximation to
-%     the one-dimensional marginalized PDF in Equation
-%     \eqref{eq:alpha-pdf}.  The favored value is near $\alpha = -4$.
-%     The fraction of the distribution with $\alpha < 0$ is 98.7\%.  The
-%     5\% quantile is $-7.66$, the median is $-4.28$, and the 95\%
-%     quantile is $-1.36$.}
-% \end{figure}
+\begin{figure}
+  \begin{center}
+    \plotone{plots/power-law}
+  \end{center}
+  \caption{\label{fig:power-law} A histogram of the MCMC samples for
+    $\alpha$ from the power-law distribution defined in Equation
+    \eqref{eq:power-law-dist}.  The histogram is an approximation to
+    the one-dimensional marginalized PDF in Equation
+    \eqref{eq:alpha-pdf}.  The favored value is near $\alpha = -4$.
+    The fraction of the distribution with $\alpha < 0$ is 98.7\%.  The
+    5\% quantile is $-7.66$, the median is $-4.28$, and the 95\%
+    quantile is $-1.36$.}
+\end{figure}
 
@@
 two-Gaussian model.
+
+\subsubsection{Log Normal}
+\label{sec:log-normal}
 
@@
 
+\section{Including High-Mass, Wind-Fed Systems}
+\label{sec:higher-mass}
+
 \section{The Minimum Mass of the Black Hole Mass Distribution}

### 2010-08-06 plots/doepstopdf.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-
-set -e
-
-for file in *.eps; do
-    epstopdf $file
-done

### 2010-10-20 paper.tex
As submitted to collaborators.

@@
 
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{amssymb}
@@
 \newcommand{\vtheta}{\vec{\theta}}
+\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
 
@@
   stellar-mass black holes using the observed masses of 15 low-mass
-  X-ray binary systems and five higher-mass, wind-fed X-ray binary
-  systems.  Modeling the mass distribution both parametrically---as a
-  power law, exponential, gaussian, combination of two gaussians, or
-  log-normal distribution---and non-parametrically---as histograms
-  with varying numbers of bins---we explore the distribution in the
-  context of Markov Chain Monte Carlo calculations of model
-  parameters.  We give confidence bounds on the shape of the mass
-  distribution in the context of each model and compare the models
-  with each other by calculating their Bayesian evidence.  The mass
-  distribution of the low-mass systems is best fit by a power-law,
-  while the distribution of the complete sample is best fit by the
-  exponential model.  We address the existence of a ``gap'' between
+  X-ray binary systems undergoing Roche lobe overflow and five
+  higher-mass, wind-fed X-ray binary systems.  Modeling the mass
+  distribution both parametrically---as a power law, exponential,
+  gaussian, combination of two gaussians, or log-normal
+  distribution---and non-parametrically---as histograms with varying
+  numbers of bins---we explore the distribution in the context of
+  Markov Chain Monte Carlo calculations of model parameters.  We give
+  confidence bounds on the shape of the mass distribution in the
+  context of each model and compare the models with each other by
+  calculating their Bayesian evidence.  The mass distribution of the
+  low-mass systems is best fit by a power-law, while the distribution
+  of the complete sample is best fit by the exponential and power-law
+  models (but with significantly different parameters than the
+  lower-mass sample).  We address the existence of a ``gap'' between
   the most massive neutron stars and least massive black holes,
   finding that the best model (the power law) fitted to the low-mass
-  systems only gives a minimum black-hole mass of 4.7 $\Msun$ (90\%
-  confidence), while the best model (the exponential) fitted to all 20
-  systems gives a minumum black-hole mass of 4.4 $\Msun$ (90\%
-  confidence).  We conclude that our sample of black hole masses
-  provides strong evidence of a gap between the maximum neutron star
-  mass and minimum black hole mass.
+  systems only gives a minimum black-hole mass of 4.3 $\Msun$ (90\%
+  confidence), while the best models (the exponential and power-law)
+  fitted to all 20 systems gives a minumum black-hole mass of 4.5 and
+  5.2 $\Msun$ (90\% confidence), respectively.  We conclude that our
+  sample of black hole masses provides strong evidence of a gap
+  between the maximum neutron star mass and minimum black hole mass.
 \end{abstract}
@@
 \begin{figure}
-  \plotone{plots/masses-high}
-
-  \caption{\label{fig:masses} The mass distributions in Figures
-    \ref{fig:low-masses} and \ref{fig:high-masses} combined into a
-    single distribution. Note that this combination is not a good
-    estimator of the true mass distribution of black hole masses in
-    the galaxy---see, e.g., \citet{Mandel2010}.  We include it here
-    only to give a sense of the range and general shape of the mass
+  \plottwo{plots/masses}{plots/masses-high}
+
+  \caption{\label{fig:masses} The mass distributions in Figure
+    \ref{fig:low-masses} (left) and both \ref{fig:low-masses} and
+    \ref{fig:high-masses} (right) combined into a single
+    distribution. Note that this combination is not a good estimator
+    of the true mass distribution of black hole masses in the
+    galaxy---see, e.g., \citet{Mandel2010}.  We include it here only
+    to give a sense of the range and general shape of the mass
     distribution of black holes in our sample.}
@@
   \begin{center}
-    \plottwo{plots/dist-parameteric}{plots/dist-non-parameteric}
+    \plottwo{plots/dist-parametric}{plots/dist-non-parametric}
   \end{center}
@@
     posterior $p(\theta|d)$ for the models discussed in Sections
-    \ref{subsec:parameteric-models} and
-    \ref{subsec:non-parameteric-models}.  These distributions use only
+    \ref{sec:parametric-models} and
+    \ref{sec:non-parametric-models}.  These distributions use only
     the 15 low-mass observations in Table \ref{tab:sources} (the
@@
 \subsection{Priors}
+\label{sec:priors}
 
@@
 
-\subsection{Parameteric Models for the Black Hole Mass Distribution}
-\label{subsec:parameteric-models}
-
-In this subsection, we discuss the various parameteric models of the
+\subsection{Parametric Models for the Black Hole Mass Distribution}
+\label{sec:parametric-models}
+
+In this subsection, we discuss the various parametric models of the
 underlying black hole mass distribution that we have analyzed using
@@
 models including the higher-mass systems appears in Section
-\ref{sec:higher-mass}.)
+\ref{sec:higher-mass}.)  Table \ref{tab:low-mass-parametric} gives
+quantiles of the marginalized parameter distributions of the
+parametric models.
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{|l|c|c|c|c|c|c|}
+      \hline
+      Model & Parameter & 5\% & 15\% & 50\% & 85\% & 95\% \\
+      \hline \hline
+      Power Law (Equation \eqref{eq:power-law-dist}) & $\Mmin$ & 
+      1.2786 &  4.1831 &  6.1001 &  6.5011 &  6.6250 \\
+      \hline
+       & $\Mmax$ & 8.5578 &  8.9214 & 23.3274 & 36.0002 & 38.8113 \\
+       \hline
+       & $\alpha$ & -12.4191 & -10.1894 & -6.3861 &  2.8476 &  5.6954 \\
+       \hline \hline
+       Decaying Exponential (Equation \eqref{eq:exp-def}) & $\Mmin$ & 
+       5.0185 &  5.4439 &  6.0313 &  6.3785 &  6.5316 \\
+       \hline
+       & $M_0$ & 0.7796 &  0.9971 & 1.5516 &  2.4635 &  3.2518 \\
+       \hline \hline
+       Gaussian (Equation \eqref{eq:gaussian-def}) & $\mu$ & 
+       6.6349 &  6.9130 &  7.3475 & 7.7845 & 8.0798 \\
+       \hline
+       & $\sigma$ & 0.7478 &  0.9050  & 1.2500 &  1.7335 & 2.1134 \\
+       \hline \hline
+       Two Gaussian (Equation \eqref{eq:two-gaussian-def}) & $\mu_1$ & 
+       5.4506 &  6.3877 &  7.1514 &  7.6728  & 7.9803 \\
+       \hline
+       & $\mu_2$ & 7.2355 &  7.7387 & 12.3986 & 25.2456 & 31.4216 \\
+       \hline
+       & $\sigma_1$ & 0.3758 &  0.7626 &  1.2104 &  1.7981 &  2.3065 \\
+       \hline
+       & $\sigma_2$ & 0.2048 & 0.6421 & 1.9182 &  5.2757  & 7.2625 \\
+       \hline
+       & $\alpha$ & 0.0983 &  0.3526 & 0.8871 &  0.9792 &  0.9936 \\
+       \hline \hline
+       Log Normal (Equation \eqref{eq:log-normal-def}) & $\langle M \rangle$ & 
+       6.7619 &  7.0122 &  7.4336  &  7.9159  &  8.2942 \\
+       \hline 
+       & $\sigma_M$ & 0.7292  &  0.8920  & 1.2704  &  1.8695  &  2.4069 \\
+       \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:low-mass-parametric} Quantiles of the
+    marginalized distribution for each of the parameters in the models
+    discussed in Section \ref{sec:parametric-models}.  We indicate
+    the 5\%, 15\%, 50\% (median), 85\%, and 95\% quantiles.  The
+    marginalized distribution can be misleading when there are strong
+    correlations between variables.  For example, while the
+    marginalized distributions for the power law parameters are quite
+    broad, the distribution of mass distributions implied by the power
+    law MCMC samples is similar to the other models.  This occurs in
+    spite of the broad marginalized distributions because of the
+    correlations between the slope and limits of the power law
+    discussed in Section \ref{sec:power-law}.}
+\end{table}
 
 \subsubsection{Power-Law Models}
-\label{subsubsec:power-law}
+\label{sec:power-law}
 
@@
 
+\begin{figure}
+  \begin{center}
+    \plotone{plots/power-law}
+  \end{center}
+  \caption{\label{fig:power-law} A histogram of the marginalized
+    distribution for the three parameters $\Mmin$, $\Mmax$, and
+    $\alpha$ from the power-law model.  The marginalized distribution
+    for $\alpha$ is broad, with $-11.8 < \alpha < 6.8$ enclosing 90\%
+    of the probability.  We have $p(\alpha < 0) = 0.6$; the median
+    value is $\alpha = -3.35$.  The broad distribution for $\alpha$
+    (and the other parameters) is due to correlations between the
+    parameters discussed in the main text; see Figure
+    \ref{fig:power-law-2D}.}
+\end{figure}
+
 The marginalized distribution for $\alpha$ is broad, with
@@
 \end{equation}
-enclosing 90\% of the probability.  We have $p(\alpha < 0) = 0.6$.
-The median value is $\alpha = -3.35$.  The distribution of the minimum
-mass is also quite interesting, but we defer this discussion until
-Section \ref{sec:minimum-mass}, where we discuss the distribution for
-the minimum black hole masses in the context of all our models.
-
-\begin{figure}
-  \begin{center}
-    \plotone{plots/power-law}
-  \end{center}
-  \caption{\label{fig:power-law} A histogram of the MCMC samples for
-    $\alpha$ from the power-law distribution defined in Equation
-    \eqref{eq:power-law-dist}.  The histogram is an approximation to
-    the one-dimensional marginalized PDF in Equation
-    \eqref{eq:alpha-pdf}.  The favored value is near $\alpha = -4$.
-    The fraction of the distribution with $\alpha < 0$ is 98.7\%.  The
-    5\% quantile is $-7.66$, the median is $-4.28$, and the 95\%
-    quantile is $-1.36$.}
+enclosing 90\% of the probability (excluding 5\% on each side).  We
+have $p(\alpha < 0) = 0.6$.  The median value is $\alpha = -3.35$.
+The broadness of the marginalized distribution for $\alpha$ comes from
+the need to match the relatively narrow peak in the mass distribution
+apparent in Figure \ref{fig:masses}.  When $\alpha$ is negative, the
+resulting mass distribution slopes down; $\Mmin$ is constrained to be
+near the lowest mass of the observed black holes, while $\Mmax$ is
+essentially irrelevant.  Conversely, when $\alpha$ is positive and the
+mass distribution slopes up, $\Mmax$ must be close to the largest mass
+observed, while $\Mmin$ is essentially irrelevant.  Figure
+\ref{fig:power-law-2D} illustrates this effect, showing the
+correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
+When we include the higher-mass systems in the analysis, the long tail
+will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
+play for all values of $\alpha$.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/power-law-2D}
+  \end{center}
+  \caption{\label{fig:power-law-2D} MCMC samples in the $\Mmin,
+    \alpha$ (top) and $\Mmax, \alpha$ (bottom) planes for the
+    power-law model discussed in Section \ref{sec:power-law}.  The
+    correlations between $\alpha$ and the power law bounds discussed
+    in the text are apparent: when $\alpha$ is positive, the mass
+    distribution slopes upward and $\Mmax$ is constrained to be near
+    the maximum observed mass while $\Mmin$ is unconstrained.  When
+    $\alpha$ is negative, the mass distribution slopes down and
+    $\Mmin$ is constrained to be near the lowest mass observed, while
+    $\Mmax$ is unconstrained. }
 \end{figure}
@@
 \subsubsection{Decaying Exponential}
-\label{subsubsec:exponential}
-
-Theoretical arguments in \citet{Fryer2001} suggest that the black-hole
-mass distribution may be well-represented by a decaying exponential
-with a minimum mass:
+\label{sec:exponential}
+
+\citet{Fryer2001} suggest that the black-hole mass distribution may be
+well-represented by a decaying exponential with a minimum mass:
 \begin{equation}
   \label{eq:exp-def}
-  p(m|\theta) = p(m|\{\Mmin, M_0\}) = 
+  p(M|\vtheta) = p(M|\{\Mmin, M_0\}) = 
   \begin{cases}
-    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{m}{M_0}
-    \right] & \Mmin \leq m \\
+    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
+    \right] & \Mmin \leq M \\
     0 & \textnormal{otherwise}
@@
 \end{equation}
-We choose again a broad, flat prior
-\begin{equation}
-  p(\{\Mmin, M_0 \}) = 
+We choose a prior for this model where $\Mmin$ is uniform between 0
+and $40 \Msun$.  For each $\Mmin$, we choose $M_0$ uniformly within a
+range ensuring that $40\Msun$ is at least two scale masses above the
+cutoff: $40\Msun \geq \Mmin + 2 M_0$.  This ensures that the majority
+of the mass probability lies in the range $0 \leq M \leq 40\Msun$.
+The resulting prior is
+\begin{equation}
+  p(\vtheta) = p(\{\Mmin, M_0\}) = 
   \begin{cases}
-    \frac{1}{40^2} & 0 \leq \Mmin, M_0 \leq 40 \\
+    \frac{4}{40^2} & 0 \leq \Mmin \leq 40, \quad 0 < M_0, \quad \Mmin
+    + 2 M_0 \leq 40, \\
+    0 & \textnormal{otherwise}
+  \end{cases}
+\end{equation}
+
+Figure \ref{fig:exp-marginal} displays the marginalized posterior
+distribution for the scale mass of the exponential, $M_0$, and the
+cutoff mass, $\Mmin$.  The median scale mass is $M_0 = 1.55$, and
+$0.78 \leq M_0 \leq 3.25$ with 90\% confidence.  Figure
+\ref{fig:exp-2D} displays the MCMC samples in the $\Mmin$, $M_0$ plane
+for this model.  There is a small correlation between smaller $\Mmin$
+and larger $M_0$, which is driven by the need to widen the
+distribution to encompass the peak of the mass measurements in Figure
+\ref{fig:low-masses} when the minimum mass is smaller.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/exp-cutoff}
+  \end{center}
+  \caption{\label{fig:exp-marginal} The distribution of scale masses,
+    $M_0$, and minimum masses, $\Mmin$, both measured in units of a
+    solar mass for the exponential underlying mass distribution
+    defined in Equation \eqref{eq:exp-def}.  The median scale mass is
+    $M_0 = 1.55$, and $0.78 \leq M_0 \leq 3.25$ with 90\%
+    confidence.}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/exp-cutoff-2d}
+  \end{center}
+  \caption{\label{fig:exp-2D} The MCMC samples in the $\Mmin$, $M_0$
+    plane for the decaying exponential underlying mass distribution
+    model.  The slight correlation between smaller $\Mmin$ and larger
+    $M_0$ is driven by the need to widen the mass distribution to
+    encompass the peak of the measurements in Figure
+    \ref{fig:low-masses} when the minimum mass decreases.}
+\end{figure}
+
+\subsubsection{Gaussian and Two-Gaussian Models}
+\label{sec:gaussian}
+
+The mass distribution in Figure \ref{fig:low-masses} appears to have a
+single peak.  The prototypical single-peaked probability distribution
+is a Gaussian:
+\begin{equation}
+  \label{eq:gaussian-def}
+  p(M|\vtheta) = p(M|\{\mu, \sigma\}) = \frac{1}{\sigma \sqrt{2\pi}}
+  \exp\left[ - \left(\frac{M - \mu}{\sqrt{2} \sigma} \right)^2 \right].
+\end{equation}
+We use a prior on the mean mass, $\mu$, and the standard deviation,
+$\sigma$, that ensures that the majority of the mass distribution lies
+below $40 \Msun$:
+\begin{equation}
+  \label{eq:gaussian-prior-def}
+  p(\{\mu,\sigma\}) = 
+  \begin{cases}
+    \frac{8}{40^2} & 0 \leq \mu \leq 40, \quad \sigma \geq 0, \quad
+    \mu + 2\sigma \leq 40 \\
     0 & \textnormal{otherwise}
@@
 \end{equation}
-with $\Mmin$ and $M_0$ measured in solar masses.
-
-%Figure \ref{fig:exp-M0}
- displays the marginalized posterior
-distribution for the scale mass of the exponential, $M_0$.  The median
-scale mass is $M_0 = 3.00$, and $1.57 \leq M_0 \leq 5.71$ with 90\%
-confidence.  We defer discussion of the distribution of $\Mmin$ until
-Section \ref{sec:minimum-mass}.  Refer to Figure \ref{fig:dists}
-for the resulting shape of the distribution of distributions.
-
-% \begin{figure}
-%   \begin{center}
-%     \plotone{plots/exp-M0}
-%   \end{center}
-%   \caption{\label{fig:exp-M0} The distribution of scale masses, $M_0$,
-%     measured in units of a solar mass for the exponential underlying
-%     mass distribution defined in Equation \eqref{eq:exp-def}.  The
-%     median scale mass is $M_0 = 3.00$; with 90\% confidence $1.57 \leq
-%     M_0 \leq 5.71$. }
-% \end{figure}
-
-\subsubsection{Gaussian and Two-Gaussian Models}
-\label{subsubsec:gaussian}
-
-The prototypical single-peaked probability distribution is a
-Gaussian:
-\begin{equation}
-  p(m|\theta) = p(m|\{\mu, \sigma\}) = \frac{1}{\sigma \sqrt{2\pi}}
-  \exp\left[ - \left(\frac{m - \mu}{\sqrt{2} \sigma} \right)^2 \right].
-\end{equation}
-We choose a broad, flat prior 
-\begin{equation}
-  \label{eq:gaussian-def}
-  p(\{\mu,\sigma\}) = 
-  \begin{cases}
-    \frac{1}{40^2} & 0 \leq \mu, \sigma \leq 40 \\
-    0 & \textnormal{otherwise}
-  \end{cases},
-\end{equation}
 where both $\mu$ and $\sigma$ are measured in solar masses.  Figure
 \ref{fig:gaussian} shows the resulting marginalized distributions for
-the parameters $\mu$ and $\sigma$.  We constrain $\mu$ in $7.02 \leq
-\mu \leq 9.23$ with 90\% confidence.  (Refer to Figure \ref{fig:dists}
-for the shape of the resulting distribution of distributions.)
+the parameters $\mu$ and $\sigma$.  We constrain the peak of the
+Gaussian between $6.63 \leq \mu \leq 8.08$ with 90\% confidence.
+(Refer to Figure \ref{fig:dists} for the shape of the resulting
+distribution of distributions.)
 
@@
     Equation \eqref{eq:gaussian-def}.  The peak of the Gaussian,
-    $\mu$, is constrained in $7.02 \leq \mu \leq 9.23$ with 90\%
+    $\mu$, is constrained in $6.63 \leq \mu \leq 8.08$ with 90\%
     confidence.}
@@
 
-To look for a second peak in the black-hole mass distribution, we also
-examined a two-Gaussian model:
+Though we do not expect to find a second peak in the low-mass
+distribution, we may find evidence of one when exploring the combined
+low- and higher-mass samples.  To look for a second peak in the
+black-hole mass distribution, we use a two-Gaussian model:
 \begin{multline}
-  p(m|\theta) = p(m|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
-  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
-        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
-        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right],
+  \label{eq:two-gaussian-def}
+  p(M|\vtheta) = p(M|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
+  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
+        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
+        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right].
 \end{multline}
-where we restrict $\mu_1 < \mu_2$ with our broad, flat prior:
+The probability is a linear combination of two Gaussians with weight
+$\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose combined
+conditions on $\mu_i$ and $\sigma_i$ that ensure that most of the mass
+probability lies below $40 \Msun$ with the prior 
 \begin{equation}
@@
   \begin{cases}
-    2 \frac{1}{40^4} & 0 \leq \mu_1 \leq \mu_2 \leq 40,
-    \quad 0 \leq \sigma_1, \sigma_2 \leq 40, \quad 0 \leq \alpha \leq
-    1 \\
+    2 p(\{\mu_1, \sigma_1\}) p(\{\mu_2, \sigma_2\}) & \mu_1 \leq
+    \mu_2, \quad 0 \leq \alpha \leq 1 \\
     0 & \textnormal{otherwise}
-  \end{cases}.
-\end{equation}
-The one-Gaussian model is a subset of the Gaussian model: when
-$\alpha$ is small, then the presence of the first Gaussian is
-irrelevant, and the second gives the same mass distribution as a
-single Gaussian; when $\alpha$ is large, the reverse situation again
-reproduces the single-Gaussian model; and when $\alpha$ is in the
-middle of its range, the two means and standard deviations can be
-nearly equal, reproducing again the single-Gaussian model.  This is
-precisely what we observe from our MCMC samples in the two-Gaussian
-model.  In Section \ref{sec:model-selection}, we will quantify more
-precisely how much redundancy we have in this model relative to the
-one-Gaussian model, but we should emphasize here that there is no
-evidence of a significant contribution from a second peak in the
-two-Gaussian model.
+  \end{cases},
+\end{equation}
+where the single-Gaussian prior, $p(\{\mu_i, \sigma_i\})$, is defined
+in Equation \eqref{eq:gaussian-prior-def}.
+
+Figure \ref{fig:two-gaussian} shows the marginalized distributions for
+the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
+> 0.8$ with 62\% probability, clearly favoring the first Gaussian.
+The distribution for $\mu_1$ and $\sigma_1$ are similar to those of
+the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
+that the first Gaussian has found the peak mass samples in Figure
+\ref{fig:masses}.  The second Gaussian's parameter distributions are
+much broader.  The second Gaussian appears to be sampling the tail of
+the mass samples (see Figure \ref{fig:masses}).  In spite of the
+ability to match a more complicated distribution, we find that this
+model is strongly disfavored relative to the single-Gaussian model for
+this dataset: $p(\textnormal{Gaussian}|d) / p(\textnormal{Two
+  Gaussian}|d) \simeq 4.7$ (see Section \ref{sec:model-selection} for
+discussion).
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/two-gaussian}
+  \end{center}
+  \caption{\label{fig:two-gaussian} The marginal distributions for the
+    five parameters of the two-Gaussian model.  We have $\alpha > 0.8$
+    with 62\% probability, favoring the first of the two Gaussians.
+    The distributions for $\mu_1$ and $\sigma_1$ are similar to those
+    of the single Gaussian model displayed in Figure
+    \ref{fig:gaussian}; the second Gaussian's parameter distributions
+    are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
+    second Gaussian is attempting to fit the tail of the mass samples
+    (see Figure \ref{fig:masses}).  The extra degrees of freedom in
+    the distribution from the second Gaussian do not provide enough
+    extra fitting power to compensate for the increase in parmeter
+    space, however: the two-Gaussian model is disfavored relative to
+    the single Gaussian by a factor of $4.7$ on this dataset (see
+    Section \ref{sec:model-selection} for discussion).}
+\end{figure}
 
@@
 
-\subsection{Non-Parameteric Models for the Black Hole Mass Distribution}
-\label{subsec:non-parameteric-models}
+Many of the mass distributions for the systems in Figure
+\ref{fig:low-masses} rise rapidly to a peak and then fall off more
+slowly in a longer tail toward high masses.  So far, none of the
+parameterized distributions we have discussed have this property.  In
+this section, we consider a log-normal model for the underlying mass
+distribution; the log-normal distribution has a rise to a peak with a
+slower falloff in a long tail.
+
+The log-normal distribution gives $\log M$ a Gaussian distribution
+with mean $\mu$ and standard deviation $\sigma$:
+\begin{equation}
+  \label{eq:log-normal-def}
+  p(M|\vtheta) = p(M|\{\mu, \sigma \}) = \frac{1}{
+    \sqrt{2\pi} M \sigma} \exp\left[ -\frac{\left(\log M - \mu\right)^2}{2 \sigma^2} \right].
+\end{equation}
+The parameters $\mu$ and $\sigma$ are dimensionless; the mean mass
+$\langle M \rangle$ and mass standard deviation $\sigma_M$ are related
+to $\mu$ and $\sigma$ by
+\begin{eqnarray}
+  \label{eq:avg-M}
+  \langle M \rangle & = & \exp\left( \mu + \frac{1}{2} \sigma^2
+  \right) \\
+  \label{eq:sigma-M}
+  \sigma_M & = & \langle M \rangle \sqrt{\exp\left( \sigma^2 \right) - 1}.
+\end{eqnarray}
+For a fair comparison with the other models, we impose a prior that is
+flat in $\langle M \rangle$ and $\sigma_M$.  To ensure that most of
+the probability in this model occurs for masses below $40 \Msun$, we
+require $\langle M \rangle + 2 \sigma_M \leq 40$, resulting in a
+prior
+\begin{equation}
+  p(\vtheta) = p(\{\mu,\sigma\}) = 
+  \begin{cases}
+    \frac{4}{40^2} \left| \frac{\partial \left(\langle M \rangle,
+          \sigma_M \right)}{\partial \left( \mu, \sigma \right)}
+    \right| & \sigma > 0, \quad \langle M \rangle + 2 \sigma_M \leq 40
+    \\
+    0 & \textnormal{otherwise}
+  \end{cases},
+\end{equation}
+where 
+\begin{equation}
+  \left| \frac{\partial \left(\langle M \rangle,
+          \sigma_M \right)}{\partial \left( \mu, \sigma \right)}
+    \right| = \frac{\exp\left( 2 \left( \mu + \sigma^2 \right)\right)
+      \sigma}{\sqrt{\exp\left( \sigma^2 \right) - 1}}
+\end{equation}
+is the determinant of the Jacobian of the map in Equations
+\eqref{eq:avg-M} and \eqref{eq:sigma-M}.
+
+The marginal distributions for $\langle M \rangle$ and $\sigma_M$
+appear in Figure \ref{fig:log-normal}.  The distributions are similar
+to those for $\mu$ and $\sigma$ from the Gaussian model in Section
+\ref{sec:gaussian}.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/log-normal}
+  \end{center}
+  \caption{\label{fig:log-normal} Marginalized distributions of the
+    mean mass, $\langle M \rangle$, and standard deviation of the
+    mass, $\sigma_M$, for the log-normal model in Section
+    \ref{sec:log-normal}.  The distributions are similar to the
+    distributions of $\mu$ and $\sigma$ in the Gaussian model of
+    Section \ref{sec:gaussian}.}
+\end{figure}
+
+\subsection{Non-Parametric Models for the Black Hole Mass Distribution}
+\label{sec:non-parametric-models}
 
@@
 assume a priori a shape for the black hole mass distribution.  The
-fundamental non-parameteric distribution in this section is a
+fundamental non-parametric distribution in this section is a
 histogram with some number of bins, $\Nbin$.  Such a distribution is
-piecewise-constant in $m$.
+piecewise-constant in $M$.
 
@@
 We choose to represent a histogram mass distribution with $\Nbin$ bins
-by allocating a fixed total probability, $1/\Nbin$, to each bin.  The
-lower and upper bounds for each bin are allowed to vary; when these
-are close to each other (i.e.\ the bin is narrow), the distribution
-will have a large value, and conversely when the bounds are far from
-each other.  We assume that the non-zero region of the distribution is
+by allocating a fixed probability, $1/\Nbin$, to each bin.  The lower
+and upper bounds for each bin are allowed to vary; when these are
+close to each other (i.e.\ the bin is narrow), the distribution will
+have a large value, and conversely when the bounds are far from each
+other.  We assume that the non-zero region of the distribution is
 contiguous, so we can represent the boundaries of the bins as a
 non-decreasing array of masses, $w_0 \leq w_1 \leq \ldots \leq
-w_{\Nbin}$, with $w_0$ the minimum and $w_{\Nbin}$ the maximum mass for
-which the distribution has support.  This gives the distribution
+w_{\Nbin}$, with $w_0$ the minimum and $w_{\Nbin}$ the maximum mass
+for which the distribution has support.  This gives the distribution
 \begin{equation}
   \label{eq:hist-def}
-  p(m|\theta) = p(m|\{w_0, \ldots, w_{\Nbin}\}) = 
+  p(M|\theta) = p(M|\{w_0, \ldots, w_{\Nbin}\}) = 
   \begin{cases}
-    0 & m < w_0 \textnormal{ or } w_{\Nbin} \leq m \\
-    \frac{1}{\Nbin} \frac{1}{w_{i+1} - w_i} & w_i \leq m < w_{i+1}
+    0 & M < w_0 \textnormal{ or } w_{\Nbin} \leq M \\
+    \frac{1}{\Nbin} \frac{1}{w_{i+1} - w_i} & w_i \leq M < w_{i+1}
   \end{cases}.
@@
 parameters for one-, two-, three-, four-, and five-bin histogram
-models are shown in Figure \ref{fig:dists}.  As the number of bins
-increases, the models are better able to capture features of the mass
-distribution.  (But note that in Section \ref{sec:model-selection}, we
-will find that the one-bin histogram---which is a flat distribution
-between a maximum and minimum mass---is the best model from this
-group.)  We defer discussion of the distribution of the minimum mass
-from the histogram distributions until Section \ref{sec:minimum-mass}.
+models are shown in Figure \ref{fig:dists}.  Table
+\ref{tab:low-mass-non-parametric} gives quantiles of the marginalized
+bin boundary distributions for the histogram models.
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|c|c|c|}
+      \hline
+      Bins & Boundary & 5\% & 15\% & 50\% & 85\% & 95\% \\
+      \hline \hline
+      1 & $w_0$ & 3.94488 & 4.55603 & 5.43333 & 6.02557 & 6.29749 \\
+      \hline
+        & $w_1$ & 8.50844 & 8.69262 & 9.11784 & 9.83477 & 10.5128 \\
+      \hline \hline
+      2 & $w_0$ & 3.3426 & 4.2047 & 5.39132 & 6.18413 & 6.47553 \\
+      \hline
+        & $w_1$ & 6.41972 & 6.72605 & 7.43421 & 8.2489 & 8.52885 \\
+      \hline
+        & $w_2$ & 8.46161 & 8.65077 & 9.12694 & 10.1113 & 11.2595 \\
+      \hline \hline
+      3 & $w_0$ & 2.18176 & 3.54345 & 5.16094 & 6.16473 & 6.44697 \\
+      \hline
+        & $w_1$ & 5.68876 & 6.14223 & 6.68829 & 7.38725 & 8.04235 \\
+      \hline
+        & $w_2$ & 6.8297 & 7.22718 & 8.1451 & 8.7512 & 9.27296 \\
+      \hline
+        & $w_3$ & 8.44307 & 8.67362 & 9.25718 & 12.1688 & 21.92 \\
+      \hline \hline
+      4 & $w_0$ & 1.32131 & 2.7934 & 4.66156 & 5.78459 & 6.17946 \\
+      \hline
+        & $w_1$ & 5.20112 & 5.77331 & 6.42501 & 6.98427 & 7.44584 \\
+      \hline
+        & $w_2$ & 6.41805 & 6.73535 & 7.43826 & 8.32958 & 8.64212 \\
+      \hline
+        & $w_3$ & 7.40302 & 7.95608 & 8.58976 & 9.33897 & 10.3992 \\
+      \hline
+        & $w_4$ & 8.56724 & 8.8059 & 10.2451 & 24.3573 & 34.2423 \\
+      \hline \hline
+      5 & $w_0$ & 0.9392 & 2.28789 & 4.33389 & 5.7012 & 6.21166 \\
+      \hline
+        & $w_1$ & 4.69778 & 5.44302 & 6.26575 & 6.76407 & 7.14427 \\
+      \hline
+        & $w_2$ & 6.1388 & 6.47155 & 7.00606 & 7.97325 & 8.38259 \\
+      \hline
+        & $w_3$ & 6.82058 & 7.28677 & 8.22514 & 8.81555 & 9.41012 \\
+      \hline
+        & $w_4$ & 8.02335 & 8.36993 & 8.94879 & 11.3206 & 17.3349 \\
+      \hline
+        & $w_5$ & 8.7112 & 9.25208 & 16.2059 & 31.897 & 37.2738 \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:low-mass-non-parametric} The 5\%, 15\%, 50\%
+    (median), 85\%, and 95\% quantiles for the bin boundaries in the
+    one- through five-bin histogram models discussed in Section
+    \ref{sec:non-parametric-models}.}
+\end{table}
+
+As the number of bins increases, the models are better able to capture
+features of the mass distribution, but we find that the one-bin
+histogram is the most probable of the histogram models for the
+low-mass data (see Section \ref{sec:model-selection} for discussion).
+This occurs because the extra fitting power does not sufficiently
+improve the fit to compensate for the vastly larger parameter space of
+the models with more bins.
 
@@
 set of models, $\{M_i| i = 1, \ldots\}$, each with corresponding
-parameters $\theta_i$.  Re-writing Equation \eqref{eq:Bayes-rule} to
+parameters $\vtheta_i$.  Re-writing Equation \eqref{eq:Bayes-rule} to
 be explicit about the assumption of a particular model, we have
 \begin{equation}
+  p(\vtheta_i | d, M_i) = \frac{p(d|\vtheta_i, M_i) p(\vtheta_i | M_i)}{p(d|M_i)}.
+\end{equation}
+This gives the posterior probability of the parameters $\vtheta_i$ in
+the context of model $M_i$.  But, the model itself can be regarded as
+a discrete parameter in a larger ``super-model'' that encompasses all
+the $M_i$.  The parameters for the super-model are $\{M_i,
+\vtheta_i\}$: a choice of model and the corresponding parameter value
+within that model.  The posterior probability of the super-model
+parameters is given by Bayes' rule:
+\begin{equation}
   \label{eq:bayes-explicit-model}
-  p(\theta_i | d, M_i) = \frac{p(d|\theta_i, M_i) p(\theta_i | M_i)}{p(d|M_i)}.
-\end{equation}
-This gives the posterior probability of the parameters $\theta_i$ in
-the context of model $M_i$.  But, the particular model itself can be
-regarded as a parameter in a larger ``super-model'' that encompasses
-all the $M_i$.  If we let the model become a parameter, Bayes' rule
-gives us
-\begin{equation}
-  p(\theta_i, M_i|d) = \frac{p(d|\theta_i, M_i) p(\theta_i |M_i) p(M_i)}{p(d)},
+  p(\vtheta_i, M_i|d) = \frac{p(d|\vtheta_i, M_i) p(\vtheta_i |M_i) p(M_i)}{p(d)},
 \end{equation}
@@
 \begin{equation}
-  p(d) = \sum_i \int d\theta_i\, p(\theta_i, M_i|d).
-\end{equation}
+  \label{eq:multi-model-evidence-def}
+  p(d) = \sum_i \int d\vtheta_i\, p(\vtheta_i, M_i|d) = \sum_i
+  p(d|M_i) p(M_i),
+\end{equation}
+writing the single-model evidence from Equation
+\eqref{eq:evidence-def} as $p(d|M_i)$ to be explicit about the
+dependence on the choice of model.
 
 To compare the various models $M_i$, we are interested in the
-marginalized posterior probability of $M_i$ (also called the evidence
-for model $M_i$):
-\begin{equation}
-  p(M_i|d) \equiv \int d\theta_i\, p(\theta_i, M_i|d).
-\end{equation}
-This quantity can be re-written in terms of the single-model evidence
+marginalized posterior probability of $M_i$:
+\begin{equation}
+  \label{eq:model-posterior-def}
+  p(M_i|d) \equiv \int d\vtheta_i\, p(\vtheta_i, M_i|d).
+\end{equation}
+This is the total posterior probability found in the parameter space
+of model $M_i$.  The marginalized posterior probability of model $M_i$
+can be re-written in terms of the single-model evidence, $p(d|M_i)$
 (see Equations \eqref{eq:bayes-explicit-model} and
@@
   \label{eq:model-evidence-def}
-  p(M_i|d) = \frac{p(M_i)}{p(d)} \int d\theta_i
-  p(d|\theta_i,M_i) p(\theta_i|M_i) = \frac{p(d|M_i) p(M_i)}{p(d)}.
+  p(M_i|d) = \int d\vtheta_i\, p(\vtheta_i, M_i|d) = \frac{p(M_i)}{p(d)} \int d\vtheta_i
+  p(d|\vtheta_i,M_i) p(\vtheta_i|M_i) = \frac{p(d|M_i) p(M_i)}{p(d)}.
+\end{equation}
+
+Here and throughout, we assume that any of the models in Section
+\ref{sec:models} are equally likely a priori, so the model priors are
+equal:
+\begin{equation}
+  p(M_i) = \textnormal{const}.
 \end{equation}
@@
 analysis of Section \ref{sec:models}.  The first way, which we discuss
-in Section \ref{subsec:direct-evidence}, is to use the MCMC samples
-from the various models to estimate the value of the integral in
-Equation \eqref{eq:model-evidence-def}.  The second way is to treat
-the model literally as a parameter in a second MCMC analysis.  One
-then proposes jumps (see the discussion in Section \ref{sec:models})
-not only for the $\theta_i$ within a particular model, but from
-parameters $\theta_i$ in model $M_i$ to parameters $\theta_j$ in
-$M_j$.  This technique is called a ``reversible-jump MCMC''
-\citep{Green1995}; we discuss it in Section
-\ref{subsec:reversible-jump-mcmc}.
+in Section \ref{sec:direct-evidence}, is to use the MCMC samples
+from the various models to estimate the single model evidence,
+$p(d|M_i)$, appearing in Equation \eqref{eq:model-evidence-def}.  The
+second way is to treat the model literally as a parameter in a second
+MCMC analysis.  In the second MCMC analysis, one proposes jumps not
+between different $\vtheta_i$ within a particular model, but from
+parameters $\vtheta_i$ in model $M_i$ to parameters $\vtheta_j$ in
+$M_j$.  The evidence for model $M_i$ is proportional to the number of
+times this MCMC has a sample in the $M_i$ parameter space.  This
+technique is called a ``reversible-jump MCMC'' \citep{Green1995}; we
+discuss it in Section \ref{sec:reversible-jump-mcmc}.
 
 \subsection{Direct Evidence Calculations}
-\label{subsec:direct-evidence}
-
-In this section we discuss techniques for using the MCMC samples from
-Section \ref{sec:models} to estimate the integral in Equation
-\eqref{eq:model-evidence-def}.  The first technique we use comes from
+\label{sec:direct-evidence}
+
+To compute $p(M_i|d)$ from Equation \eqref{eq:model-evidence-def}
+using the MCMC samples we discussed in Section \ref{sec:models} we
+need a way to compute the single-model evidence, $p(d|M_i)$.  The
+remaining terms in Equation \eqref{eq:model-evidence-def} are either
+under our control (the prior $p(M_i)$) or a normalization factor that
+we can compute using Equation \eqref{eq:multi-model-evidence-def} once
+we have the single model evidence (the evidence $p(d)$).  Begin by
 re-writing Equation \eqref{eq:bayes-explicit-model}:
@@
 \begin{equation}
-  \frac{1}{p(d|M_i)} \approx \frac{1}{N_i} \sum_j \frac{1}{p(d|\theta_{ij}, M_i)};
-\end{equation}
-the evidence for model $M_i$ can be estimated from the harmonic mean
-of the likelihood over the MCMC samples.
-
-Alternately, we can attempt to perform the integral over $\theta_i$
-for the model evidence directly using our MCMC samples.  If we could
-assign non-overlapping volumes, $V_j$, to each of the $N_i$ MCMC
-samples, $\theta_{ij}$, then we could approximate the integral in
-Equation \eqref{eq:model-evidence-def} via
-\begin{equation}
-  \label{eq:direct-evidence-estimate}
-  \int d\theta_i
-  p(d|\theta_i,M_i) p(\theta_i|M_i) \approx \sum_j
-  p(d|\theta_{ij},M_i) p(\theta_{ij}|M_i) V_j.
-\end{equation}
-
-To associate a volume to each MCMC sample, we use a data structure
-known as a kD-tree.  (Our algorithm is similar to Algorithm P of
-\citet{Weinberg2010}.)  A kD-tree decomposes a volume in
-$\mathbb{R}^n$ into non-overlapping boxes, each of which contains some
-number of points from a given sample.  The decomposition proceeds
-recursively: given a region to subdivide and a sample of points, a
-dimension is chosen (we choose the dimension of largest extent of the
-sample, but other choices are possible) and the region is split along
-this dimension such that the numbers of points to each side of the
-partition are equal or differ by one.  Each of the resulting
-sub-regions and the sub-samples contained in them are further split in
-the same manner, until every region contains fewer than some chosen
-number of points.  The sequence of splittings forms a binary tree,
-with the initial region and point sample at the root, and the final
-boxes and subsamples at the leaves.  From experimentation on
-artificial point distributions, we find that terminating the splitting
-when each box contains fewer than $\sim 64$ to $\sim 256$ samples is
-reasonably effective in estimating evidence integrals in $\sim 5$
-dimensions.  (See also \citet{Weinberg2010}.)  To estimate the
-integral as in Equation \eqref{eq:direct-evidence-estimate}, we walk
-the leaves of the kD-tree; for each leaf, we compute the volume
-spanned by the samples in that leaf, and assign an equal fraction of
-this volume to each sample (note that the volume spanned by the
-samples is \emph{smaller} than the leaf region itself).  The integral
-is computed using the assigned volumes as in Equation
-\eqref{eq:direct-evidence-estimate}.
+  \frac{1}{p(d|M_i)} \approx \frac{1}{N_i} \sum_j \frac{1}{p(d|\theta_{ij}, M_i)}.
+\end{equation}
+This technique is known as the ``harmonic mean approximation'' for the
+single-model evidence $p(d|M_i)$.
+
+\citet{Weinberg2010} discusses techniques to estimate the integral
+over parameter space in Equation \eqref{eq:model-evidence-def} by
+adaptively partitioning the parameter volume into non-overlapping
+boxes surrounding the points in the MCMC sample.  We implemented the
+algorithms from that paper, but found them, in this case, to give
+results that depended strongly on the parameters in the algorithm and
+were sometimes inconsistent with both the harmonic mean calculations
+in this Section and the reversible jump calculations in Section
+\ref{sec:reversible-jump-mcmc}.  Since the harmonic mean and
+reversible jump calculations are in agreement with each other, we will
+not discuss these direct integration techniques further.
 
@@
 statistical uncertainty due to the tendency for a few low-likelihood
-points to dominate the sum \citep{Weinberg2010}; the statistics of the
-kD-tree estimate are less well-known.  In any case, it is useful to
-have multiple independent estimates of the evidence for a given model.
-In Figure \ref{fig:direct-evidence} we use both methods to estimate
-the evidence for the various models discussed in Section
-\ref{sec:models}.  We assume that no model is a priori more likely
-than any other, so the model priors $p(M_i)$ are equal.  We also
-ignore the model-independent constant $p(d)$ in Equation
-\eqref{eq:model-evidence-def}, so only ratios of evidence are
-relevant.  Due to the large statistical uncertainty expected in the
-harmonic mean estimate of the evidence, we perform bootstrap estimates
-of the range of the harmonic mean, reporting the 10\% and 90\% values
-in the error bars of Figure \ref{fig:direct-evidence}.
+points to dominate the sum \citep{Weinberg2010}.  Therefore, it is
+useful to have multiple independent estimates of the evidence for a
+given model.  In Section \ref{sec:reversible-jump-mcmc}, we will
+discuss the reversible jump algorithm for estimating the evidence for
+our models; the results presented here are in broad agreement with the
+results in that Section.  In Figure \ref{fig:harmonic-evidence} we
+display the harmonic mean estimator for the evidence for the various
+models discussed in Section \ref{sec:models}.  In addition, we have
+performed 10000 calculations of the harmonic mean estimator using
+``mock'' MCMC data drawn randomly from our true MCMC with replacement.
+The errorbars in Figure \ref{fig:harmonic-evidence} indicate the 10\%
+and 90\% quantiles from this bootstrap calculation.  Recall that we
+assume that no model is a priori more likely than any other, so the
+model priors $p(M_i)$ are equal.
 
@@
   \end{center}
-  \caption{\label{fig:direct-evidence} Direct integration and harmonic
-    mean evidence for the various models from Section \ref{sec:models}
-    (see Section \ref{subsec:direct-evidence}).  In increasing
-    order along the $x$-axis, the models are the power-law of
-    Section \ref{subsubsec:power-law} (PL), the decaying
-    exponential of Section \ref{subsubsec:exponential} (E), the
-    single Gaussian of Section \ref{subsubsec:gaussian} (G), the
-    double Gaussian of Section \ref{subsubsec:gaussian} (TG),
-    and the one-, two-, three-, four-, and five-bin histogram models
-    of Section \ref{subsec:non-parameteric-models} (H1, H2, H3, H4,
-    H5, respectively).  The cross symbols give the value of the direct
-    integration; the plus symbols give the harmonic mean estimator
-    with bootstrap estimates of the 10\% and 90\% bounds on the
-    harmonic mean.  It appears that the power law and one-bin
-    histogram are the two best models, but the error ranges and
-    inconsistency between the two techniques for calculating evidence
-    cast some doubt on this conclusion.}
-\end{figure}
-
-The results of the direct evidence calculation appear to point to the
-power law and one-bin histogram as the two most favored models, but
-the error ranges and inconsistency between the two techniques for
-calculating evidence cast some doubt on this conclusion.  In the next
-section, we will confirm these results with a more accurate method for
-estimating the evidence.
+  \caption{\label{fig:harmonic-evidence} Harmonic mean evidence for
+    the various models from Section \ref{sec:models} (see Section
+    \ref{sec:direct-evidence}).  In increasing order along the
+    $x$-axis, the models are the power-law of Section
+    \ref{sec:power-law} (PL), the decaying exponential of Section
+    \ref{sec:exponential} (E), the single Gaussian of Section
+    \ref{sec:gaussian} (G), the double Gaussian of Section
+    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
+    five-bin histogram models of Section
+    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
+    respectively).  The errorbars are bootstrap estimates of the 10\%
+    and 90\% bounds on the harmonic mean estimator.  The power-law and
+    Gaussian are the favored models, with the single-bin histogram
+    coming third (but the 10-90\% bootstrap range for the one-bin
+    histogram includes the possibility of being the most favored
+    model).  This is not surprising, given the single-peaked nature of
+    the lower-mass sample.}
+\end{figure}
+
+The results of the harmonic mean calculation indicate that the
+power-law and Gaussian models are most favored, with the single-bin
+histogram coming third (though the 10-90\% range for the models
+includes the possibility of it being the most favored).  This is not
+surprising, given the single-peaked nature of the lower-mass sample.
+The log normal and exponential models are likely disfavored due to
+their long tails.  The multiple-bin histogram models are disfavored
+because the structure of the lower-mass mass distribution is not
+complicated enough to warrant the extra parameter space implied by the
+larger number of bins.  The two Gaussian model is disfavored because
+the distribution is single-peaked, so the second Gaussian is
+superfluous.
 
 \subsection{Reversible-Jump MCMC}
-\label{subsec:reversible-jump-mcmc}
-
-The technique of reversible-jump MCMC \citep{Green1995} follows the
-``super-model'' concept, in which the choice of model, $M_i$, becomes
-a parameter, to its logical conclusion.  One performs an MCMC that
-proposes jumps between the parameter spaces of the various models that
-are under consideration (i.e.\ from parameters $\theta_i$ in model
-$M_i$ to parameters $\theta_j$ in model $M_j$).  For this MCMC to be
-efficient, proposed jumps into a model should favor regions with large
-posterior; when the posterior is highly-peaked in a small region of
-parameter space, proposed jumps outside this region are unlikely to be
-accepted, and the reversible-jump MCMC samples will require a very
-long chain to properly sample the ``super-model'' posterior.  The
-evidence for a particular model is proportional to the number of
-reversible-jump MCMC samples that live in that model's parameter
-space.
-
-We use the kD-trees from the last subsection to interpolate the
-posterior between MCMC samples from single-model MCMCs when proposing
-jumps in our reversible-jump MCMC.  Thus, proposed jumps track the
-posteriors computed from the single-model MCMCs.  (For an alternative
-approach to generating jump proposals for reversible-jump MCMCs, see
-\citet{Littenberg2009}.)  The interpolation algorithm works as
-follows.  We construct a kD-tree as described in the last subsection,
-continuing the subdivision until each region contains exactly one
-sample point.  To propose a jump, we first choose a model uniformly
-from the set of models (i.e.\ according to the model priors).  Then we
-choose a point uniformly from the sample chain (i.e.\ a point is
-chosen according to the posterior).  Then we find the leaf region in
-the kD-tree that contains this point, and propose a jump uniformly
-within this region.  This procedure for proposing a jump admits a
-straightforward calculation of the jump proposal distribution.  The
-advantage of using a kD-tree instead of the method in
-\citet{Littenberg2009} is that the tree is locally adaptive to the
-structure of the MCMC samples, and therefore more efficiently tracks
-the posterior from each of the single-model MCMCs.
-
-We have performed a reversible-jump MCMC that jumps between all the
-models (both parameteric and non-parameteric) in Section
-\ref{sec:models}.  The counts for the number of samples in each model
-appear in Figure \ref{fig:rj}; relative probabilities of the models
-are given in Table \ref{tab:rj}.  The most favored model is the power
-law from Section \ref{subsubsec:power-law}, followed by the one-
-and two-bin histograms from Section
-\ref{subsec:non-parameteric-models}.  This confirms the results from
-the direct integration and harmonic mean evidence estimators in the
-last subsection.  Interestingly, the theoretical curve from
-\citet{Fryer2001} places fourth in the ranking of evidence.  However,
-while the exponential distribution is a factor of 3.9 times less
-likely than the best-fit power law model, it is only 1.2 times less
-likely than the two best histogram models.  A choice of model prior
-that only slightly favored the exponential model over the
-non-parameteric histograms would put it in second place.
+\label{sec:reversible-jump-mcmc}
+
+Another technique for computing $p(M_i|d)$ is the reversible-jump MCMC
+\citep{Green1995}.  For a reversible-jump MCMC, we consider a
+``super-model'' that encompasses all the models under consideration.
+The parameter space of the super-model consists of a discrete
+parameter that identifies the choice of model, $M_i$, and the
+continuous parameters appropriate for this model, $\vtheta_i$.  We
+denote a point in the super-model parameter space by $\{M_i,
+\vtheta_i\}$; each such point is a statement that, e.g., ``the
+underlying mass distribution is a Gaussian, with parameters $\mu$ and
+$\sigma$,'' or ``the underlying mass distribution is a triple-bin
+histogram with parameters $w_1$, $w_2$, $w_3$, and $w_4$,'' or ....
+To compare models, we are interested in the quantity (see Equation
+\eqref{eq:model-posterior-def})
+\begin{equation}
+  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d).
+\end{equation}
+If we perform an MCMC in the super-model parameter space, then we
+obtain a chain of samples $\{M_i, \vtheta_i\}$ distributed in
+parameter space with density $p(M_i,\vtheta_i) d\vtheta_i$ and we can
+estimate the integral as
+\begin{equation}
+  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d) \approx \frac{N_i}{N},
+\end{equation}
+where $N_i$ is the number of MCMC samples that have discrete parameter
+$M_i$, and $N$ is the total number of samples in the MCMC.  In other
+words, the fraction of samples lying in the parameter space of model
+$M_i$ gives the probability of that model relative to the other models
+under consideration.
+
+To perform the MCMC in the super-model parameter space, we must
+propose jumps not only between points in a particular model's
+parameter space (as we have been doing up to now), but also between
+the parameter spaces of different models.  For this MCMC to be
+efficient, proposed jumps into a model from another should favor
+regions with large posterior; when the posterior is highly-peaked in a
+small region of parameter space, proposed jumps outside this region
+are unlikely to be accepted, and the reversible-jump MCMC samples will
+require a very long chain to properly sample the ``super-model''
+posterior.
+
+We can exploit the information we have from the single-model MCMCs
+described in Section \ref{sec:models} to generate efficient jump
+proposal distributions for our reversible jump MCMC.  We would like to
+propose jumps that roughly follow the distribution of samples in the
+single-model MCMCs.  We can do this by assigning a neighborhood to
+each point in the sample using an algorithm we will describe in the
+following paragraphs; the neighborhoods are non-overlapping,
+completely cover the region of parameter space with prior support, and
+contain only one point from the MCMC samples.  To propose a jump into
+model $M_i$, we choose a point uniformly from that single-model MCMC
+and then propose a jump drawn uniformly from that point's
+neighborhood.  This is equivalent to drawing from a piecewise-constant
+approximation to the single-model posterior, where each neighborhood
+contributes a constant fraction, $1/N_i$, to the cumulative jump
+probability.  In regions of high density the neighborhoods are
+smaller, and the jump probability density is correspondingly higher.
+Because the neighborhoods cover the entire region of prior support, it
+is possible for the proposal to propose any point in parameter space
+with prior support (though points in regions of low single-model
+posterior are of course unlikely to be proposed).
+
+To assign a neighborhood to each point in a single-model MCMC we use a
+data structure called a kD-tree.  A kD-tree is a binary
+space-partitioning tree.  To construct a kD-tree, we begin with the
+set of points from a single-model MCMC and a box in parameter space
+bounding the region of prior support (which must necessarily enclose
+all the points).  The construction proceeds recursively: we choose a
+dimension%
+\footnote{Our algorithm chooses the dimension along which the
+  numerical extent of the points is largest.  Other choices are
+  possible; some algorithms cycle through the dimensions in order,
+  while others choose a random dimension for each subdivision.  Our
+  goal by picking the longest dimension is to produce neighborhoods
+  that are ``square,'' at least in the chosen parametrization.} %
+along which to divide the points, find the median point along that
+dimension and its nearest neighbor, and divide the box at the midpoint
+between these two points, producing two sub-boxes.  We then partition
+the points into those to the left (i.e.\ smaller coordinate along the
+given dimension) and right of the dividing line, and repeat this
+procedure for each subset and the corresponding bounding box, until we
+have only one point in each box.  An example of the neighborhoods that
+result from a two-dimensional kD-tree constructed around a Gaussian
+point distribution appears in Figure \ref{fig:kD-tree}.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/kdtree}
+  \end{center}
+  \caption{\label{fig:kD-tree} The neighborhoods constructed from a
+    two-dimensional kD-tree built from a sample of points with a
+    Gaussian density distribution.  Each line on the figure
+    corresponds to a sub-dividing box boundary drawn between the
+    median of a subset of the sample points and its nearest
+    neighbor. The peak of the Gaussian lies in the center of the
+    figure; here the point density is highest and the neighborhoods
+    are smallest.  Near the edges the density is lower, and the
+    neighborhoods correspondingly larger.  The tree adapts itself to
+    the local density of points.  If these were single-model MCMC
+    samples, the corresponding jump proposal would first select one of
+    the boxes uniformly at random, and then choose a point uniformly
+    within the box to propose.  Since there are many more boxes near
+    the center (each box corresponds to one point), and these boxes
+    are smaller, the proposal will tend to concentrate its points
+    there, approximately tracking the distribution of single-model
+    MCMC samples.}
+\end{figure}
+
+Construction of a kD-tree is an $\order{N\log N}$ operation, where $N$
+is the number of points in the tree.  Median finding is $\order{n}$,
+where $n$ is the number of points from which the median is to be
+obtained.  At level $i$ in the tree, there are $2^i$ subsets of
+points, each of length $\order{N/2^i}$, so the total cost of the $2^i$
+median calculations is $\order{N}$ at each level.  There are
+$\order{\log N}$ levels in the tree, yielding a total construction
+cost for the tree of $\order{N \log N}$.
+
+To find the neighborhood of a point using the tree, we begin at the
+root of the tree, and examine the two sub-boxes at the next level
+down.  The point will be in one of them; following that branch, we
+have again two sub-boxes, one of which contains the point; following
+that branch....  Eventually, the search terminates at a leaf of the
+tree, containing the point in question.  The box at the leaf defines
+the neighborhood of the point in the jump proposal algorithm described
+above.  The total cost for this operation is proportional to the
+number of levels in the tree, which is $\order{\log N}$.
+
+We have performed a suite of 500 independent reversible-jump MCMCs
+jumping between all the models (both parametric and non-parametric)
+in Section \ref{sec:models} using the single-model MCMC samples to
+construct an efficient jump proposal for each model as described
+above.  The numbers of counts in each model are consistent across the
+MCMCs in the suite; Figure \ref{fig:rj} displays the average
+probability for each model across the suite, along with the 1-$\sigma$
+errors on the average inferred from the standard deviation of the
+model counts across the suite.  Table \ref{tab:rj}
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/rj}
+  \end{center}
+  \caption{\label{fig:rj} The relative probability of the models
+    discussed in Section \ref{sec:models} as computed using the
+    reversible-jump MCMC with the efficient jump proposal algorithm
+    described in Section \ref{sec:reversible-jump-mcmc}.  In
+    increasing order along the $x$-axis, the models are the power-law
+    of Section \ref{sec:power-law} (PL), the decaying exponential of
+    Section \ref{sec:exponential} (E), the single Gaussian of Section
+    \ref{sec:gaussian} (G), the double Gaussian of Section
+    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
+    five-bin histogram models of Section
+    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
+    respectively).  The average of 500 independent reversible-jump
+    MCMCs is plotted, along with the 1-$\sigma$ error on the average
+    inferred from the standard deviation of the probability from the
+    individual MCMCs.  As discussed in the text, the power-law and
+    Gaussian models are the most favored.}
+\end{figure}
+
+The most favored model is the power law from Section
+\ref{sec:power-law}, followed by the Gaussian model from Section
+\ref{sec:gaussian}.  This confirms the results from the harmonic mean
+evidence estimator in the last subsection.  (Though note that the
+harmonic mean ranks the one-bin histogram significantly higher than
+the reversible jump calculation; given the large uncertainty in the
+harmonic mean estimator, it is likely that the true evidence for the
+one-bin histogram model is closer to the value computed in this
+section.)  Interestingly, the theoretical curve from \citet{Fryer2001}
+(the exponential model of Section \ref{sec:exponential}) places fourth
+in the ranking of evidence.
 
@@
       \hline \hline
-      Power Law (Section \ref{subsubsec:power-law}) & 0.441660 \\
-      \hline
-      Histogram (1 Bin, Section \ref{subsec:non-parameteric-models}) &
-      0.143435 \\
-      \hline
-      Histogram (2 Bin, Section \ref{subsec:non-parameteric-models}) &
-      0.135854 \\
-      \hline 
-      Exponential (Section \ref{subsubsec:exponential}) & 0.112514 \\
-      \hline
-      Histogram (3 Bin, Section \ref{subsec:non-parameteric-models}) &
-      0.075592 \\
-      \hline
-      Histogram (4 Bin, Section \ref{subsec:non-parameteric-models}) &
-      0.038839 \\
-      \hline 
-      Gaussian (Section \ref{subsubsec:gaussian}) & 0.022999 \\ 
-      \hline
-      Histogram (5 Bin, Section \ref{subsec:non-parameteric-models}) &
-      0.021011 \\
-      \hline      
-      Two Gaussian (Section \ref{subsubsec:gaussian}) & 0.008096 \\
+      Power Law (Section \ref{sec:power-law}) & 0.331488 \\
+      \hline
+      Gaussian (Section \ref{sec:gaussian}) & 0.288129 \\ 
+      \hline
+      Log Normal (Section \ref{sec:log-normal}) & 0.138435 \\
+      \hline
+      Exponential (Section \ref{sec:exponential}) & 0.0916218 \\
+      \hline
+      Two Gaussian (Section \ref{sec:gaussian}) & 0.0662577 \\
+      \hline
+      Histogram (1 Bin, Section \ref{sec:non-parametric-models}) &
+      0.0641941 \\
+      \hline
+      Histogram (2 Bin, Section \ref{sec:non-parametric-models}) &
+      0.015184 \\
+      \hline 
+      Histogram (3 Bin, Section \ref{sec:non-parametric-models}) &
+      0.00332933 \\
+      \hline
+      Histogram (4 Bin, Section \ref{sec:non-parametric-models}) &
+      0.000999976 \\
+      \hline 
+        Histogram (5 Bin, Section \ref{sec:non-parametric-models}) &
+      0.0003614  \\
       \hline      
@@
     from Section \ref{sec:models} computed from reversible-jump
-    MCMC samples.}
+    MCMC samples with the efficient jump proposal algorithm in Section \ref{sec:reversible-jump-mcmc}.}
 \end{table}
 
-Assuming that the reversible-jump MCMC has converged (and we have
-tested ours for convergence), the standard deviation on the number of
-counts in each model is given by
-\begin{equation}
-  \sigma_{N_i} = \sqrt{N_i \left( 1 - \frac{N_i}{N} \right)},
-\end{equation}
-where $N_i$ is the number of counts in model $M_i$, and $N$ is the
-total number of reversible-jump counts.  This quantity is extremely
-small relative to the statistical error in the harmonic mean estimator
-or the likely truncation error on the direct integration estimate, so
-we consider the reversible-jump evidence results to be definitive.
-
-\begin{figure}
-  \begin{center}
-    \plotone{plots/rj}
-  \end{center}
-  \caption{\label{fig:rj} Counts of samples in the parameter space of
-    each model from Section \ref{sec:models} in the reversible-jump
-    MCMC of Section \ref{subsec:reversible-jump-mcmc}.  The models
-    are the power law of Section \ref{subsubsec:power-law} (PL),
-    the exponential of Section \ref{subsubsec:exponential} (E),
-    the Gaussian (G) and two Gaussian (TG) models of Section
-    \ref{subsubsec:gaussian}, and the one-, two-, three-, four-, and
-    five-bin histograms of Section
-    \ref{subsec:non-parameteric-models} (H1, H2, H3, H4, H5,
-    respectively).  The number of counts in each model is proportional
-    to the evidence for that model; therefore, the most favored model
-    is the power-law, followed by the one- and two-bin histogram.  }
-\end{figure}
+Though the model probabilities presented in this section have small
+statistical error, they are subject to large ``systematic'' error.
+The source of this error is both the particular choice of model prior
+(uniform across models) and the choice of priors on the parameters
+within each model used for this work.  For example, the
+theoretically-preferred exponential model (Section
+\ref{sec:exponential}) is only a factor of $\sim 3$ away from the
+most-favored Power Law model (Section \ref{sec:power-law}), which does
+not have theoretical support.  Is such support worth a factor of three
+in the model prior?  Alternately, we may say we know (in advance of
+any mass measurements) that black holes must exist with mass $\lesssim
+10\Msun$; then we could, for example, impose a prior on the minimum
+mass in the exponential model ($\Mmin$) that is uniform between $0$
+and $10 \Msun$, which would reduce the prior volume available for the
+model by a factor of 4 without significantly reducing the posterior
+support for the model.  This has the same effect as increasing the
+model prior by a factor of 4, which would move this model from fourth
+to first place.  Of course, we would then have to modify the prior
+support for the other models to take into account the restriction that
+there must be black holes with $M \lesssim 10\Msun$....
+\citet{Linder2008} discuss these issues in the context of cosmological
+model selection, concluding with a similar warning against
+over-reliance on model selection probabilities.
+
+Nevertheless, we believe that our model comparison is reasonably fair
+(see the discussion of priors in Section \ref{sec:priors}).  It seems
+safe to conclude that ``single-peaked'' models (the power-law and
+Gaussian) are preferred over ``extended'' models (the exponential or
+log-normal), or those with ``structure'' (the many-bin histograms or
+two-Gaussian model).  Previous studies have also supported the
+``single, narrow peak'' mass distribution \citep{Bailyn1998,Ozel2010}.
+In this light, the difference between the evidence calculated from the
+harmonic mean estimator and the reversible-jump MCMC for the
+single-bin histogram is surprising; one would expect that a
+narrowly-peaked single-bin histogram would perform well (as it does
+with the harmoinc-mean estimator) instead of being disfavored relative
+to the other models (as with the reversible-jump MCMC).
 
@@
 
+This section repeats the analysis of Sections \ref{sec:models} and
+\ref{sec:model-selection} but including the higher-mass, wind-fed
+systems from Table \ref{tab:sources} (see also Figure
+\ref{fig:high-masses}) in the sample.  Figure
+\ref{fig:high-mass-dists} displays bounds on the value of the
+underlying mass distribution for the various models in Section
+\ref{sec:models} applied to this data set; compare to Figure
+\ref{fig:dists}.  The inclusion of the higher-mass, wind-fed systems
+tends to widen the distribution toward the high-mass end and, in
+models that allow it, produce a second, higher-mass peak in addition
+to the one in Figure \ref{fig:dists}.
+
+\begin{figure}
+  \begin{center}
+    \plottwo{plots/dist-parametric-high}{plots/dist-non-parametric-high}
+  \end{center}
+  \caption{\label{fig:high-mass-dists} The median values of the black
+    hole mass distribution, $p(M|\theta)$, at various masses implied
+    by the posterior $p(\theta|d)$ for the models discussed in
+    Sections \ref{sec:parametric-models} and
+    \ref{sec:non-parametric-models}.  These distributions use the
+    complete sample of 20 observations in Table \ref{tab:sources},
+    including the higher-mass, wind-fed systems.  Error bars span the
+    10\% to 90\% range.  Note that these ``distributions of
+    distributions'' are not necessarily normalized, and need not be
+    ``shaped'' like the underlying model distributions.  Compare to
+    Figure \ref{fig:dists}, which includes only the low-mass systems
+    in the analysis.  Including the higher-mass systems tends to widen
+    the distribution toward the high-mass end and, in models that
+    allow it, produce a second, higher-mass peak in addition to the
+    one in Figure \ref{fig:dists}. }
+\end{figure}
+
+Figures \ref{fig:power-law-high} presents the marginalized
+distribution for the three power-law parameters $\Mmin$, $\Mmax$, and
+$\alpha$ (Section \ref{sec:power-law}) from an analysis including the
+higher-mass systems.  The distribution for $\Mmax$ is quite broad
+because the best fit power laws slope downward ($\alpha < 0$), making
+this parameter less relevant.  The range $-5.05 \leq \alpha \leq
+-1.77$ encloses 90\% of the probability; the median value of $\alpha$
+is -3.23.  The presence of the higher-mass samples in the analysis
+produces a distinctive tail, eliminating the correlations discussed in
+Section \ref{sec:power-law} and displayed in Figure
+\ref{fig:power-law-2D} for the lower-mass subset of the observations.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/power-law-high}
+  \end{center}
+  \caption{\label{fig:power-law-high} A histogram of the marginalized
+    distribution for the three parameters $\Mmin$, $\Mmax$, and
+    $\alpha$ from the power-law model including the higher-mass
+    samples in the MCMC.  The distribution for $\Mmax$ is quite broad
+    because the best fit power laws slope downward ($\alpha < 0$),
+    making this parameter less relevant.  The range $-5.05 \leq \alpha
+    \leq -1.77$ encloses 90\% of the probability; the median value of
+    $\alpha$ is -3.23.  The presence of the higher-mass samples in the
+    analysis produces a distinctive tail, eliminating the correlations
+    discussed in Section \ref{sec:power-law} and displayed in Figure
+    \ref{fig:power-law-2D} for the lower-mass subset of the
+    observations. }
+\end{figure}
+
+Figure \ref{fig:exp-cutoff-high} displays the marginalized
+distributions for the exponential parameters $\Mmin$ and $M_0$
+(Section \ref{sec:exponential}) from an analysis including the
+higher-mass systems.  The distribution for the scale mass, $M_0$, has
+moved to higher masses relative to Figure \ref{fig:exp-marginal} to
+fit the tail of the mass distribution; the distribution for $\Mmin$ is
+less affected, though it has broadened somewhat toward low masses.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/exp-cutoff-high}
+  \end{center}
+  \caption{\label{fig:exp-cutoff-high} The marginalized distributions
+    for the exponential parameters $\Mmin$ and $M_0$ (Section
+    \ref{sec:exponential}) from an analysis including the higher-mass
+    systems.  The distribution for the scale mass, $M_0$, has moved to
+    higher masses relative to Figure \ref{fig:exp-marginal} to fit the
+    tail of the mass distribution; we now have $2.8292 \leq M_0 \leq
+    7.9298$ with 90\% confidence, with median 4.7003.  The
+    distribution for $\Mmin$ is less affected, though it has broadened
+    somewhat toward low masses.}
+\end{figure}
+
+Figure \ref{fig:gaussian-high} displays the marginalized distributions
+for the Gaussian parameters (Section \ref{sec:gaussian}) when the
+higher-mass objects are included in the mass distribution.  The mean
+mass, $\mu$, and the mass standard deviation, $\sigma$, are both
+increased relative to Figure \ref{fig:gaussian} to account for the
+broader distribution and higher-mass tail.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/gaussian-high}
+  \end{center}
+  \caption{\label{fig:gaussian-high} The marginalized distributions
+    for the Gaussian parameters when the higher-mass objects are
+    included in the mass distribution.  The mean mass, $\mu$, and the
+    mass standard deviation, $\sigma$, are both increased relative to
+    Figure \ref{fig:gaussian} to account for the broader distribution
+    and higher-mass tail.  The peak of the underlying mass
+    distribution lies in the range $7.8660 \leq \mu \leq 10.9836$ with
+    90\% confidence; the median value is 9.2012.}
+\end{figure}
+
+The analysis of the two-Gaussian model shows the largest change when
+the higher-mass samples are included.  Figure
+\ref{fig:two-gaussian-high} shows the marginalized distributions for
+the two-Gaussian parameters (Section \ref{sec:gaussian}) when the
+higher-mass samples are included in the analysis.  In stark contrast
+to Figure \ref{fig:two-gaussian}, there are two well-defined,
+separated peaks; the lower-mass peak reproduces the results from the
+low-mass samples, while the higher-mass peak ($13.5534 \leq \mu_2 \leq
+27.9481$ with 90\% confidence; median 20.3839) matches the new
+higher-mass samples.  The peak in $\alpha$ near 0.8 is consistent with
+approximately 4/5 the total probability being concentrated in the 15
+low-mass samples.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/two-gaussian-high}
+  \end{center}
+  \caption{\label{fig:two-gaussian-high} The marginalized
+    distributions for the two-Gaussian parameters (Section
+    \ref{sec:gaussian}) when the higher-mass samples are included in
+    the analysis.  In stark contrast to Figure \ref{fig:two-gaussian},
+    there are two well-defined, separated peaks; the lower-mass peak
+    reproduces the results from the low-mass samples, while the
+    higher-mass peak ($13.5534 \leq \mu_2 \leq 27.9481$ with 90\%
+    confidence; median 20.3839) matches the new higher-mass samples.
+    The peak in $\alpha$ near 0.8 is consistent with approximately 15
+    out of 20 samples belonging to the low-mass peak.}
+\end{figure}
+
+The marginalized distributions for the log-normal parameters (Section
+\ref{sec:log-normal}) when the higher-mass samples are included in the
+analysis are displayed in Figure \ref{fig:log-normal-high}.  The
+changes when the higher-mass samples are included (compare to Figure
+\ref{fig:log-normal}) are similar to the changes in the Gaussian
+distribution: the mean mass moves to higher masses, and the
+distribution broadens.  Because the log-normal distribution is
+inherently asymmetric, with a high-mass tail, it does not need to
+widen as much as the Gaussian distribution did.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/log-normal-high}
+  \end{center}
+  \caption{\label{fig:log-normal-high} The marginalized distributions
+    for the log-normal parameters (Section \ref{sec:log-normal}) when
+    the higher-mass samples are included in the analysis.  The changes
+    when the higher-mass samples are included (compare to Figure
+    \ref{fig:log-normal}) are similar to the changes in the Gaussian
+    distribution: the mean mass moves to higher masses, and the
+    distribution broadens.}
+\end{figure}
+
+The confidence limits on the parameters for the parametric models of
+the underlying mass distribution are displayed in Table
+\ref{tab:high-mass-parametric} (compare to Table
+\ref{tab:low-mass-parametric}).
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{|l|c|c|c|c|c|c|}
+      \hline
+      Model & Parameter & 5\% & 15\% & 50\% & 85\% & 95\% \\
+      \hline \hline
+      Power Law (Equation \eqref{eq:power-law-dist}) & $\Mmin$ & 
+      4.87141 & 5.29031 & 5.85019 & 6.26118 & 6.45674 \\
+      \hline
+      & $\Mmax$ & 19.1097 & 23.4242 & 31.5726 & 37.7519 & 39.3369 \\
+      \hline
+      & $\alpha$ & -5.04879 & -4.30368 & -3.23404 & -2.31365 & -1.77137 \\
+      \hline \hline
+      Decaying Exponential (Equation \eqref{eq:exp-def}) & $\Mmin$ & 
+      4.0865 & 4.60236 & 5.32683 & 5.94097 & 6.22952 \\
+      \hline
+      & $M_0$ & 2.82924 & 3.41139 & 4.70034 & 6.52214 & 7.92979 \\
+      \hline \hline
+      Gaussian (Equation \eqref{eq:gaussian-def}) & $\mu$ & 
+      7.86599 & 8.33118 & 9.20116 & 10.2493 & 10.9836 \\
+      \hline
+      & $\sigma$ & 2.23643 & 2.58899 & 3.33545 & 4.17886 & 4.67881 \\
+      \hline \hline
+      Two Gaussian (Equation \eqref{eq:two-gaussian-def}) & $\mu_1$ & 
+      6.741 & 7.02724 & 7.48174 & 8.0139 & 8.46626 \\
+      \hline
+      & $\mu_2$ & 13.5534 & 16.202 & 20.3839 & 24.9259 & 27.9481 \\
+      \hline
+      & $\sigma_1$ & 0.742824 & 0.913941 & 1.31244 & 1.94862 & 2.50238 \\
+      \hline
+      & $\sigma_2$ & 0.511159 & 1.5025 & 4.39824 & 7.04612 & 8.25905 \\
+      \hline
+      & $\alpha$ & 0.575692 & 0.670978 & 0.798227 & 0.891522 & 0.932143 \\
+      \hline \hline
+      Log Normal (Equation \eqref{eq:log-normal-def}) & $\langle M \rangle$ & 
+      8.00086 & 8.51192 & 9.6264 & 11.1851 & 12.3986 \\
+      \hline
+      & $\sigma_M$ & 2.19262 & 2.8137 & 4.16742 & 6.25101 & 8.11839 \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:high-mass-parametric} Quantiles of the
+    marginalized distribution for each of the parameters in the models
+    discussed in Section \ref{sec:parametric-models} when the
+    higher-mass samples are included in the analysis (compare to Table
+    \ref{tab:low-mass-parametric}).  We indicate the 5\%, 15\%, 50\% (median), 85\%, and 95\% quantiles.}
+\end{table}
+
+The non-parametric (histogram; see Section \ref{sec:non-parametric-models}) models
+also show evidence of a long tail from the inclusion of the
+higher-mass samples.  Table \ref{tab:high-mass-non-parametric}
+displays confidence limits on the histogram parameters for the
+analysis including the higher-mass systems; compare to Table
+\ref{tab:low-mass-non-parametric}.
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{|c|c|c|c|c|c|c|}
+      \hline
+      Bins & Boundary & 5\% & 15\% & 50\% & 85\% & 95\% \\
+      \hline \hline
+      \hline
+      1 & $w_0$ & 2.22294 & 3.12695 & 4.2456 & 5.15132 & 5.58265 \\
+      \hline
+      & $w_1$ & 15.93 & 16.2535 & 17.7836 & 20.5449 & 22.5836 \\
+      \hline \hline
+      2 & $w_0$ & 3.87202 & 4.49983 & 5.41234 & 6.08334 & 6.35933 \\
+      \hline
+      & $w_1$ & 7.22163 & 8.25079 & 8.93669 & 9.71551 & 10.4287 \\
+      \hline
+      & $w_2$ & 18.4762 & 19.9798 & 24.941 & 32.5972 & 36.8615 \\
+      \hline \hline
+      3 & $w_0$ & 3.39289 & 4.24509 & 5.41694 & 6.15087 & 6.42822 \\
+      \hline
+      & $w_1$ & 6.41849 & 6.71984 & 7.47263 & 8.2942 & 8.61785 \\
+      \hline
+      & $w_2$ & 8.41449 & 8.64664 & 9.17056 & 10.4075 & 12.2718 \\
+      \hline
+      & $w_3$ & 18.5705 & 21.0481 & 27.1494 & 34.7753 & 38.0652 \\
+      \hline \hline
+      4 & $w_0$ & 2.42094 & 3.69875 & 5.2596 & 6.25449 & 6.54316 \\
+      \hline
+      & $w_1$ & 5.83725 & 6.2836 & 6.84987 & 7.8033 & 8.27706 \\
+      \hline
+      & $w_2$ & 6.94919 & 7.43628 & 8.38531 & 9.13401 & 9.91845 \\
+      \hline
+      & $w_3$ & 8.50371 & 8.75188 & 9.86694 & 17.1848 & 22.1086 \\
+      \hline
+      & $w_4$ & 18.5823 & 21.4628 & 28.367 & 35.8118 & 38.5278 \\      
+      \hline \hline
+      5 & $w_0$ & 1.73691 & 3.19184 & 4.89769 & 5.9547 & 6.35522 \\
+      \hline
+      & $w_1$ & 5.46124 & 5.95881 & 6.59431 & 7.26795 & 7.91821 \\
+      \hline
+      & $w_2$ & 6.63468 & 6.9804 & 7.93239 & 8.60918 & 9.06926 \\
+      \hline
+      & $w_3$ & 7.89654 & 8.35634 & 8.91766 & 10.6568 & 13.9644 \\
+      \hline
+      & $w_4$ & 8.74064 & 9.42672 & 15.8004 & 22.7101 & 27.6399 \\
+      \hline
+      & $w_5$ & 20.0202 & 22.9065 & 29.6307 & 36.6606 & 38.8573 \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:high-mass-non-parametric} The 5\%, 15\%, 50\%
+    (median), 85\%, and 95\% quantiles for the bin boundaries in the
+    one- through five-bin histogram models discussed in Section
+    \ref{sec:non-parametric-models} in an analysis including the higher-mass, wind-fed systems.
+The tails evident in Figure \ref{fig:high-mass-dists} are apparent here as well; compare to Table \ref{tab:low-mass-non-parametric}.}
+\end{table}
+
+\subsection{Model Selection}
+
+Repeating the model selection analysis discussed in Section
+\ref{sec:model-selection} for the sample including the higher-mass
+systems, we find that the model probabilities have changed with the
+inclusion of the extra five systems.  This is conclusive evidence (as
+if more were needed) that the higher-mass, wind-fed systems come from
+a different underlying population than the lower-mass, Roche-lobe
+overflow systems.  As before, we assume for this analysis that the
+model priors are equal.
+
+The harmonic mean estimators for the model probabilities are displayed
+in Figure \ref{fig:high-harmonic-evidence}; compare Figure
+\ref{fig:harmonic-evidence}.  Reversible jump MCMC calculations of the
+model probabilities are displayed in Figure
+\ref{fig:high-rj-evidence}; compare Figure \ref{fig:rj}.  Though the
+resulting rankings are not in complete agreement, a striking feature
+of both calculations is that the two-Gaussian model is much more
+likely with the higher-mass samples.  In the harmonic mean
+calculation, the power-law model is the most likely; in the
+reversible-jump calculation the exponential model is most likely.  In
+either case, the most likely model differs significantly from the most
+likely model for the lower-mass samples.  The model selection provides
+more strong evidence that the wind-fed and Roche-lobe systems come
+from different underlying populations.  The improvement of the
+exponential model relative to the lower-mass analysis (it is the most
+likely model in the reversible-jump MCMC) is encouraging for
+theoretical calculations that attempt to model the entire population
+of X-ray binaries with this mass model.  Note also that the increasing
+structure of the mass distribution favors histogram models with three
+bins over those with fewer bins.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/evidence-high}
+  \end{center}
+  \caption{\label{fig:high-harmonic-evidence} Harmonic mean evidence
+    for the various models from Section \ref{sec:models} applied to
+    all 20 systems in Table \ref{tab:sources} (see Section
+    \ref{sec:direct-evidence}).  In increasing order along the
+    $x$-axis, the models are the power-law of Section
+    \ref{sec:power-law} (PL), the decaying exponential of Section
+    \ref{sec:exponential} (E), the single Gaussian of Section
+    \ref{sec:gaussian} (G), the double Gaussian of Section
+    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
+    five-bin histogram models of Section
+    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
+    respectively).  The errorbars are bootstrap estimates of the 10\%
+    and 90\% bounds on the harmonic mean estimator.  Compare to Figure
+    \ref{fig:harmonic-evidence}.}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/rj-high}
+  \end{center}
+  \caption{\label{fig:high-rj-evidence} The relative probability of
+    the models discussed in Section \ref{sec:models} as computed using
+    the reversible-jump MCMC with the efficient jump proposal
+    algorithm described in Section \ref{sec:reversible-jump-mcmc},
+    applied to all 20 systems in Table \ref{tab:sources} (i.e.\
+    including the higher-mass systems).  In increasing order along the
+    $x$-axis, the models are the power-law of Section
+    \ref{sec:power-law} (PL), the decaying exponential of Section
+    \ref{sec:exponential} (E), the single Gaussian of Section
+    \ref{sec:gaussian} (G), the double Gaussian of Section
+    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
+    five-bin histogram models of Section
+    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
+    respectively).  The average of 500 independent reversible-jump
+    MCMCs is plotted, along with the 1-$\sigma$ error on the average
+    inferred from the standard deviation of the probability from the
+    individual MCMCs.  Compare to Figure \ref{fig:rj}.}
+\end{figure}
+
 \section{The Minimum Mass of the Black Hole Mass Distribution}
@@
 
-Using the MCMC samples for the parameters of the distributions in
-Section \ref{sec:models}, we can form probability distributions for
-the minimum black hole mass implied by the model posteriors.  (For the
-Gaussian models without a sharp cutoff we use the mass corresponding
-to the 1\% quantile of the distribution as the ``minimum black hole
-mass.'')  Figure \ref{fig:min-mass} displays the probability
-distributions for the minimum mass in the various models.
-
-\begin{figure}
-  \begin{center}
-    \plottwo{plots/mmin-parameteric}{plots/mmin-non-parameteric}
+It is interesting to use our models for the underlying mass
+distribution of X-ray binary black holes to try to place constraints
+on the minimum black hole mass.  \citet{Bailyn1998} addressed this
+question in the context of a ``mass gap'' between the most massive
+neutron stars and the least massive black holes.  The more recent
+study of \citet{Ozel2010} also looked for a mass gap using a subset of
+the models and systems presented here.  Both works found that the
+minimum black hole mass is significantly above the maximum neutron
+star mass.
+
+The distribution of the minimum black hole mass from the analysis of
+the lower-mass samples is displayed in Figure \ref{fig:min-mass}.  For
+all distributions, the minimum black hole mass is defined as the 1\%
+mass quantile (i.e.\ the mass lying below 99\% of the mass
+distribution).  (A cutoff quantile is necessary in the case of those
+distributions that do not have a hard cutoff mass; even for those that
+do, like the power-law model, it can be useful to define a ``soft''
+cutoff in the event that the lower mass hard cutoff becomes an
+irrelevant parameter is discussed in Section \ref{sec:power-law}.
+
+\begin{figure}
+  \begin{center}
+    \plottwo{plots/mmin-parametric}{plots/mmin-non-parametric}
   \end{center}
@@
     black hole mass calculated from the MCMC samples for the models in
-    Section \ref{sec:models}.  For the Gaussian distributions without
-    a sharp cutoff, we use the mass corresponding to the 1\% mass
-    quantile for the ``minimum black hole mass.''  For the three
-    most-favored models (the power law, one-, and two-bin histogram),
-    the minimum black hole mass is above 5.00, 3.23, and 3.82 solar
-    masses, respectively (at 95\% confidence). }
-\end{figure}
-
-For the three most-favored models (the power law, one-, and two-bin
-histograms), the minimum black hole mass is above 5.00, 3.23, and 3.82
-solar masses, respectively, with 95\% confidence.  While these limits
-are lower than those of a previous study of fewer systems
-\citep{Bailyn1998}, they are still significantly above the maximum
-neutron star mass.  The measured masses of the systems in Section
-\ref{sec:systems} show strong evidence of a ``gap'' between the
-maximum neutron star and minimum black hole mass.
+    Section \ref{sec:models}.  The minimum black hole mass is defined
+    as the 1\% mass quantile.  For the most favored models, the
+    power-law and Gaussian, the 90\% confidence limit on the minimum
+    black hole mass is 4.3 $\Msun$ and 2.9 $\Msun$, respectively.}
+\end{figure}
+
+We find that the best-fit models (the power-law and Gaussian) have a
+90\% confidence limit on the minimum black hole mass of 4.3 $\Msun$
+and 2.9 $\Msun$, respectively.  These are both significantly above the
+maximum neutron star mass, so we conclude that the lower-mass systems
+show strong evidence of a mass gap.
+
+The distribution of minimum black hole masses for the analysis of the
+complete sample (i.e.\ including the higher-mass systems) is shown in
+Figure \ref{fig:high-min-mass}.  For the most favored models, the
+power-law, exponential and two-Gaussian, the 90\% confidence limit on
+the minimum black hole mass is 5.2 $\Msun$, 4.5 $\Msun$, and 2.2
+$\Msun$, respectively.  Except for the two-Gaussian model, these are
+also well above the maximum neutron star mass.  Recall that the
+two-Gaussian model is forced to have a wide, symmetric, high-mass peak
+to match the high-mass samples when, in fact, the mass distribution is
+asymmetric.  This symmetry of the Gaussian forces the distribution
+artificially high at low masses; it seems likely that the low minimum
+black hole mass for the two-Gaussian is not a true feature of the
+black hole mass distribution, but rather due to this effect.  We
+therefore conclude that there is strong evidence for a mass gap in the
+complete sample as well.
+
+\begin{figure}
+\begin{center}
+    \plottwo{plots/mmin-parametric-high}{plots/mmin-non-parametric-high}
+  \end{center}
+  \caption{\label{fig:high-min-mass} The distributions for the minimum
+    black hole mass calculated from the MCMC samples for the models in
+    Section \ref{sec:models} using the complete sample of systems.
+    The minimum black hole mass is defined as the 1\% mass quantile.
+    For the most favored models, the power-law, exponential and
+    two-Gaussian, the 90\% confidence limit on the minimum black hole
+    mass is 5.2 $\Msun$, 4.5 $\Msun$, and 2.2 $\Msun$, respectively.}
+\end{figure}
+
+It is interesting that the minimum black hole mass for the power-law
+model \emph{increases} significantly when the higher-mass samples are
+added.  We hypothesize that the minimum black hole mass for the
+lower-mass samples alone is artificially lowered due to the
+correlations among the power-law parameters that lead to a wide
+distribution of power-law slopes discussed in Section
+\ref{sec:power-law}, and that the higher-mass result may be more
+indicative of the true minimum mass of the distribution.  In either
+case, however, the minimum mass is significantly above the maximum
+neutron star mass.
 

### 2010-10-20 paper.tex
First stab at introduction.

@@
 \label{sec:intro}
+
+(X-Ray binaries are important---supernova, evolutionary processes, GW,
+etc.)
+
+\citet{Bailyn1998} examined a sample of seven transient X-ray binaries
+thought to contain a black hole, concluding in a Bayesian analysis
+that the mass function was strongly peaked around seven solar masses.
+They found strong evidence for a ``gap'' between the least massive
+black hole and the upper limit for neutron star masses of 3 $\Msun$.
+
+A more recent study \citep{Ozel2010}, also in a Bayesian framework,
+largely confirmed these findings, examining 16 low-mass X-ray binary
+systems containing black holes and finding a strongly peaked
+distribution at $7.8 \pm 1.2 \, \Msun$.  \citet{Ozel2010} used two
+models for the mass function: a Gaussian and a decaying exponential
+with a minimum ``turn-on'' mass (motivated by the analytical model of
+the black-hole mass function in \citet{Fryer2001}), finding in both
+cases strong evidence of a mass gap.  They argue that this mass gap
+cannot be explained by observational selection effects.
+
+In this work, we use observations of 20 low- and higher-mass galactic
+X-ray binaries that contain black holes to constrain the mass function
+for stellar mass black holes in the galaxy.  We consider separately
+the 15 low-mass X-ray binaries containing Roche-lobe-filling
+secondaries with masses smaller than the black hole primary, and a
+complete sample including these systems and five higher-mass, wind-fed
+systems where the secondary is heavier than the black hole primary.
+
+We use a Bayesian Markov-Chain Monte Carlo analysis to explore various
+models for the black hole mass function for both samples, including
+those of \citet{Ozel2010} and \citet{Bailyn1998}.  We include both
+parameteric models, such as a Gaussian, and non-parameteric models
+where the mass function is represented by histograms with various
+numbers of bins.  After computing posterior distributions for the
+model parameters, we use model selection techniques (including a new
+technique for efficient reversible-jump MCMC \citep{Farr2010}) to
+compare the evidence for the various models from both samples.
+
+We find, like \citet{Ozel2010} and \citet{Bailyn1998}, strong evidence
+for a mass gap among the best models for both systems, with the best
+model for the lower-mass systems giving a minimum black hole mass of
+4.3 $\Msun$ (at 90\% confidence), and the best model for the combined
+sample of lower- and high-mass systems giving a minimum black hole
+mass of 4.5 $\Msun$ (at 90\% confidence).  Among the models with lower
+evidence, most also have a mass gap.
+
+The theoretical model from \citet{Fryer2001}, a decaying exponential,
+considered for the low-mass systems in \citet{Ozel2010}, has
+relatively low evidence.  We find that the lower-mass systems are best
+described by a Gaussian (the other model from \citet{Ozel2010}).  We
+find that the theoretical model from \citet{Fryer2001} is, however,
+the preferred model for the combined sample, while a Gaussian has much
+lower evidence in this case.  A model with two separate Gaussian peaks
+also has high evidence for the combined sample of systems, suggesting
+\fixme{differences in the distributions of lower- and high-mass X-ray
+  binary systems}.
+
+The structure of this paper is as follows \fixme{This will change
+  dramatically}.
 

### 2010-10-22 paper.tex
Made Vicky's brief changes.

@@
   X-ray binary systems undergoing Roche lobe overflow and five
-  higher-mass, wind-fed X-ray binary systems.  Modeling the mass
-  distribution both parametrically---as a power law, exponential,
-  gaussian, combination of two gaussians, or log-normal
-  distribution---and non-parametrically---as histograms with varying
-  numbers of bins---we explore the distribution in the context of
-  Markov Chain Monte Carlo calculations of model parameters.  We give
-  confidence bounds on the shape of the mass distribution in the
-  context of each model and compare the models with each other by
-  calculating their Bayesian evidence.  The mass distribution of the
-  low-mass systems is best fit by a power-law, while the distribution
-  of the complete sample is best fit by the exponential and power-law
-  models (but with significantly different parameters than the
-  lower-mass sample).  We address the existence of a ``gap'' between
-  the most massive neutron stars and least massive black holes,
-  finding that the best model (the power law) fitted to the low-mass
-  systems only gives a minimum black-hole mass of 4.3 $\Msun$ (90\%
-  confidence), while the best models (the exponential and power-law)
+  higher-mass, wind-fed X-ray binary systems.  Using MCMC
+  calculations, we model the mass distribution both
+  parametrically---as a power law, exponential, gaussian, combination
+  of two gaussians, or log-normal distribution---and
+  non-parametrically---as histograms with varying numbers of bins.  We
+  provide confidence bounds on the shape of the mass distribution in
+  the context of each model and compare the models with each other by
+  calculating their Bayesian evidence as supported by the
+  measurements.  The mass distribution of the low-mass systems is best
+  fit by a power-law, while the distribution of the complete sample is
+  best fit by the exponential model.  We examine the existence of a
+  ``gap'' between the most massive neutron stars and least massive
+  black holes, finding that the best model (the power law) fitted to
+  the low-mass systems only gives a minimum black-hole mass of 4.3
+  $\Msun$ (90\% confidence), while the best model (the exponential)
   fitted to all 20 systems gives a minumum black-hole mass of 4.5 and
@@
 rotational broadening of spectral lines from the secondary, but even
-in the ideal case (see, e.g., \citet{Cantrell2010} on A0620) this
-procedure is complicated.  In particular, contributions from an
+in the most studied case (see, e.g., \citet{Cantrell2010} on A0620)
+this procedure is complicated.  In particular, contributions from an
 accretion disk and hot spots in the disk can significantly distort the
 measured inclination and mass ratios.  For some systems (e.g.\ GS 1354
-\citep{Casares2009}) strong variability prevents determination of the
-inclination from the lightcurve at all; in these cases an upper limit
-on the inclination often comes from the observed lack of eclipses in
-the lightcurve.  In general, accurately determining $q$ and $i$
-requires a careful system-by-system analysis.
+\citep{Casares2009}) strong variability completely prevents
+determination of the inclination from the lightcurve; in these cases
+an upper limit on the inclination often comes from the observed lack
+of eclipses in the lightcurve.  In general, accurately determining $q$
+and $i$ requires a careful system-by-system analysis.
 
@@
 
-In Figure \ref{fig:masses} we show the combination of these individual
-mass distributions into one single distribution.  Note that this
-combination is not a good estimator of the true mass distribution of
-black hole masses.  To understand why this is not a good estimator of
-the underlying mass distribution, suppose that the observations
-implied an identical mass distribution for each object.  Then the
-combined distribution is equal to the individual distributions, but in
-fact we can be confident that each of the objects was drawn from a
-distribution that was narrower by a factor of approximately
-$\sqrt{N}$.  \citet{Mandel2010} discusses this issue in detail.  We
-include Figure \ref{fig:masses} here only to give a sense of the range
-of masses of black holes in our sample; we will come to estimators for
-the mass distribution.
-
-\begin{figure}
-  \plottwo{plots/masses}{plots/masses-high}
-
-  \caption{\label{fig:masses} The mass distributions in Figure
-    \ref{fig:low-masses} (left) and both \ref{fig:low-masses} and
-    \ref{fig:high-masses} (right) combined into a single
-    distribution. Note that this combination is not a good estimator
-    of the true mass distribution of black hole masses in the
-    galaxy---see, e.g., \citet{Mandel2010}.  We include it here only
-    to give a sense of the range and general shape of the mass
-    distribution of black holes in our sample.}
-\end{figure}
-
 \section{Models}
@@
 
-In this section we apply a Bayesian analysis to various models for the
-underlying mass distribution from which the low-mass systems in Table
-\ref{tab:sources} were drawn.  We will treat the combined low- and
-higher-mass systems in Section \ref{sec:higher-mass}.  The end result
-will be the probability distribution for the parameters of each model
-implied by the data from Section \ref{sec:systems} in combination with
-our prior assumptions about the probability distribution for the
-parameters.  Bayes' rule relates these quantities.  For a model with
-parameters $\vtheta$ in the presence of data $d$, Bayes' rule states
+In this section we apply a statistical analysis to various models for
+the underlying mass distribution from which the low-mass systems in
+Table \ref{tab:sources} were drawn.  We will treat the combined low-
+and higher-mass systems in Section \ref{sec:higher-mass}.  The end
+result will be the probability distribution for the parameters of each
+model implied by the data from Section \ref{sec:systems} in
+combination with our prior assumptions about the probability
+distribution for the parameters.  Bayes' rule relates these
+quantities.  For a model with parameters $\vtheta$ in the presence of
+data $d$, Bayes' rule states
 \begin{equation}
@@
 from the underlying mass distribution is equally likely to be
-observed, leaving the analysis of selection effects to future work.
-(During the preparation of this work, we became aware of
-\citet{Ozel2010}, which finds that selection effects do not greatly
-affect an analysis similar to ours.)
+observed.  (See also \citet{Ozel2010}.)
 
@@
 The broadness of the marginalized distribution for $\alpha$ comes from
-the need to match the relatively narrow peak in the mass distribution
-apparent in Figure \ref{fig:masses}.  When $\alpha$ is negative, the
-resulting mass distribution slopes down; $\Mmin$ is constrained to be
-near the lowest mass of the observed black holes, while $\Mmax$ is
-essentially irrelevant.  Conversely, when $\alpha$ is positive and the
-mass distribution slopes up, $\Mmax$ must be close to the largest mass
+the need to match the relatively narrow range in mass of the
+lower-mass systems.  When $\alpha$ is negative, the resulting mass
+distribution slopes down; $\Mmin$ is constrained to be near the lowest
+mass of the observed black holes, while $\Mmax$ is essentially
+irrelevant.  Conversely, when $\alpha$ is positive and the mass
+distribution slopes up, $\Mmax$ must be close to the largest mass
 observed, while $\Mmin$ is essentially irrelevant.  Figure
@@
 the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
-that the first Gaussian has found the peak mass samples in Figure
-\ref{fig:masses}.  The second Gaussian's parameter distributions are
-much broader.  The second Gaussian appears to be sampling the tail of
-the mass samples (see Figure \ref{fig:masses}).  In spite of the
-ability to match a more complicated distribution, we find that this
-model is strongly disfavored relative to the single-Gaussian model for
-this dataset: $p(\textnormal{Gaussian}|d) / p(\textnormal{Two
-  Gaussian}|d) \simeq 4.7$ (see Section \ref{sec:model-selection} for
-discussion).
+that the first Gaussian is centered around the peaks of the lower-mass
+distributions.  The second Gaussian's parameter distributions are much
+broader.  The second Gaussian appears to be sampling the tail of the
+mass samples.  In spite of the ability to match a more complicated
+distribution, we find that this model is strongly disfavored relative
+to the single-Gaussian model for this dataset:
+$p(\textnormal{Gaussian}|d) / p(\textnormal{Two Gaussian}|d) \simeq
+4.7$ (see Section \ref{sec:model-selection} for discussion).
 
@@
     are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
-    second Gaussian is attempting to fit the tail of the mass samples
-    (see Figure \ref{fig:masses}).  The extra degrees of freedom in
-    the distribution from the second Gaussian do not provide enough
-    extra fitting power to compensate for the increase in parmeter
-    space, however: the two-Gaussian model is disfavored relative to
-    the single Gaussian by a factor of $4.7$ on this dataset (see
-    Section \ref{sec:model-selection} for discussion).}
+    second Gaussian is attempting to fit the tail of the mass samples.
+    The extra degrees of freedom in the distribution from the second
+    Gaussian do not provide enough extra fitting power to compensate
+    for the increase in parmeter space, however: the two-Gaussian
+    model is disfavored relative to the single Gaussian by a factor of
+    $4.7$ on this dataset (see Section \ref{sec:model-selection} for
+    discussion).}
 \end{figure}
@@
 
-There are two broad ways to compare models in the context of the MCMC
-analysis of Section \ref{sec:models}.  The first way, which we discuss
-in Section \ref{sec:direct-evidence}, is to use the MCMC samples
-from the various models to estimate the single model evidence,
-$p(d|M_i)$, appearing in Equation \eqref{eq:model-evidence-def}.  The
-second way is to treat the model literally as a parameter in a second
-MCMC analysis.  In the second MCMC analysis, one proposes jumps not
-between different $\vtheta_i$ within a particular model, but from
-parameters $\vtheta_i$ in model $M_i$ to parameters $\vtheta_j$ in
-$M_j$.  The evidence for model $M_i$ is proportional to the number of
-times this MCMC has a sample in the $M_i$ parameter space.  This
-technique is called a ``reversible-jump MCMC'' \citep{Green1995}; we
-discuss it in Section \ref{sec:reversible-jump-mcmc}.
-
-\subsection{Direct Evidence Calculations}
-\label{sec:direct-evidence}
-
-To compute $p(M_i|d)$ from Equation \eqref{eq:model-evidence-def}
-using the MCMC samples we discussed in Section \ref{sec:models} we
-need a way to compute the single-model evidence, $p(d|M_i)$.  The
-remaining terms in Equation \eqref{eq:model-evidence-def} are either
-under our control (the prior $p(M_i)$) or a normalization factor that
-we can compute using Equation \eqref{eq:multi-model-evidence-def} once
-we have the single model evidence (the evidence $p(d)$).  Begin by
-re-writing Equation \eqref{eq:bayes-explicit-model}:
-\begin{equation}
-  \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} = \frac{p(\theta_i|M_i)}{p(d|M_i)}.
-\end{equation}
-Integrating both sides over all $\theta_i$, recalling that the prior
-is normalized, we have 
-\begin{equation}
-  \int d\theta_i\, \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} = \frac{1}{p(d|M_i)}.
-\end{equation}
-We can estimate the integral using our MCMC samples (which are
-distributed with local density $p(\theta_i|d,M_i)d\theta_i$) via
-\begin{equation}
-  \int d\theta_i\, \frac{p(\theta_i|d,M_i)}{p(d|\theta_i,M_i)} \approx
-  \frac{1}{N_i} \sum_j \frac{1}{p(d|\theta_{ij}, M_i)},
-\end{equation}
-where the $j$th MCMC parameter sample (from a total $N_i$) is
-$\theta_{ij}$.  Thus we have
-\begin{equation}
-  \frac{1}{p(d|M_i)} \approx \frac{1}{N_i} \sum_j \frac{1}{p(d|\theta_{ij}, M_i)}.
-\end{equation}
-This technique is known as the ``harmonic mean approximation'' for the
-single-model evidence $p(d|M_i)$.
-
-\citet{Weinberg2010} discusses techniques to estimate the integral
-over parameter space in Equation \eqref{eq:model-evidence-def} by
-adaptively partitioning the parameter volume into non-overlapping
-boxes surrounding the points in the MCMC sample.  We implemented the
-algorithms from that paper, but found them, in this case, to give
-results that depended strongly on the parameters in the algorithm and
-were sometimes inconsistent with both the harmonic mean calculations
-in this Section and the reversible jump calculations in Section
-\ref{sec:reversible-jump-mcmc}.  Since the harmonic mean and
-reversible jump calculations are in agreement with each other, we will
-not discuss these direct integration techniques further.
-
-The harmonic mean estimate of the evidence is known for having large
-statistical uncertainty due to the tendency for a few low-likelihood
-points to dominate the sum \citep{Weinberg2010}.  Therefore, it is
-useful to have multiple independent estimates of the evidence for a
-given model.  In Section \ref{sec:reversible-jump-mcmc}, we will
-discuss the reversible jump algorithm for estimating the evidence for
-our models; the results presented here are in broad agreement with the
-results in that Section.  In Figure \ref{fig:harmonic-evidence} we
-display the harmonic mean estimator for the evidence for the various
-models discussed in Section \ref{sec:models}.  In addition, we have
-performed 10000 calculations of the harmonic mean estimator using
-``mock'' MCMC data drawn randomly from our true MCMC with replacement.
-The errorbars in Figure \ref{fig:harmonic-evidence} indicate the 10\%
-and 90\% quantiles from this bootstrap calculation.  Recall that we
-assume that no model is a priori more likely than any other, so the
-model priors $p(M_i)$ are equal.
-
-\begin{figure}
-  \begin{center}
-    \plotone{plots/evidence}
-  \end{center}
-  \caption{\label{fig:harmonic-evidence} Harmonic mean evidence for
-    the various models from Section \ref{sec:models} (see Section
-    \ref{sec:direct-evidence}).  In increasing order along the
-    $x$-axis, the models are the power-law of Section
-    \ref{sec:power-law} (PL), the decaying exponential of Section
-    \ref{sec:exponential} (E), the single Gaussian of Section
-    \ref{sec:gaussian} (G), the double Gaussian of Section
-    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
-    five-bin histogram models of Section
-    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
-    respectively).  The errorbars are bootstrap estimates of the 10\%
-    and 90\% bounds on the harmonic mean estimator.  The power-law and
-    Gaussian are the favored models, with the single-bin histogram
-    coming third (but the 10-90\% bootstrap range for the one-bin
-    histogram includes the possibility of being the most favored
-    model).  This is not surprising, given the single-peaked nature of
-    the lower-mass sample.}
-\end{figure}
-
-The results of the harmonic mean calculation indicate that the
-power-law and Gaussian models are most favored, with the single-bin
-histogram coming third (though the 10-90\% range for the models
-includes the possibility of it being the most favored).  This is not
-surprising, given the single-peaked nature of the lower-mass sample.
-The log normal and exponential models are likely disfavored due to
-their long tails.  The multiple-bin histogram models are disfavored
-because the structure of the lower-mass mass distribution is not
-complicated enough to warrant the extra parameter space implied by the
-larger number of bins.  The two Gaussian model is disfavored because
-the distribution is single-peaked, so the second Gaussian is
-superfluous.
-
 \subsection{Reversible-Jump MCMC}
@@
 
-Another technique for computing $p(M_i|d)$ is the reversible-jump MCMC
-\citep{Green1995}.  For a reversible-jump MCMC, we consider a
-``super-model'' that encompasses all the models under consideration.
-The parameter space of the super-model consists of a discrete
-parameter that identifies the choice of model, $M_i$, and the
-continuous parameters appropriate for this model, $\vtheta_i$.  We
-denote a point in the super-model parameter space by $\{M_i,
-\vtheta_i\}$; each such point is a statement that, e.g., ``the
-underlying mass distribution is a Gaussian, with parameters $\mu$ and
-$\sigma$,'' or ``the underlying mass distribution is a triple-bin
-histogram with parameters $w_1$, $w_2$, $w_3$, and $w_4$,'' or ....
-To compare models, we are interested in the quantity (see Equation
+A powerful technique for computing $p(M_i|d)$ is the reversible-jump
+MCMC \citep{Green1995}.  Consider the ``super-model'' that encompasses
+all the models under consideration.  The parameter space of the
+super-model consists of a discrete parameter that identifies the
+choice of model, $M_i$, and the continuous parameters appropriate for
+this model, $\vtheta_i$.  We denote a point in the super-model
+parameter space by $\{M_i, \vtheta_i\}$; each such point is a
+statement that, e.g., ``the underlying mass distribution is a
+Gaussian, with parameters $\mu$ and $\sigma$,'' or ``the underlying
+mass distribution is a triple-bin histogram with parameters $w_1$,
+$w_2$, $w_3$, and $w_4$,'' or ....  To compare models, we are
+interested in the quantity (see Equation
 \eqref{eq:model-posterior-def})
@@
     one- through five-bin histogram models discussed in Section
-    \ref{sec:non-parametric-models} in an analysis including the higher-mass, wind-fed systems.
-The tails evident in Figure \ref{fig:high-mass-dists} are apparent here as well; compare to Table \ref{tab:low-mass-non-parametric}.}
+    \ref{sec:non-parametric-models} in an 
+    analysis including the higher-mass, wind-fed systems.  
+    The tails evident in Figure \ref{fig:high-mass-dists} are apparent
+    here as well; compare to Table \ref{tab:low-mass-non-parametric}.}
 \end{table}
@@
 
-The harmonic mean estimators for the model probabilities are displayed
-in Figure \ref{fig:high-harmonic-evidence}; compare Figure
-\ref{fig:harmonic-evidence}.  Reversible jump MCMC calculations of the
-model probabilities are displayed in Figure
-\ref{fig:high-rj-evidence}; compare Figure \ref{fig:rj}.  Though the
-resulting rankings are not in complete agreement, a striking feature
-of both calculations is that the two-Gaussian model is much more
-likely with the higher-mass samples.  In the harmonic mean
-calculation, the power-law model is the most likely; in the
-reversible-jump calculation the exponential model is most likely.  In
-either case, the most likely model differs significantly from the most
-likely model for the lower-mass samples.  The model selection provides
-more strong evidence that the wind-fed and Roche-lobe systems come
-from different underlying populations.  The improvement of the
-exponential model relative to the lower-mass analysis (it is the most
-likely model in the reversible-jump MCMC) is encouraging for
-theoretical calculations that attempt to model the entire population
-of X-ray binaries with this mass model.  Note also that the increasing
-structure of the mass distribution favors histogram models with three
-bins over those with fewer bins.
-
-\begin{figure}
-  \begin{center}
-    \plotone{plots/evidence-high}
-  \end{center}
-  \caption{\label{fig:high-harmonic-evidence} Harmonic mean evidence
-    for the various models from Section \ref{sec:models} applied to
-    all 20 systems in Table \ref{tab:sources} (see Section
-    \ref{sec:direct-evidence}).  In increasing order along the
-    $x$-axis, the models are the power-law of Section
-    \ref{sec:power-law} (PL), the decaying exponential of Section
-    \ref{sec:exponential} (E), the single Gaussian of Section
-    \ref{sec:gaussian} (G), the double Gaussian of Section
-    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
-    five-bin histogram models of Section
-    \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
-    respectively).  The errorbars are bootstrap estimates of the 10\%
-    and 90\% bounds on the harmonic mean estimator.  Compare to Figure
-    \ref{fig:harmonic-evidence}.}
-\end{figure}
+Reversible jump MCMC calculations of the model probabilities are
+displayed in Figure \ref{fig:high-rj-evidence}; compare Figure
+\ref{fig:rj}.  The exponential model is the most favored model for the
+complete sample, with the two-Gaussian model the second-most favored.
+The ranking of models differs significantly from the lower-mass
+samples.  The model selection provides more strong evidence that the
+wind-fed and Roche-lobe systems come from different underlying
+populations.  The improvement of the exponential model relative to the
+lower-mass analysis is encouraging for theoretical calculations that
+attempt to model the entire population of X-ray binaries with this
+mass model.  Note also that the increasing structure of the mass
+distribution favors histogram models with three bins over those with
+fewer bins.
 

### 2010-10-22 paper.tex
Charles' quick changes.

@@
 the analysis through both the shape and the range of prior support in
-parameter space.
+parameter space.  The prior should reflect the ``best guess'' for the
+distribution of parameters before examining any of the data.  In the
+absence of any information about the distribution of parameters, it is
+best to choose a prior that is broad and uniformative to avoid biasing
+the posterior as much as possible.
 
@@
 well-represented by a decaying exponential with a minimum mass:
+\fixme{Vicky, do you want to say anything more here?}
 \begin{equation}
@@
     \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
-    \right] & \Mmin \leq M \\
+    \right] & M \geq \Mmin \\
     0 & \textnormal{otherwise}
@@
   \begin{center}
-    \plottwo{plots/dist-parametric-high}{plots/dist-non-parametric-high}
+    \plottwo{plots/dist-parametric-high}{plots/dist-non-parametric-high-mass}
   \end{center}

### 2010-10-22 paper.tex
Ilya's quick changes.

@@
 
-\author{Ilya Mandel \and Vicky Kalogera}
+\author{Ilya Mandel\altaffilmark{1}}
+
+\affil{MIT Kavli Institute, Cambridge, MA 02139}
+
+\email{ilyamandel@chgk.info}
+
+\and
+
+\author{Vicky Kalogera}
 
@@
 
- \email{ilyamandel@chgk.info, vicky@northwestern.edu}
+ \email{vicky@northwestern.edu}
+
+ \altaffiltext{1}{School of Physics and Astronomy, University of
+   Birmingham, Edgbaston, Birmingham B15 2TT, UK}
 
@@
 within the quoted range.  Table \ref{tab:sources} gives the assumed
-distribution for the observables in the 17 systems we use.  We do not
+distribution for the observables in the 20 systems we use.  We do not
 attempt to deal with the systematic biases in the observational
@@
 equivalent to taking a finite section of an infinite chain that begins
-with the chosen point.  Since every point in parameter space occurs in
-an infinite chain, and no section of the chain is better than any
-other, our chosen finite chain will have the appropriate distribution.
-However, because consecutive samples in a chain are correlated with
-each other, the beginning of our finite chain has a ``memory'' of the
-starting point; we discard enough points at the beginning of the
-finite chain that we can be confident it does not retain a memory of
-the arbitrary starting point.  The points discarded in this way are
-commonly called ``burn-in'' points.
+with the chosen point.  Every point in parameter space occurs in an
+infinite chain, and no section of the chain is better than any other,
+so a sufficiently long, but finite, section of the infinite chain
+chosen in this manner can be representative of the statistics of the
+chain as a whole.  However, because consecutive samples in a chain are
+correlated with each other, the beginning of our finite chain has a
+``memory'' of the starting point; we discard enough points at the
+beginning of the finite chain that we can be confident it does not
+retain a memory of the arbitrary starting point.  The points discarded
+in this way are commonly called ``burn-in'' points.
 
@@
 
-Many astrophysical distributions are power-laws.  Let us assume that
+Many astrophysical distributions are power laws.  Let us assume that
 the BH mass distribution is given by
@@
 \end{equation}
-This is the total posterior probability found in the parameter space
+This is the integral of the posterior over the entire parameter space
 of model $M_i$.  The marginalized posterior probability of model $M_i$
@@
 obtain a chain of samples $\{M_i, \vtheta_i\}$ distributed in
-parameter space with density $p(M_i,\vtheta_i) d\vtheta_i$ and we can
+parameter space with density $p(M_i,\vtheta_i|d) d\vtheta_i$ and we can
 estimate the integral as
@@
 We have performed a suite of 500 independent reversible-jump MCMCs
-jumping between all the models (both parametric and non-parametric)
-in Section \ref{sec:models} using the single-model MCMC samples to
+jumping between all the models (both parametric and non-parametric) in
+Section \ref{sec:models} using the single-model MCMC samples to
 construct an efficient jump proposal for each model as described
@@
 errors on the average inferred from the standard deviation of the
-model counts across the suite.  Table \ref{tab:rj}
+model counts across the suite.  Table \ref{tab:rj} gives the numerical
+values of the average probability for each model across the suite of
+MCMCs.
 
@@
       \hline
-      Model & Relative Probability \\
+      Model & Relative Evidence \\
       \hline \hline
@@
 \ref{sec:exponential}) is only a factor of $\sim 3$ away from the
-most-favored Power Law model (Section \ref{sec:power-law}), which does
-not have theoretical support.  Is such support worth a factor of three
-in the model prior?  Alternately, we may say we know (in advance of
-any mass measurements) that black holes must exist with mass $\lesssim
+power law model (Section \ref{sec:power-law}), which does not have
+theoretical support.  Is such support worth a factor of three in the
+model prior?  Alternately, we may say we know (in advance of any mass
+measurements) that black holes must exist with mass $\lesssim
 10\Msun$; then we could, for example, impose a prior on the minimum
@@
 ``single, narrow peak'' mass distribution \citep{Bailyn1998,Ozel2010}.
-In this light, the difference between the evidence calculated from the
-harmonic mean estimator and the reversible-jump MCMC for the
-single-bin histogram is surprising; one would expect that a
-narrowly-peaked single-bin histogram would perform well (as it does
-with the harmoinc-mean estimator) instead of being disfavored relative
-to the other models (as with the reversible-jump MCMC).
+In this light, poor performance of the single-bin histogram is
+surprising.
 
@@
 cutoff in the event that the lower mass hard cutoff becomes an
-irrelevant parameter is discussed in Section \ref{sec:power-law}.
+irrelevant parameter as discussed in Section \ref{sec:power-law}.)
 
@@
 
-We find that the best-fit models (the power-law and Gaussian) have a
-90\% confidence limit on the minimum black hole mass of 4.3 $\Msun$
-and 2.9 $\Msun$, respectively.  These are both significantly above the
-maximum neutron star mass, so we conclude that the lower-mass systems
-show strong evidence of a mass gap.
+We find that the best-fit model (the power-law) has a 90\% confidence
+limit on the minimum black hole mass of 4.3 $\Msun$.  This is
+significantly above the maximum theoretically-allowed neutron star
+mass, $\sim 3 \Msun$ \citep{Kalogera1996}\fixme{Vicky, do you have a
+  more recent reference than this?}, so we conclude that the
+lower-mass systems show strong evidence of a mass gap.
 
@@
 complete sample (i.e.\ including the higher-mass systems) is shown in
-Figure \ref{fig:high-min-mass}.  For the most favored models, the
-power-law, exponential and two-Gaussian, the 90\% confidence limit on
-the minimum black hole mass is 5.2 $\Msun$, 4.5 $\Msun$, and 2.2
-$\Msun$, respectively.  Except for the two-Gaussian model, these are
-also well above the maximum neutron star mass.  Recall that the
-two-Gaussian model is forced to have a wide, symmetric, high-mass peak
-to match the high-mass samples when, in fact, the mass distribution is
-asymmetric.  This symmetry of the Gaussian forces the distribution
-artificially high at low masses; it seems likely that the low minimum
-black hole mass for the two-Gaussian is not a true feature of the
-black hole mass distribution, but rather due to this effect.  We
-therefore conclude that there is strong evidence for a mass gap in the
-complete sample as well.
+Figure \ref{fig:high-min-mass}.  For the most favored model, the
+exponential, the 90\% confidence limit on the minimum black hole mass
+is 4.5 $\Msun$.  We therefore conclude that there is strong evidence
+for a mass gap in the complete sample as well.
 
@@
 
-It is interesting that the minimum black hole mass for the power-law
-model \emph{increases} significantly when the higher-mass samples are
-added.  We hypothesize that the minimum black hole mass for the
-lower-mass samples alone is artificially lowered due to the
-correlations among the power-law parameters that lead to a wide
-distribution of power-law slopes discussed in Section
-\ref{sec:power-law}, and that the higher-mass result may be more
-indicative of the true minimum mass of the distribution.  In either
-case, however, the minimum mass is significantly above the maximum
-neutron star mass.
-
 \section{Conclusion}
@@
 
-This work was supported by grant \fixme{XXX}.  Calculations for this
-work were performed on the Northwestern Fugu cluster, which was
-partially funded by NSF MRI grant \fixme{XXX}.  We thank Jonathan Gair
-for helpful discussions.
+This work was supported by grant \fixme{XXX}.  IM acknowledges support
+from the NSF AAPF under award AST-0901985.  Calculations for this work
+were performed on the Northwestern Fugu cluster, which was partially
+funded by NSF MRI grant \fixme{XXX}.  We thank Jonathan Gair for
+helpful discussions.
 

### 2010-10-23 paper.tex
Major re-ordering, added conclusion.  Shipping as v3 (but figures still need work).

@@
 
- \altaffiltext{1}{School of Physics and Astronomy, University of
+ \altaffiltext{1}{Also: School of Physics and Astronomy, University of
    Birmingham, Edgbaston, Birmingham B15 2TT, UK}
@@
   X-ray binary systems undergoing Roche lobe overflow and five
-  higher-mass, wind-fed X-ray binary systems.  Using MCMC
-  calculations, we model the mass distribution both
+  higher-mass, wind-fed X-ray binary systems.  Using Markov Chain
+  Monte Carlo calculations, we model the mass distribution both
   parametrically---as a power law, exponential, gaussian, combination
@@
   the context of each model and compare the models with each other by
-  calculating their Bayesian evidence as supported by the
+  calculating their relative Bayesian evidence as supported by the
   measurements.  The mass distribution of the low-mass systems is best
@@
   $\Msun$ (90\% confidence), while the best model (the exponential)
-  fitted to all 20 systems gives a minumum black-hole mass of 4.5 and
-  5.2 $\Msun$ (90\% confidence), respectively.  We conclude that our
-  sample of black hole masses provides strong evidence of a gap
-  between the maximum neutron star mass and minimum black hole mass.
+  fitted to all 20 systems gives a minumum black-hole mass of 4.5
+  $\Msun$ (90\% confidence).  We conclude that our sample of black
+  hole masses provides strong evidence of a gap between the maximum
+  neutron star mass and minimum black hole mass.
 \end{abstract}
@@
 
-(X-Ray binaries are important---supernova, evolutionary processes, GW,
-etc.)
-
-\citet{Bailyn1998} examined a sample of seven transient X-ray binaries
-thought to contain a black hole, concluding in a Bayesian analysis
-that the mass function was strongly peaked around seven solar masses.
-They found strong evidence for a ``gap'' between the least massive
-black hole and the upper limit for neutron star masses of 3 $\Msun$.
+The most massive stars probably end their lives with a supernova
+explosion, becoming stellar-mass black holes.  The mass distribution
+of such black holes can provide important clues to the end stages of
+evolution of these stars.  In addition, the mass distribution of
+stellar-mass black holes is an important input in calculations of
+rates of gravitational wave emission events from coalescing neutron
+star-black hole and black hole-black hole binaries in the LIGO
+gravitational wave observatory \citep{Abadie2010}.
+
+Observations of low-mass X-ray binaries in both the optical and X-ray
+can provide a measurement of the mass of the compact object in these
+systems.  There are now enough such measurements where the mass of the
+compact object is large enough to classify it as a black hole to
+attempt to statistically reconstruct the mass distribution of these
+stellar-mass black holes.
+
+The first study of the mass distribution of stellar-mass black holes,
+in \citet{Bailyn1998}, examined a sample of seven transient X-ray
+binaries thought to contain a black hole, concluding in a Bayesian
+analysis that the mass function was strongly peaked around seven solar
+masses.  \citet{Bailyn1998} found evidence of a ``gap'' between the
+least massive black hole and the upper limit for neutron star masses
+of 3 $\Msun$ \citep{Kalogera1996}.  Such a gap is puzzling in light of
+theoretical studies that predict a continuous distribution of compact
+object supernova remnant masses with a smooth transition from neutron
+stars to black holes \citep{Fryer2001}.
 
@@
 systems containing black holes and finding a strongly peaked
-distribution at $7.8 \pm 1.2 \, \Msun$.  \citet{Ozel2010} used two
+distribution at $7.8 \pm 1.2 \, \Msun$.  \citet{Ozel2010} use two
 models for the mass function: a Gaussian and a decaying exponential
 with a minimum ``turn-on'' mass (motivated by the analytical model of
-the black-hole mass function in \citet{Fryer2001}), finding in both
-cases strong evidence of a mass gap.  They argue that this mass gap
-cannot be explained by observational selection effects.
-
-In this work, we use observations of 20 low- and higher-mass galactic
-X-ray binaries that contain black holes to constrain the mass function
-for stellar mass black holes in the galaxy.  We consider separately
-the 15 low-mass X-ray binaries containing Roche-lobe-filling
-secondaries with masses smaller than the black hole primary, and a
-complete sample including these systems and five higher-mass, wind-fed
-systems where the secondary is heavier than the black hole primary.
+the black-hole mass function in \citet{Fryer2001}).  \citet{Ozel2010}
+do not provide confidence limits for the minimum black hole mass,
+instead discussing only the maximum-posterior parameters for their
+models.  Nevertheless, it appears that their analysis confirms the
+existence of a mass gap.  \citet{Ozel2010} discuss possible selection
+effects that could lead to the appearance of a mass gap, but conclude
+these effects could not produce the observed gap, which they therefore
+claim is a real property of the black hole mass distribution.
+
+In this work, we use observations of 20 galactic X-ray binaries that
+contain black holes to constrain the mass function for stellar mass
+black holes.  We consider separately 15 low-mass X-ray binaries
+containing Roche-lobe-filling secondaries with masses smaller than the
+black hole primary, and a sample including these systems and five
+higher-mass, wind-fed systems where the secondary is heavier than the
+black hole primary.
 
 We use a Bayesian Markov-Chain Monte Carlo analysis to explore various
-models for the black hole mass function for both samples, including
-those of \citet{Ozel2010} and \citet{Bailyn1998}.  We include both
-parameteric models, such as a Gaussian, and non-parameteric models
-where the mass function is represented by histograms with various
-numbers of bins.  After computing posterior distributions for the
-model parameters, we use model selection techniques (including a new
-technique for efficient reversible-jump MCMC \citep{Farr2010}) to
-compare the evidence for the various models from both samples.
+models for the black hole mass function for both samples.  We include
+both parameteric models, such as a Gaussian, and non-parameteric
+models where the mass function is represented by histograms with
+various numbers of bins.  (Our set of models includes those of
+\citet{Ozel2010} and \citet{Bailyn1998}.)  After computing posterior
+distributions for the model parameters, we use model selection
+techniques (including a new technique for efficient reversible-jump
+MCMC \citep{Farr2010}) to compare the evidence for the various models
+from both samples.
 
@@
 The theoretical model from \citet{Fryer2001}, a decaying exponential,
-considered for the low-mass systems in \citet{Ozel2010}, has
-relatively low evidence.  We find that the lower-mass systems are best
-described by a Gaussian (the other model from \citet{Ozel2010}).  We
-find that the theoretical model from \citet{Fryer2001} is, however,
-the preferred model for the combined sample, while a Gaussian has much
-lower evidence in this case.  A model with two separate Gaussian peaks
-also has high evidence for the combined sample of systems, suggesting
-\fixme{differences in the distributions of lower- and high-mass X-ray
-  binary systems}.
-
-The structure of this paper is as follows \fixme{This will change
-  dramatically}.
+considered for the low-mass systems in \citet{Ozel2010}, is strongly
+disfavored by our model selection.  We find that the lower-mass
+systems are best described by a power law, followed closely by a
+Gaussian (which is the other model from \citet{Ozel2010}).  We find
+that the theoretical model from \citet{Fryer2001} is, however, the
+preferred model for the combined sample of lower- and high-mass
+systems.  A model with two separate Gaussian peaks also has high
+evidence for the combined sample of systems.  While the failure of the
+theoretically-motivated model to explain the lower-mass, Roche
+lobe-filling systems suggests gaps in our understanding of the
+formation of these systems, it is encouraging for population synthesis
+studies \fixme{Vicky, do you have a couple of go-to papers for
+  pop. synth we could cite here?} that the model of \citet{Fryer2001}
+is a good fit to the full sample including both lower- and high-mass
+systems.
 
@@
 which the black hole is fed via winds (these systems all have a
-secondary that appears more massive than the black hole).  The low-
-and higher-mass systems undoubtedly have different evolutionary
+secondary that appears to be more massive than the black hole).  The
+low- and higher-mass systems undoubtedly have different evolutionary
 tracks, and therefore it is reasonable that they would have different
 black-hole mass distributions.  We will first analyze the 15 low-mass
-systems alone (Sections \ref{sec:models} and
-\ref{sec:model-selection}), and then the complete sample of 20 systems
-(Section \ref{sec:higher-mass}).
+systems alone (Section \ref{sec:results-lower-mass}), and then the
+complete sample of 20 systems (Section \ref{sec:higher-mass}).
 
 In each of these systems, spectroscopic measurements of the secondary
-star provide an orbital period for the system and a half-amplitude for
+star provide an orbital period for the system and a semi-amplitude for
 the secondary's velocity curve.  These measurements can be combined
@@
 mass distribution away from the ``best-fit'' mass in the
-observations).  This careful treatment will appear in future work.
+observations).  This treatment will appear in future work.
 
@@
 work are shown in Figure \ref{fig:low-masses} and Figure
-\ref{fig:high-masses}.  Note that systems for which $i$ is poorly
-constrained have broad ``tails'' on their mass distributions.  These
-mass distributions constitute the ``observational data'' we will use
-in the remainder of this paper.
+\ref{fig:high-masses}.  Systems for which $i$ is poorly constrained
+have broad ``tails'' on their mass distributions.  These mass
+distributions constitute the ``observational data'' we will use in the
+remainder of this paper.
 
@@
 
-\section{Models}
+\section{Statistical Analysis}
 \label{sec:models}
 
-In this section we apply a statistical analysis to various models for
-the underlying mass distribution from which the low-mass systems in
-Table \ref{tab:sources} were drawn.  We will treat the combined low-
-and higher-mass systems in Section \ref{sec:higher-mass}.  The end
-result will be the probability distribution for the parameters of each
-model implied by the data from Section \ref{sec:systems} in
-combination with our prior assumptions about the probability
-distribution for the parameters.  Bayes' rule relates these
-quantities.  For a model with parameters $\vtheta$ in the presence of
-data $d$, Bayes' rule states
+In this section we describe the statistical analysis we will apply to
+various models for the underlying mass distribution from which the
+low-mass systems in Table \ref{tab:sources} were drawn.  Results of
+the analysis appear in Section \ref{sec:results}.
+
+\subsection{Bayesian Inference}
+
+The end result of our statistical analysis will be the probability
+distribution for the parameters of each model implied by the data from
+Section \ref{sec:systems} in combination with our prior assumptions
+about the probability distribution for the parameters.  Bayes' rule
+relates these quantities.  For a model with parameters $\vtheta$ in
+the presence of data $d$, Bayes' rule states
 \begin{equation}
@@
 In our context, the data are the mass distributions given in Section
-\ref{sec:systems}: $d = \{ p_i(M)| i = 1, 2, \ldots, 17 \}$.  We
+\ref{sec:systems}: $d = \{ p_i(M)| i = 1, 2, \ldots, 20 \}$.  We
 assume that the measurements in Section \ref{sec:systems} are
@@
 from the underlying mass distribution is equally likely to be
-observed.  (See also \citet{Ozel2010}.)
+observed.  The results of \citet{Ozel2010} suggest that selection
+effects are unlikely to significantly bias our analysis.
 
 For a mass distribution with several parameters, $p(\vtheta | d)$
-lives in a multi-dimensional space.  Exploring the entirety of this
-space systematically rapidly becomes prohibitive as the number of
-parameters increases.  A more efficient way to explore the
-distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
-(MCMC) method.  MCMC methods produce a chain (sequence) of parameter
-samples, $\{ \vtheta_i, i = 1, \ldots \}$, such that a particular
-parameter set, $\vtheta$, appears in the sequence with a frequency
-equal to its probability according to $p(\vtheta|d)$.  In this way,
-regions of parameter space where $p(\vtheta|d)$ is large are sampled
-densely while regions where $p(\vtheta|d)$ is small are effectively
-ignored.  A Markov chain has the property that the transition
-probability from one element to the next, $p(\vtheta_i \to
-\vtheta_{i+1})$, depends only on the value of $\vtheta_i$, not on any
-previous values in the chain.
-
-One way to produce a sequence of MCMC samples is via the following
-algorithm, first proposed by \citet{Metropolis1953}:
-\begin{enumerate}
-  \item Begin with the current sample, $\vtheta_i$.
-  \item Propose a new sample, $\vtheta_p$, by drawing randomly from a
-    ``jump proposal distribution'' with probability $Q(\vtheta_i \to
-    \vtheta_p)$.  Note that $Q(\vtheta_i \to \vtheta_p)$ can depend on
-    the current parameters, $\vtheta_i$, and any other ``constant''
-    data, but cannot examine the history of the chain beyond the most
-    recent point.  This is necessary to preserve the Markovian
-    property of the chain.
-  \item Compute the ``acceptance'' probability,
-    \begin{equation}
-      \label{eq:paccept}
-      p_{\textnormal{accept}} \equiv
-      \frac{p(\vtheta_p|d)}{p(\vtheta_i|d)} \frac{Q(\vtheta_p \to
-        \vtheta_i)}{Q(\vtheta_i \to \vtheta_p)}
-    \end{equation}
-  \item With probability $\min(1,p_{\textnormal{accept}})$ ``accept''
-    the propsed $\vtheta_p$, setting $\vtheta_{i+1} = \vtheta_p$;
-    otherwise set $\vtheta_{i+1} = \vtheta_i$.
-\end{enumerate}
-This algorithm is more likely to accept a proposed jump when it
-increases the posterior (the first factor in Equation
-\eqref{eq:paccept}) and when it is to a location in parameter space
-from which it is easy to return (the second factor in Equation
-\eqref{eq:paccept}); the combination of these influences in Equation
-\eqref{eq:paccept} ensures that the equilibrium distribution of the
-chain is $p(\vtheta|d)$.  As $i \to \infty$ the samples $\vtheta_i$
-are distributed according to $p(\vtheta|d)$.
-
-In practice the number of samples required before the chain
-appropriately samples $p(\vtheta|d)$ depends strongly on the jump
-proposal distribution; proposal distributions that often propose jumps
-toward or within regions of large $p(\vtheta|d)$ can be very
-efficient, while poor proposal distributions can require prohibitively
-large numbers of samples before convergence.
-
-There is no foolproof test for the convergence of a chain.  In this
-work, we test the convergence of our chains by comparing the
-statistics calculated from the entire chain to statistics calculated
-from only the first half of the chain; when the chain has converged,
-the two calculations agree.  This is a necessary, but not sufficient,
-condition for convergence.
-
-We begin the chain at an arbitrary point in parameter space; this is
-equivalent to taking a finite section of an infinite chain that begins
-with the chosen point.  Every point in parameter space occurs in an
-infinite chain, and no section of the chain is better than any other,
-so a sufficiently long, but finite, section of the infinite chain
-chosen in this manner can be representative of the statistics of the
-chain as a whole.  However, because consecutive samples in a chain are
-correlated with each other, the beginning of our finite chain has a
-``memory'' of the starting point; we discard enough points at the
-beginning of the finite chain that we can be confident it does not
-retain a memory of the arbitrary starting point.  The points discarded
-in this way are commonly called ``burn-in'' points.
-
-Once we have a chain of samples from $p(\vtheta|d)$---a distribution
-of distributions---the distribution for any quantity of interest can
-be computed.  For example, Figure \ref{fig:dists} shows the median
-values, and the 10\% and 90\% quantiles for the value of the BH mass
-probability distribution at various masses.  (Recall that we are only
-using the 15 low-mass observations in this Section's analysis.)  Such
-a figure can be generated from the MCMC output by computing the value
-of the parameterized BH mass distribution at various masses for each
-parameter sample output by the MCMC, and then plotting the quantiles
-at each mass.
-
-\begin{figure}
-  \begin{center}
-    \plottwo{plots/dist-parametric}{plots/dist-non-parametric}
-  \end{center}
-  \caption{\label{fig:dists} The median values of the black hole mass
-    distribution, $p(M|\theta)$, at various masses implied by the
-    posterior $p(\theta|d)$ for the models discussed in Sections
-    \ref{sec:parametric-models} and
-    \ref{sec:non-parametric-models}.  These distributions use only
-    the 15 low-mass observations in Table \ref{tab:sources} (the
-    complete sample is analyzed in Section \ref{sec:higher-mass}).
-    Error bars span the 10\% to 90\% range.  Note that these
-    ``distributions of distributions'' are not necessarily normalized,
-    and need not be ``shaped'' like the underlying model
-    distributions.}
-\end{figure}
-
-It is also common to look at the one-dimensional distribution for a
-single parameter obtained by integrating over all other dimensions in
-parameter space; this is approximated by a histogram of the MCMC
-sample values for that parameter.  Such a distribution is called the
-``marginalized'' distribution.
+lives in a multi-dimensional space.  Previous works
+\citep{Ozel2010,Bailyn1998} have considered models with only two
+parameters, evaluating $p(\vtheta|d)$ on a grid.  Many of our models
+for the underlying mass distribution have three or more parameters.
+Exploring the entirety of these parameter spaces with a grid rapidly
+becomes prohibitive as the number of parameters increases.  A more
+efficient way to explore the distribution $p(\vtheta | d)$ is to use a
+Markov Chain Monte Carlo (MCMC) method.  MCMC methods produce a chain
+(sequence) of parameter samples, $\{ \vtheta_i \, | \, i = 1, \ldots
+\}$, such that a particular parameter sample, $\vtheta$, appears in
+the sequence with a frequency equal to its posterior probability,
+$p(\vtheta|d)$.  In this way, regions of parameter space where
+$p(\vtheta|d)$ is large are sampled densely while regions where
+$p(\vtheta|d)$ is small are effectively ignored.
+
+Once we have a chain of samples from $p(\vtheta|d)$, the distribution
+for any quantity of interest can be computed by evaluating it on each
+sample in the chain and forming a histogram of these values.  For
+example, to compute the one-dimensional distribution for a single
+parameter obtained by integrating over all other dimensions in
+parameter space, called the ``marginalized'' distribution, one plots
+the histogram of the values of that parameter appearing in the chain.
 
@@
 
-In this subsection, we discuss the various parametric models of the
-underlying black hole mass distribution that we have analyzed using
-the low-mass data from Section \ref{sec:systems}.  (Analysis of these
-models including the higher-mass systems appears in Section
-\ref{sec:higher-mass}.)  Table \ref{tab:low-mass-parametric} gives
-quantiles of the marginalized parameter distributions of the
-parametric models.
+Here we discuss the various parametric models of the underlying black
+hole mass distribution considered in this paper.
+
+\subsubsection{Power-Law Models}
+\label{sec:power-law}
+
+Many astrophysical distributions are power laws.  Let us assume that
+the BH mass distribution is given by
+\begin{equation}
+  \label{eq:power-law-dist}
+  p(M|\vtheta) = p(M|\{\Mmin, \Mmax, \alpha \}) =
+  \begin{cases}
+    A M^\alpha & \Mmin \leq m \leq \Mmax \\
+    0 & \textnormal{otherwise}
+  \end{cases}.
+\end{equation}
+The normalizing constant $A$ is 
+\begin{equation}
+  A = \frac{1+\alpha}{\Mmax^{1+\alpha} - \Mmin^{1+\alpha}}.
+\end{equation}
+We choose uniform priors on $\Mmin$ and $\Mmax \geq \Mmin$ between 0 and
+$40 \Msun$, and uniform priors on the exponent $\alpha$ in a broad
+range between $-15$ and $13$:
+\begin{equation}
+  p(\vtheta) = p(\{\Mmin, \Mmax, \alpha\}) = 
+  \begin{cases}
+    2 \frac{1}{40^2}\frac{1}{28} & 0 \leq \Mmin \leq \Mmax
+    \leq 40, \quad -15 \leq \alpha \leq 13 \\
+    0 & \textnormal{otherwise}
+  \end{cases}.
+\end{equation}
+
+Our MCMC analysis output is a list of $\{\Mmin, \Mmax, \alpha\}$
+values distributed according to the posterior 
+\begin{equation}
+  p(\vtheta|d) = p(\{\Mmin, \Mmax, \alpha\}|d) \propto p(d|\{\Mmin,
+  \Mmax, \alpha\}) p(\{\Mmin, \Mmax, \alpha\}),
+\end{equation}
+with the likelihood $p(d|\{\Mmin, \Mmax, \alpha\})$ defined in
+Equation \eqref{eq:likelihood-def}.
+
+\subsubsection{Decaying Exponential}
+\label{sec:exponential}
+
+\citet{Fryer2001} studied the relation between progenitor and remnant
+mass in simulations of supernova explosions.  Combining this with the
+mass function for supernova progenitors, they suggested that the
+black-hole mass distribution may be well-represented by a decaying
+exponential with a minimum mass: \fixme{Vicky, do you want to say
+  anything more here?}
+\begin{equation}
+  \label{eq:exp-def}
+  p(M|\vtheta) = p(M|\{\Mmin, M_0\}) = 
+  \begin{cases}
+    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
+    \right] & M \geq \Mmin \\
+    0 & \textnormal{otherwise}
+  \end{cases}.
+\end{equation}
+We choose a prior for this model where $\Mmin$ is uniform between 0
+and $40 \Msun$.  For each $\Mmin$, we choose $M_0$ uniformly within a
+range ensuring that $40\Msun$ is at least two scale masses above the
+cutoff: $40\Msun \geq \Mmin + 2 M_0$.  This ensures that the majority
+of the mass probability lies in the range $0 \leq M \leq 40\Msun$.
+The resulting prior is
+\begin{equation}
+  p(\vtheta) = p(\{\Mmin, M_0\}) = 
+  \begin{cases}
+    \frac{4}{40^2} & 0 \leq \Mmin \leq 40, \quad 0 < M_0, \quad \Mmin
+    + 2 M_0 \leq 40, \\
+    0 & \textnormal{otherwise}
+  \end{cases}
+\end{equation}
+
+\subsubsection{Gaussian and Two-Gaussian Models}
+\label{sec:gaussian}
+
+The mass distributions in Figure \ref{fig:low-masses} all peak in a
+relatively narrow range near $\sim 10 \Msun$.  The prototypical
+single-peaked probability distribution is a Gaussian:
+\begin{equation}
+  \label{eq:gaussian-def}
+  p(M|\vtheta) = p(M|\{\mu, \sigma\}) = \frac{1}{\sigma \sqrt{2\pi}}
+  \exp\left[ - \left(\frac{M - \mu}{\sqrt{2} \sigma} \right)^2 \right].
+\end{equation}
+We use a prior on the mean mass, $\mu$, and the standard deviation,
+$\sigma$, that ensures that the majority of the mass distribution lies
+below $40 \Msun$:
+\begin{equation}
+  \label{eq:gaussian-prior-def}
+  p(\{\mu,\sigma\}) = 
+  \begin{cases}
+    \frac{8}{40^2} & 0 \leq \mu \leq 40, \quad \sigma \geq 0, \quad
+    \mu + 2\sigma \leq 40 \\
+    0 & \textnormal{otherwise}
+  \end{cases},
+\end{equation}
+where both $\mu$ and $\sigma$ are measured in solar masses.
+
+Though we do not expect to find a second peak in the low-mass
+distribution, we may find evidence of one when exploring the combined
+low- and higher-mass samples.  To look for a second peak in the
+black-hole mass distribution, we use a two-Gaussian model:
+\begin{multline}
+  \label{eq:two-gaussian-def}
+  p(M|\vtheta) = p(M|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
+  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
+        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
+        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right].
+\end{multline}
+The probability is a linear combination of two Gaussians with weight
+$\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose combined
+conditions on $\mu_i$ and $\sigma_i$ that ensure that most of the mass
+probability lies below $40 \Msun$ with the prior 
+\begin{equation}
+  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
+  \begin{cases}
+    2 p(\{\mu_1, \sigma_1\}) p(\{\mu_2, \sigma_2\}) & \mu_1 \leq
+    \mu_2, \quad 0 \leq \alpha \leq 1 \\
+    0 & \textnormal{otherwise}
+  \end{cases},
+\end{equation}
+where the single-Gaussian prior, $p(\{\mu_i, \sigma_i\})$, is defined
+in Equation \eqref{eq:gaussian-prior-def}.
+
+\subsubsection{Log Normal}
+\label{sec:log-normal}
+
+Many of the mass distributions for the systems in Figure
+\ref{fig:low-masses} rise rapidly to a peak and then fall off more
+slowly in a longer tail toward high masses.  So far, none of the
+parameterized distributions we have discussed have this property.  In
+this section, we consider a log-normal model for the underlying mass
+distribution; the log-normal distribution has a rise to a peak with a
+slower falloff in a long tail.
+
+The log-normal distribution gives $\log M$ a Gaussian distribution
+with mean $\mu$ and standard deviation $\sigma$:
+\begin{equation}
+  \label{eq:log-normal-def}
+  p(M|\vtheta) = p(M|\{\mu, \sigma \}) = \frac{1}{
+    \sqrt{2\pi} M \sigma} \exp\left[ -\frac{\left(\log M - \mu\right)^2}{2 \sigma^2} \right].
+\end{equation}
+The parameters $\mu$ and $\sigma$ are dimensionless; the mean mass
+$\langle M \rangle$ and mass standard deviation $\sigma_M$ are related
+to $\mu$ and $\sigma$ by
+\begin{eqnarray}
+  \label{eq:avg-M}
+  \langle M \rangle & = & \exp\left( \mu + \frac{1}{2} \sigma^2
+  \right) \\
+  \label{eq:sigma-M}
+  \sigma_M & = & \langle M \rangle \sqrt{\exp\left( \sigma^2 \right) - 1}.
+\end{eqnarray}
+For a fair comparison with the other models, we impose a prior that is
+flat in $\langle M \rangle$ and $\sigma_M$.  To ensure that most of
+the probability in this model occurs for masses below $40 \Msun$, we
+require $\langle M \rangle + 2 \sigma_M \leq 40$, resulting in a
+prior
+\begin{equation}
+  p(\vtheta) = p(\{\mu,\sigma\}) = 
+  \begin{cases}
+    \frac{4}{40^2} \left| \frac{\partial \left(\langle M \rangle,
+          \sigma_M \right)}{\partial \left( \mu, \sigma \right)}
+    \right| & \sigma > 0, \quad \langle M \rangle + 2 \sigma_M \leq 40
+    \\
+    0 & \textnormal{otherwise}
+  \end{cases},
+\end{equation}
+where 
+\begin{equation}
+  \left| \frac{\partial \left(\langle M \rangle,
+          \sigma_M \right)}{\partial \left( \mu, \sigma \right)}
+    \right| = \frac{\exp\left( 2 \left( \mu + \sigma^2 \right)\right)
+      \sigma}{\sqrt{\exp\left( \sigma^2 \right) - 1}}
+\end{equation}
+is the determinant of the Jacobian of the map in Equations
+\eqref{eq:avg-M} and \eqref{eq:sigma-M}.
+
+\subsection{Non-Parametric Models for the Black Hole Mass Distribution}
+\label{sec:non-parametric-models}
+
+The previous subsection discussed models for the underlying black hole
+mass distribution that assumed particular parameterized shapes for the
+distribution.  In this subsection, we will discuss models that do not
+assume a priori a shape for the black hole mass distribution.  The
+fundamental non-parametric distribution in this section is a
+histogram with some number of bins, $\Nbin$.  Such a distribution is
+piecewise-constant in $M$.
+
+One choice for representing such a histogram would be to fix the bin
+locations, and allow the heights to vary.  With this approach, one
+should be careful not to ``split'' features of the mass distribution
+across more than one bin in order to avoid diluting the sensitivity to
+such features; similarly, one should avoid including more than ``one''
+feature in each bin.  The locations of the bins, then, are crucial.
+An alternative representation of histogram mass distributions avoids
+this difficulty.
+
+We choose to represent a histogram mass distribution with $\Nbin$ bins
+by allocating a fixed probability, $1/\Nbin$, to each bin.  The lower
+and upper bounds for each bin are allowed to vary; when these are
+close to each other (i.e.\ the bin is narrow), the distribution will
+have a large value, and conversely when the bounds are far from each
+other.  We assume that the non-zero region of the distribution is
+contiguous, so we can represent the boundaries of the bins as a
+non-decreasing array of masses, $w_0 \leq w_1 \leq \ldots \leq
+w_{\Nbin}$, with $w_0$ the minimum and $w_{\Nbin}$ the maximum mass
+for which the distribution has support.  This gives the distribution
+\begin{equation}
+  \label{eq:hist-def}
+  p(M|\theta) = p(M|\{w_0, \ldots, w_{\Nbin}\}) = 
+  \begin{cases}
+    0 & M < w_0 \textnormal{ or } w_{\Nbin} \leq M \\
+    \frac{1}{\Nbin} \frac{1}{w_{i+1} - w_i} & w_i \leq M < w_{i+1}
+  \end{cases}.
+\end{equation}
+
+For priors on the histogram model with $\Nbin$ bins, we assume that
+the bin boundaries are uniformly distributed between 0 and $40 \Msun$
+subject only to the constraint that the boundaries are non-decreasing
+from $w_0$ to $w_{\Nbin}$:
+\begin{equation}
+  p(\{w_0, \ldots, w_{\Nbin}\}) = 
+  \begin{cases}
+    \frac{\left(\Nbin+1\right)!}{40^{\Nbin+1}} & 0 \leq w_0 \leq w_1
+    \leq \ldots \leq w_{\Nbin} \leq 40 \\
+    0 & \textnormal{otherwise}
+  \end{cases}.
+\end{equation}
+
+We consider histograms with up to five bins in this work.  We will see
+that the evidence for the histogram models (see Sections
+\ref{sec:model-selection}, \ref{sec:lower-mass-model-selection}, and
+\ref{sec:high-mass-model-selection}) from both the lower-mass and
+complete datasets is decreasing as the number of bins reaches five,
+indicating that increasing the number of bins beyond five would not
+sufficiently improve the fit to the mass distribution to compensate
+for the extra parameter-space volume implied by the additional
+parameters.
+
+\subsection{Bayesian Model Selection}
+\label{sec:model-selection}
+
+In Sections \ref{sec:parametric-models} and
+\ref{sec:non-parametric-models}, we discussed a series of models for
+the underlying black hole mass distribution.  Our MCMC analysis will
+provide the posterior distribution of the parameters within each
+model, but does not tell us which models are more likely to correspond
+to the actual distribution.  This model selection problem is the topic
+of this section.
+
+Consider a set of models, $\{M_i| i = 1, \ldots\}$, each with
+corresponding parameters $\vtheta_i$.  Re-writing Equation
+\eqref{eq:Bayes-rule} to be explicit about the assumption of a
+particular model, we have
+\begin{equation}
+  p(\vtheta_i | d, M_i) = \frac{p(d|\vtheta_i, M_i) p(\vtheta_i | M_i)}{p(d|M_i)}.
+\end{equation}
+This gives the posterior probability of the parameters $\vtheta_i$ in
+the context of model $M_i$.  But, the model itself can be regarded as
+a discrete parameter in a larger ``super-model'' that encompasses all
+the $M_i$.  The parameters for the super-model are $\{M_i,
+\vtheta_i\}$: a choice of model and the corresponding parameter value
+within that model.  Each point in the super-model parameter space is a
+statement that, e.g., ``the underlying mass distribution is a
+Gaussian, with parameters $\mu$ and $\sigma$,'' or ``the underlying
+mass distribution is a triple-bin histogram with parameters $w_1$,
+$w_2$, $w_3$, and $w_4$,'' or ....  The posterior probability of the
+super-model parameters is given by Bayes' rule:
+\begin{equation}
+  \label{eq:bayes-explicit-model}
+  p(\vtheta_i, M_i|d) = \frac{p(d|\vtheta_i, M_i) p(\vtheta_i |M_i) p(M_i)}{p(d)},
+\end{equation}
+where we have introduced the model prior $p(M_i)$, which represents
+our estimate on the probability that model $M_i$ is correct in the
+absence of the data $d$.  The normalizing evidence is now
+\begin{equation}
+  \label{eq:multi-model-evidence-def}
+  p(d) = \sum_i \int d\vtheta_i\, p(\vtheta_i, M_i|d) = \sum_i
+  p(d|M_i) p(M_i),
+\end{equation}
+writing the single-model evidence from Equation
+\eqref{eq:evidence-def} as $p(d|M_i)$ to be explicit about the
+dependence on the choice of model.
+
+To compare the various models $M_i$, we are interested in the
+marginalized posterior probability of $M_i$:
+\begin{equation}
+  \label{eq:model-posterior-def}
+  p(M_i|d) \equiv \int d\vtheta_i\, p(\vtheta_i, M_i|d).
+\end{equation}
+This is the integral of the posterior over the entire parameter space
+of model $M_i$.  The marginalized posterior probability of model $M_i$
+can be re-written in terms of the single-model evidence, $p(d|M_i)$
+(see Equations \eqref{eq:bayes-explicit-model} and
+\eqref{eq:evidence-def}):
+\begin{equation}
+  \label{eq:model-evidence-def}
+  p(M_i|d) = \int d\vtheta_i\, p(\vtheta_i, M_i|d) = \frac{p(M_i)}{p(d)} \int d\vtheta_i
+  p(d|\vtheta_i,M_i) p(\vtheta_i|M_i) = \frac{p(d|M_i) p(M_i)}{p(d)}.
+\end{equation}
+
+Here and throughout, we assume that any of the models in Section
+\ref{sec:models} are equally likely a priori, so the model priors are
+equal:
+\begin{equation}
+  p(M_i) = \frac{1}{N_{\textnormal{model}}}.
+\end{equation}
+
+A powerful technique%
+\footnote{We also attempted to compute $p(M_i|d)$ using two other
+  methods: the well-known harmonic-mean estimator and the direct
+  integration methods described in \citet{Weinberg2010}.  The harmonic
+  mean is known to be very sensitive to outlying points in the MCMC in
+  general, and we found this to be true in our specific application.
+  The statistical properties of the direct integration algorithm from
+  \citet{Weinberg2010} are less certain, but we found that it was
+  quite noisy in our application compared to the reversible-jump MCMC.
+  Due to the statistical noise in the other two methods, we use the
+  results from our reversible jump MCMC analysis for model
+  selection.} %
+for computing $p(M_i|d)$ is the reversible-jump MCMC
+\citep{Green1995}.  Reversible jump MCMC, discussed in more detail in
+Appendix \ref{sec:reversible-jump-mcmc}, is a standard MCMC analysis
+conducted in the super-model.  The result of a reversible jump MCMC is
+a chain of samples, $\{ M_i, \vtheta_i\, | \, i = 1, \ldots \}$, from the
+super-model parameter space.  The integral in Equation
+\eqref{eq:model-evidence-def} can be estimated by counting the number
+of times that a given model $M_i$ appears in the reversible jump MCMC
+chain:
+\begin{equation}
+  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d) \approx \frac{N_i}{N},
+\end{equation}
+where $N_i$ is the number of MCMC samples that have discrete parameter
+$M_i$, and $N$ is the total number of samples in the MCMC.
+
+Naively implemented reversible jump MCMCs can be very inefficient when
+the posteriors for a model or models are strongly peaked.  In this
+circumstance, a proposed MCMC jump into one of the peaked models is
+unlikely to land on the peak by chance; since it is rare to propose a
+jump into the important regions of parameter space of the peaked model
+in a naive reversible jump MCMC, the output chain must be very long to
+ensure that all models have been compared fairly.  We describe a new
+algorithm in Appendix \ref{sec:reversible-jump-mcmc} that produces
+very efficient jump proposals for a reversible jump MCMC by exploiting
+the information about the model posteriors we have from the
+single-model MCMC samples.  (See also \citet{Farr2010}.)  With this
+algorithm, reasonable chain lengths can fairly compare all the models
+under consideration.  We have used this algorithm to perform 10-way
+reversible jump MCMCs to calculate the relative evidence for both the
+parametric and non-parametric models in this study.  These results
+appear in Section \ref{sec:results}.
+
+\section{Results}
+\label{sec:results}
+
+In this section we discuss the results of our MCMC analysis of the
+posterior distributions of parameters for the models in Sections
+\ref{sec:parametric-models} and \ref{sec:non-parametric-models}.  We
+also discuss model selection results.  The results in Section
+\ref{sec:results-lower-mass} apply to the lower-mass sample of
+systems, while those of Section \ref{sec:higher-mass} apply to the
+complete sample of systems.
+
+\subsection{Lower-Mass Systems}
+\label{sec:results-lower-mass}
+
+Table \ref{tab:low-mass-parametric} gives quantiles of the
+marginalized parameter distributions of the parametric models implied
+by the low-mass data.  Table \ref{tab:low-mass-non-parametric} gives
+the quantiles of the histogram bin boundaries in the non-parametric
+analysis implied by the low-mass data.
 
@@
        \hline \hline
-       Decaying Exponential (Equation \eqref{eq:exp-def}) & $\Mmin$ & 
+       Exponential (Equation \eqref{eq:exp-def}) & $\Mmin$ & 
        5.0185 &  5.4439 &  6.0313 &  6.3785 &  6.5316 \\
@@
     marginalized distribution for each of the parameters in the models
-    discussed in Section \ref{sec:parametric-models}.  We indicate
+    discussed in Section \ref{sec:parametric-models} implied by the low-mass data.  We indicate
     the 5\%, 15\%, 50\% (median), 85\%, and 95\% quantiles.  The
@@
 
-\subsubsection{Power-Law Models}
-\label{sec:power-law}
-
-Many astrophysical distributions are power laws.  Let us assume that
-the BH mass distribution is given by
-\begin{equation}
-  \label{eq:power-law-dist}
-  p(M|\vtheta) = p(M|\{\Mmin, \Mmax, \alpha \}) =
-  \begin{cases}
-    A M^\alpha & \Mmin \leq m \leq \Mmax \\
-    0 & \textnormal{otherwise}
-  \end{cases}.
-\end{equation}
-The normalizing constant $A$ is given by 
-\begin{equation}
-  A = \frac{1+\alpha}{\Mmax^{1+\alpha} - \Mmin^{1+\alpha}}.
-\end{equation}
-We choose uniform priors on $\Mmin$ and $\Mmax \geq \Mmin$ between 0 and
-$40 \Msun$, and uniform priors on the exponent $\alpha$ in a broad
-range between $-15$ and $13$:
-\begin{equation}
-  p(\vtheta) = p(\{\Mmin, \Mmax, \alpha\}) = 
-  \begin{cases}
-    2 \frac{1}{40^2}\frac{1}{28} & 0 \leq \Mmin \leq \Mmax
-    \leq 40, \quad -15 \leq \alpha \leq 13 \\
-    0 & \textnormal{otherwise}
-  \end{cases}.
-\end{equation}
-
-Our MCMC analysis output is a list of $\{\Mmin, \Mmax, \alpha\}$
-values distributed according to the posterior 
-\begin{equation}
-  p(\vtheta|d) = p(\{\Mmin, \Mmax, \alpha\}|d) \propto p(d|\{\Mmin,
-  \Mmax, \alpha\}) p(\{\Mmin, \Mmax, \alpha\}),
-\end{equation}
-with the likelihood $p(d|\{\Mmin, \Mmax, \alpha\})$ defined in
-Equation \eqref{eq:likelihood-def}.  Figure \ref{fig:dists} presents
-the resulting distribution of mass distributions inferred from the
-parameters output by the MCMC.
-
-In Figure \ref{fig:power-law} , we display a histogram of the resulting
-samples in each of the parameters $\Mmin$, $\Mmax$, and $\alpha$
-individually; this represents the one-dimensional ``marginalized''
-distribution
+Recall that each MCMC sample in our analysis gives the parameters for
+a model of the black hole mass distribution.  The chain of samples of
+parameters for a particular model gives us a distribution of black
+hole mass distributions.  Figure \ref{fig:dists} gives a sense of the
+shape and range of the distributions of black hole mass distributions
+that result from our MCMC analysis.  In Figure \ref{fig:dists} we plot
+the median, 10\% and 90\% values of the black hole mass distributions
+that result from the MCMC chains.  Because the choice of parameters
+that gives, for example, the median distribution value at one mass
+need not give the median distribution at another mass, these curves do
+not necessarily look like the underlying model for the mass
+distribution.  For the same reason, they are not necessarily
+normalized.
+
+\begin{figure}
+  \begin{center}
+    \plottwo{plots/dist-parametric}{plots/dist-non-parametric}
+  \end{center}
+  \caption{\label{fig:dists} The median (solid line), 10\% (lower
+    dashed line) and 90\% (upper dashed line) values of the black hole
+    mass distribution, $p(M|\theta)$, at various masses implied by the
+    posterior $p(\theta|d)$ for the models discussed in Sections
+    \ref{sec:parametric-models} and \ref{sec:non-parametric-models}.
+    These distributions use only the 15 low-mass observations in Table
+    \ref{tab:sources} (the complete sample is analyzed in Section
+    \ref{sec:higher-mass}).  Error bars span the 10\% to 90\% range.
+    Note that these ``distributions of distributions'' are not
+    necessarily normalized, and need not be shaped like the underlying
+    model distributions.}
+\end{figure}
+
+\subsubsection{Power Law}
+
+In Figure \ref{fig:power-law} , we display a histogram of the
+resulting samples in each of the parameters $\Mmin$, $\Mmax$, and
+$\alpha$ for the power law model (see Equation
+\eqref{eq:power-law-dist}); this represents the one-dimensional
+``marginalized'' distribution
 \begin{equation}
@@
 \subsubsection{Decaying Exponential}
-\label{sec:exponential}
-
-\citet{Fryer2001} suggest that the black-hole mass distribution may be
-well-represented by a decaying exponential with a minimum mass:
-\fixme{Vicky, do you want to say anything more here?}
-\begin{equation}
-  \label{eq:exp-def}
-  p(M|\vtheta) = p(M|\{\Mmin, M_0\}) = 
-  \begin{cases}
-    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
-    \right] & M \geq \Mmin \\
-    0 & \textnormal{otherwise}
-  \end{cases}.
-\end{equation}
-We choose a prior for this model where $\Mmin$ is uniform between 0
-and $40 \Msun$.  For each $\Mmin$, we choose $M_0$ uniformly within a
-range ensuring that $40\Msun$ is at least two scale masses above the
-cutoff: $40\Msun \geq \Mmin + 2 M_0$.  This ensures that the majority
-of the mass probability lies in the range $0 \leq M \leq 40\Msun$.
-The resulting prior is
-\begin{equation}
-  p(\vtheta) = p(\{\Mmin, M_0\}) = 
-  \begin{cases}
-    \frac{4}{40^2} & 0 \leq \Mmin \leq 40, \quad 0 < M_0, \quad \Mmin
-    + 2 M_0 \leq 40, \\
-    0 & \textnormal{otherwise}
-  \end{cases}
-\end{equation}
 
@@
 distribution for the scale mass of the exponential, $M_0$, and the
-cutoff mass, $\Mmin$.  The median scale mass is $M_0 = 1.55$, and
-$0.78 \leq M_0 \leq 3.25$ with 90\% confidence.  Figure
-\ref{fig:exp-2D} displays the MCMC samples in the $\Mmin$, $M_0$ plane
-for this model.  There is a small correlation between smaller $\Mmin$
-and larger $M_0$, which is driven by the need to widen the
-distribution to encompass the peak of the mass measurements in Figure
-\ref{fig:low-masses} when the minimum mass is smaller.
+cutoff mass, $\Mmin$ (see Equation \ref{eq:exp-def}).  The median
+scale mass is $M_0 = 1.55$, and $0.78 \leq M_0 \leq 3.25$ with 90\%
+confidence.  This model was one of those considered by
+\citet{Ozel2010}, whose results ($M_0 \sim 1.5$ and $\Mmin \sim 6.5$)
+are broadly consistent with ours.  Figure \ref{fig:exp-2D} displays
+the MCMC samples in the $\Mmin$, $M_0$ plane for this model.  There is
+a small correlation between smaller $\Mmin$ and larger $M_0$, which is
+driven by the need to widen the distribution to encompass the peak of
+the mass measurements in Figure \ref{fig:low-masses} when the minimum
+mass is smaller.
 
@@
 
-\subsubsection{Gaussian and Two-Gaussian Models}
-\label{sec:gaussian}
-
-The mass distribution in Figure \ref{fig:low-masses} appears to have a
-single peak.  The prototypical single-peaked probability distribution
-is a Gaussian:
-\begin{equation}
-  \label{eq:gaussian-def}
-  p(M|\vtheta) = p(M|\{\mu, \sigma\}) = \frac{1}{\sigma \sqrt{2\pi}}
-  \exp\left[ - \left(\frac{M - \mu}{\sqrt{2} \sigma} \right)^2 \right].
-\end{equation}
-We use a prior on the mean mass, $\mu$, and the standard deviation,
-$\sigma$, that ensures that the majority of the mass distribution lies
-below $40 \Msun$:
-\begin{equation}
-  \label{eq:gaussian-prior-def}
-  p(\{\mu,\sigma\}) = 
-  \begin{cases}
-    \frac{8}{40^2} & 0 \leq \mu \leq 40, \quad \sigma \geq 0, \quad
-    \mu + 2\sigma \leq 40 \\
-    0 & \textnormal{otherwise}
-  \end{cases},
-\end{equation}
-where both $\mu$ and $\sigma$ are measured in solar masses.  Figure
-\ref{fig:gaussian} shows the resulting marginalized distributions for
-the parameters $\mu$ and $\sigma$.  We constrain the peak of the
-Gaussian between $6.63 \leq \mu \leq 8.08$ with 90\% confidence.
-(Refer to Figure \ref{fig:dists} for the shape of the resulting
-distribution of distributions.)
+\subsubsection{Gaussian}
+
+Figure \ref{fig:gaussian} shows the resulting marginalized
+distributions for the parameters $\mu$ and $\sigma$.  We constrain the
+peak of the Gaussian between $6.63 \leq \mu \leq 8.08$ with 90\%
+confidence.  This model also appeared in \citet{Ozel2010}; they found
+$\mu \sim 7.8$ and $\sigma \sim 1.2$, consistent with our results
+here.
 
@@
 
-Though we do not expect to find a second peak in the low-mass
-distribution, we may find evidence of one when exploring the combined
-low- and higher-mass samples.  To look for a second peak in the
-black-hole mass distribution, we use a two-Gaussian model:
-\begin{multline}
-  \label{eq:two-gaussian-def}
-  p(M|\vtheta) = p(M|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
-  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
-        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
-        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right].
-\end{multline}
-The probability is a linear combination of two Gaussians with weight
-$\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose combined
-conditions on $\mu_i$ and $\sigma_i$ that ensure that most of the mass
-probability lies below $40 \Msun$ with the prior 
-\begin{equation}
-  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
-  \begin{cases}
-    2 p(\{\mu_1, \sigma_1\}) p(\{\mu_2, \sigma_2\}) & \mu_1 \leq
-    \mu_2, \quad 0 \leq \alpha \leq 1 \\
-    0 & \textnormal{otherwise}
-  \end{cases},
-\end{equation}
-where the single-Gaussian prior, $p(\{\mu_i, \sigma_i\})$, is defined
-in Equation \eqref{eq:gaussian-prior-def}.
+\subsubsection{Two Gaussian}
 
@@
 > 0.8$ with 62\% probability, clearly favoring the first Gaussian.
-The distribution for $\mu_1$ and $\sigma_1$ are similar to those of
+The distributions for $\mu_1$ and $\sigma_1$ are similar to those of
 the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
@@
 $p(\textnormal{Gaussian}|d) / p(\textnormal{Two Gaussian}|d) \simeq
-4.7$ (see Section \ref{sec:model-selection} for discussion).
+4.7$ (see Sections \ref{sec:model-selection} and
+\ref{sec:lower-mass-model-selection} for discussion).
 
@@
     model is disfavored relative to the single Gaussian by a factor of
-    $4.7$ on this dataset (see Section \ref{sec:model-selection} for
-    discussion).}
+    $4.7$ on this dataset (see Sections \ref{sec:model-selection} and
+    \ref{sec:lower-mass-model-selection} for discussion).}
 \end{figure}
@@
 \subsubsection{Log Normal}
-\label{sec:log-normal}
-
-Many of the mass distributions for the systems in Figure
-\ref{fig:low-masses} rise rapidly to a peak and then fall off more
-slowly in a longer tail toward high masses.  So far, none of the
-parameterized distributions we have discussed have this property.  In
-this section, we consider a log-normal model for the underlying mass
-distribution; the log-normal distribution has a rise to a peak with a
-slower falloff in a long tail.
-
-The log-normal distribution gives $\log M$ a Gaussian distribution
-with mean $\mu$ and standard deviation $\sigma$:
-\begin{equation}
-  \label{eq:log-normal-def}
-  p(M|\vtheta) = p(M|\{\mu, \sigma \}) = \frac{1}{
-    \sqrt{2\pi} M \sigma} \exp\left[ -\frac{\left(\log M - \mu\right)^2}{2 \sigma^2} \right].
-\end{equation}
-The parameters $\mu$ and $\sigma$ are dimensionless; the mean mass
-$\langle M \rangle$ and mass standard deviation $\sigma_M$ are related
-to $\mu$ and $\sigma$ by
-\begin{eqnarray}
-  \label{eq:avg-M}
-  \langle M \rangle & = & \exp\left( \mu + \frac{1}{2} \sigma^2
-  \right) \\
-  \label{eq:sigma-M}
-  \sigma_M & = & \langle M \rangle \sqrt{\exp\left( \sigma^2 \right) - 1}.
-\end{eqnarray}
-For a fair comparison with the other models, we impose a prior that is
-flat in $\langle M \rangle$ and $\sigma_M$.  To ensure that most of
-the probability in this model occurs for masses below $40 \Msun$, we
-require $\langle M \rangle + 2 \sigma_M \leq 40$, resulting in a
-prior
-\begin{equation}
-  p(\vtheta) = p(\{\mu,\sigma\}) = 
-  \begin{cases}
-    \frac{4}{40^2} \left| \frac{\partial \left(\langle M \rangle,
-          \sigma_M \right)}{\partial \left( \mu, \sigma \right)}
-    \right| & \sigma > 0, \quad \langle M \rangle + 2 \sigma_M \leq 40
-    \\
-    0 & \textnormal{otherwise}
-  \end{cases},
-\end{equation}
-where 
-\begin{equation}
-  \left| \frac{\partial \left(\langle M \rangle,
-          \sigma_M \right)}{\partial \left( \mu, \sigma \right)}
-    \right| = \frac{\exp\left( 2 \left( \mu + \sigma^2 \right)\right)
-      \sigma}{\sqrt{\exp\left( \sigma^2 \right) - 1}}
-\end{equation}
-is the determinant of the Jacobian of the map in Equations
-\eqref{eq:avg-M} and \eqref{eq:sigma-M}.
 
@@
 
-\subsection{Non-Parametric Models for the Black Hole Mass Distribution}
-\label{sec:non-parametric-models}
-
-The previous subsection discussed models for the underlying black hole
-mass distribution that assumed particular parameterized shapes for the
-distribution.  In this subsection, we will discuss models that do not
-assume a priori a shape for the black hole mass distribution.  The
-fundamental non-parametric distribution in this section is a
-histogram with some number of bins, $\Nbin$.  Such a distribution is
-piecewise-constant in $M$.
-
-One choice for representing such a histogram would be to fix the bin
-locations, and allow the heights to vary.  With this approach, one
-should be careful not to ``split'' features of the mass distribution
-across more than one bin in order to avoid diluting the sensitivity to
-such features; similarly, one should avoid including more than ``one''
-feature in each bin.  The locations of the bins, then, are crucial.
-An alternative representation of histogram mass distributions avoids
-this difficulty.
-
-We choose to represent a histogram mass distribution with $\Nbin$ bins
-by allocating a fixed probability, $1/\Nbin$, to each bin.  The lower
-and upper bounds for each bin are allowed to vary; when these are
-close to each other (i.e.\ the bin is narrow), the distribution will
-have a large value, and conversely when the bounds are far from each
-other.  We assume that the non-zero region of the distribution is
-contiguous, so we can represent the boundaries of the bins as a
-non-decreasing array of masses, $w_0 \leq w_1 \leq \ldots \leq
-w_{\Nbin}$, with $w_0$ the minimum and $w_{\Nbin}$ the maximum mass
-for which the distribution has support.  This gives the distribution
-\begin{equation}
-  \label{eq:hist-def}
-  p(M|\theta) = p(M|\{w_0, \ldots, w_{\Nbin}\}) = 
-  \begin{cases}
-    0 & M < w_0 \textnormal{ or } w_{\Nbin} \leq M \\
-    \frac{1}{\Nbin} \frac{1}{w_{i+1} - w_i} & w_i \leq M < w_{i+1}
-  \end{cases}.
-\end{equation}
-
-For priors on the histogram model with $\Nbin$ bins, we assume that
-the bin boundaries are uniformly distributed between 0 and $40 \Msun$
-subject only to the constraint that the boundaries are non-decreasing
-from $w_0$ to $w_{\Nbin}$:
-\begin{equation}
-  p(\{w_0, \ldots, w_{\Nbin}\}) = 
-  \begin{cases}
-    \frac{\left(\Nbin+1\right)!}{40^{\Nbin+1}} & 0 \leq w_0 \leq w_1
-    \leq \ldots \leq w_{\Nbin} \leq 40 \\
-    0 & \textnormal{otherwise}
-  \end{cases}.
-\end{equation}
+\subsubsection{Histogram Models}
 
@@
 histogram is the most probable of the histogram models for the
-low-mass data (see Section \ref{sec:model-selection} for discussion).
-This occurs because the extra fitting power does not sufficiently
-improve the fit to compensate for the vastly larger parameter space of
-the models with more bins.
-
-\section{Model Selection}
-\label{sec:model-selection}
-
-In Section \ref{sec:models}, we discussed a series of models for the
-underlying black hole mass distribution.  For each model, we have
-assumed that the underlying mass distribution corresponds to the
-model, and we ask what distributions are implied by the data for the
-parameters in the model.  We have not yet asked which models are more
-likely to correspond to the actual distribution.  That is the topic of
-this section.
-
-To compare models in the context of a Bayesian analysis, consider a
-set of models, $\{M_i| i = 1, \ldots\}$, each with corresponding
-parameters $\vtheta_i$.  Re-writing Equation \eqref{eq:Bayes-rule} to
-be explicit about the assumption of a particular model, we have
-\begin{equation}
-  p(\vtheta_i | d, M_i) = \frac{p(d|\vtheta_i, M_i) p(\vtheta_i | M_i)}{p(d|M_i)}.
-\end{equation}
-This gives the posterior probability of the parameters $\vtheta_i$ in
-the context of model $M_i$.  But, the model itself can be regarded as
-a discrete parameter in a larger ``super-model'' that encompasses all
-the $M_i$.  The parameters for the super-model are $\{M_i,
-\vtheta_i\}$: a choice of model and the corresponding parameter value
-within that model.  The posterior probability of the super-model
-parameters is given by Bayes' rule:
-\begin{equation}
-  \label{eq:bayes-explicit-model}
-  p(\vtheta_i, M_i|d) = \frac{p(d|\vtheta_i, M_i) p(\vtheta_i |M_i) p(M_i)}{p(d)},
-\end{equation}
-where we have introduced the model prior $p(M_i)$, which represents
-our estimate on the probability that model $M_i$ is correct in the
-absence of the data $d$.  The normalizing evidence is now
-\begin{equation}
-  \label{eq:multi-model-evidence-def}
-  p(d) = \sum_i \int d\vtheta_i\, p(\vtheta_i, M_i|d) = \sum_i
-  p(d|M_i) p(M_i),
-\end{equation}
-writing the single-model evidence from Equation
-\eqref{eq:evidence-def} as $p(d|M_i)$ to be explicit about the
-dependence on the choice of model.
-
-To compare the various models $M_i$, we are interested in the
-marginalized posterior probability of $M_i$:
-\begin{equation}
-  \label{eq:model-posterior-def}
-  p(M_i|d) \equiv \int d\vtheta_i\, p(\vtheta_i, M_i|d).
-\end{equation}
-This is the integral of the posterior over the entire parameter space
-of model $M_i$.  The marginalized posterior probability of model $M_i$
-can be re-written in terms of the single-model evidence, $p(d|M_i)$
-(see Equations \eqref{eq:bayes-explicit-model} and
-\eqref{eq:evidence-def}):
-\begin{equation}
-  \label{eq:model-evidence-def}
-  p(M_i|d) = \int d\vtheta_i\, p(\vtheta_i, M_i|d) = \frac{p(M_i)}{p(d)} \int d\vtheta_i
-  p(d|\vtheta_i,M_i) p(\vtheta_i|M_i) = \frac{p(d|M_i) p(M_i)}{p(d)}.
-\end{equation}
-
-Here and throughout, we assume that any of the models in Section
-\ref{sec:models} are equally likely a priori, so the model priors are
-equal:
-\begin{equation}
-  p(M_i) = \textnormal{const}.
-\end{equation}
-
-\subsection{Reversible-Jump MCMC}
-\label{sec:reversible-jump-mcmc}
-
-A powerful technique for computing $p(M_i|d)$ is the reversible-jump
-MCMC \citep{Green1995}.  Consider the ``super-model'' that encompasses
-all the models under consideration.  The parameter space of the
-super-model consists of a discrete parameter that identifies the
-choice of model, $M_i$, and the continuous parameters appropriate for
-this model, $\vtheta_i$.  We denote a point in the super-model
-parameter space by $\{M_i, \vtheta_i\}$; each such point is a
-statement that, e.g., ``the underlying mass distribution is a
-Gaussian, with parameters $\mu$ and $\sigma$,'' or ``the underlying
-mass distribution is a triple-bin histogram with parameters $w_1$,
-$w_2$, $w_3$, and $w_4$,'' or ....  To compare models, we are
-interested in the quantity (see Equation
-\eqref{eq:model-posterior-def})
-\begin{equation}
-  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d).
-\end{equation}
-If we perform an MCMC in the super-model parameter space, then we
-obtain a chain of samples $\{M_i, \vtheta_i\}$ distributed in
-parameter space with density $p(M_i,\vtheta_i|d) d\vtheta_i$ and we can
-estimate the integral as
-\begin{equation}
-  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d) \approx \frac{N_i}{N},
-\end{equation}
-where $N_i$ is the number of MCMC samples that have discrete parameter
-$M_i$, and $N$ is the total number of samples in the MCMC.  In other
-words, the fraction of samples lying in the parameter space of model
-$M_i$ gives the probability of that model relative to the other models
-under consideration.
-
-To perform the MCMC in the super-model parameter space, we must
-propose jumps not only between points in a particular model's
-parameter space (as we have been doing up to now), but also between
-the parameter spaces of different models.  For this MCMC to be
-efficient, proposed jumps into a model from another should favor
-regions with large posterior; when the posterior is highly-peaked in a
-small region of parameter space, proposed jumps outside this region
-are unlikely to be accepted, and the reversible-jump MCMC samples will
-require a very long chain to properly sample the ``super-model''
-posterior.
-
-We can exploit the information we have from the single-model MCMCs
-described in Section \ref{sec:models} to generate efficient jump
-proposal distributions for our reversible jump MCMC.  We would like to
-propose jumps that roughly follow the distribution of samples in the
-single-model MCMCs.  We can do this by assigning a neighborhood to
-each point in the sample using an algorithm we will describe in the
-following paragraphs; the neighborhoods are non-overlapping,
-completely cover the region of parameter space with prior support, and
-contain only one point from the MCMC samples.  To propose a jump into
-model $M_i$, we choose a point uniformly from that single-model MCMC
-and then propose a jump drawn uniformly from that point's
-neighborhood.  This is equivalent to drawing from a piecewise-constant
-approximation to the single-model posterior, where each neighborhood
-contributes a constant fraction, $1/N_i$, to the cumulative jump
-probability.  In regions of high density the neighborhoods are
-smaller, and the jump probability density is correspondingly higher.
-Because the neighborhoods cover the entire region of prior support, it
-is possible for the proposal to propose any point in parameter space
-with prior support (though points in regions of low single-model
-posterior are of course unlikely to be proposed).
-
-To assign a neighborhood to each point in a single-model MCMC we use a
-data structure called a kD-tree.  A kD-tree is a binary
-space-partitioning tree.  To construct a kD-tree, we begin with the
-set of points from a single-model MCMC and a box in parameter space
-bounding the region of prior support (which must necessarily enclose
-all the points).  The construction proceeds recursively: we choose a
-dimension%
-\footnote{Our algorithm chooses the dimension along which the
-  numerical extent of the points is largest.  Other choices are
-  possible; some algorithms cycle through the dimensions in order,
-  while others choose a random dimension for each subdivision.  Our
-  goal by picking the longest dimension is to produce neighborhoods
-  that are ``square,'' at least in the chosen parametrization.} %
-along which to divide the points, find the median point along that
-dimension and its nearest neighbor, and divide the box at the midpoint
-between these two points, producing two sub-boxes.  We then partition
-the points into those to the left (i.e.\ smaller coordinate along the
-given dimension) and right of the dividing line, and repeat this
-procedure for each subset and the corresponding bounding box, until we
-have only one point in each box.  An example of the neighborhoods that
-result from a two-dimensional kD-tree constructed around a Gaussian
-point distribution appears in Figure \ref{fig:kD-tree}.
-
-\begin{figure}
-  \begin{center}
-    \plotone{plots/kdtree}
-  \end{center}
-  \caption{\label{fig:kD-tree} The neighborhoods constructed from a
-    two-dimensional kD-tree built from a sample of points with a
-    Gaussian density distribution.  Each line on the figure
-    corresponds to a sub-dividing box boundary drawn between the
-    median of a subset of the sample points and its nearest
-    neighbor. The peak of the Gaussian lies in the center of the
-    figure; here the point density is highest and the neighborhoods
-    are smallest.  Near the edges the density is lower, and the
-    neighborhoods correspondingly larger.  The tree adapts itself to
-    the local density of points.  If these were single-model MCMC
-    samples, the corresponding jump proposal would first select one of
-    the boxes uniformly at random, and then choose a point uniformly
-    within the box to propose.  Since there are many more boxes near
-    the center (each box corresponds to one point), and these boxes
-    are smaller, the proposal will tend to concentrate its points
-    there, approximately tracking the distribution of single-model
-    MCMC samples.}
-\end{figure}
-
-Construction of a kD-tree is an $\order{N\log N}$ operation, where $N$
-is the number of points in the tree.  Median finding is $\order{n}$,
-where $n$ is the number of points from which the median is to be
-obtained.  At level $i$ in the tree, there are $2^i$ subsets of
-points, each of length $\order{N/2^i}$, so the total cost of the $2^i$
-median calculations is $\order{N}$ at each level.  There are
-$\order{\log N}$ levels in the tree, yielding a total construction
-cost for the tree of $\order{N \log N}$.
-
-To find the neighborhood of a point using the tree, we begin at the
-root of the tree, and examine the two sub-boxes at the next level
-down.  The point will be in one of them; following that branch, we
-have again two sub-boxes, one of which contains the point; following
-that branch....  Eventually, the search terminates at a leaf of the
-tree, containing the point in question.  The box at the leaf defines
-the neighborhood of the point in the jump proposal algorithm described
-above.  The total cost for this operation is proportional to the
-number of levels in the tree, which is $\order{\log N}$.
-
+low-mass data (see Section \ref{sec:lower-mass-model-selection} for
+discussion).  This occurs because the extra fitting power does not
+sufficiently improve the fit to compensate for the vastly larger
+parameter space of the models with more bins.
+
+\subsubsection{Lower-Mass Model Selection}
+\label{sec:lower-mass-model-selection}
 We have performed a suite of 500 independent reversible-jump MCMCs
@@
 Section \ref{sec:models} using the single-model MCMC samples to
-construct an efficient jump proposal for each model as described
-above.  The numbers of counts in each model are consistent across the
-MCMCs in the suite; Figure \ref{fig:rj} displays the average
-probability for each model across the suite, along with the 1-$\sigma$
-errors on the average inferred from the standard deviation of the
-model counts across the suite.  Table \ref{tab:rj} gives the numerical
-values of the average probability for each model across the suite of
-MCMCs.
+construct an efficient jump proposal for each model as described above
+(see Appendix \ref{sec:reversible-jump-mcmc}).  The numbers of counts in
+each model are consistent across the MCMCs in the suite; Figure
+\ref{fig:rj} displays the average probability for each model across
+the suite, along with the 1-$\sigma$ errors on the average inferred
+from the standard deviation of the model counts across the suite.
+Table \ref{tab:rj} gives the numerical values of the average
+probability for each model across the suite of MCMCs.
 
@@
 \ref{sec:power-law}, followed by the Gaussian model from Section
-\ref{sec:gaussian}.  This confirms the results from the harmonic mean
-evidence estimator in the last subsection.  (Though note that the
-harmonic mean ranks the one-bin histogram significantly higher than
-the reversible jump calculation; given the large uncertainty in the
-harmonic mean estimator, it is likely that the true evidence for the
-one-bin histogram model is closer to the value computed in this
-section.)  Interestingly, the theoretical curve from \citet{Fryer2001}
-(the exponential model of Section \ref{sec:exponential}) places fourth
-in the ranking of evidence.
+\ref{sec:gaussian}.  Interestingly, the theoretical curve from
+\citet{Fryer2001} (the exponential model of Section
+\ref{sec:exponential}) places fourth in the ranking of evidence.
 
@@
   \caption{\label{tab:rj} Relative probabilities of the various models
-    from Section \ref{sec:models} computed from reversible-jump
-    MCMC samples with the efficient jump proposal algorithm in Section \ref{sec:reversible-jump-mcmc}.}
+    from Section \ref{sec:models} implied by the lower-mass data.  These probabilities have been 
+    computed from reversible-jump
+    MCMC samples using the efficient jump proposal algorithm in Appendix \ref{sec:reversible-jump-mcmc}.}
 \end{table}
@@
 
-\section{Including High-Mass, Wind-Fed Systems}
+\subsection{Complete Sample}
 \label{sec:higher-mass}
 
-This section repeats the analysis of Sections \ref{sec:models} and
-\ref{sec:model-selection} but including the higher-mass, wind-fed
-systems from Table \ref{tab:sources} (see also Figure
-\ref{fig:high-masses}) in the sample.  Figure
-\ref{fig:high-mass-dists} displays bounds on the value of the
-underlying mass distribution for the various models in Section
+This section repeats the analysis of the models from Section
+\ref{sec:models}, but including the higher-mass, wind-fed systems from
+Table \ref{tab:sources} (see also Figure \ref{fig:high-masses}) in the
+sample.  Figure \ref{fig:high-mass-dists} displays bounds on the value
+of the underlying mass distribution for the various models in Section
 \ref{sec:models} applied to this data set; compare to Figure
@@
   \end{center}
-  \caption{\label{fig:high-mass-dists} The median values of the black
-    hole mass distribution, $p(M|\theta)$, at various masses implied
-    by the posterior $p(\theta|d)$ for the models discussed in
+  \caption{\label{fig:high-mass-dists} The median (solid line), 10\%
+    (lower dashed line), and 90\% (upper dashed line) values of the
+    black hole mass distribution, $p(M|\theta)$, at various masses
+    implied by the posterior $p(\theta|d)$ for the models discussed in
     Sections \ref{sec:parametric-models} and
@@
 
-Figures \ref{fig:power-law-high} presents the marginalized
+\subsubsection{Power Law}
+
+Figure \ref{fig:power-law-high} presents the marginalized
 distribution for the three power-law parameters $\Mmin$, $\Mmax$, and
@@
 
+\subsubsection{Decaying Exponential}
+
 Figure \ref{fig:exp-cutoff-high} displays the marginalized
@@
 
+\subsubsection{Gaussian}
+
 Figure \ref{fig:gaussian-high} displays the marginalized distributions
@@
 \end{figure}
+
+\subsubsection{Two Gaussian}
 
@@
 
+\subsubsection{Log Normal}
+
 The marginalized distributions for the log-normal parameters (Section
@@
       \hline \hline
-      Decaying Exponential (Equation \eqref{eq:exp-def}) & $\Mmin$ & 
+      Exponential (Equation \eqref{eq:exp-def})& $\Mmin$ & 
       4.0865 & 4.60236 & 5.32683 & 5.94097 & 6.22952 \\
@@
     higher-mass samples are included in the analysis (compare to Table
-    \ref{tab:low-mass-parametric}).  We indicate the 5\%, 15\%, 50\% (median), 85\%, and 95\% quantiles.}
+    \ref{tab:low-mass-parametric}).  We indicate the 5\%, 15\%, 50\% 
+    (median), 85\%, and 95\% quantiles.}
 \end{table}
 
-The non-parametric (histogram; see Section \ref{sec:non-parametric-models}) models
-also show evidence of a long tail from the inclusion of the
-higher-mass samples.  Table \ref{tab:high-mass-non-parametric}
-displays confidence limits on the histogram parameters for the
-analysis including the higher-mass systems; compare to Table
-\ref{tab:low-mass-non-parametric}.
+\subsubsection{Histogram Models}
+
+The non-parametric (histogram; see Section
+\ref{sec:non-parametric-models}) models also show evidence of a long
+tail from the inclusion of the higher-mass samples.  Table
+\ref{tab:high-mass-non-parametric} displays confidence limits on the
+histogram parameters for the analysis including the higher-mass
+systems; compare to Table \ref{tab:low-mass-non-parametric}.
 
@@
 
-\subsection{Model Selection}
+\subsubsection{High-Mass Model Selection}
+\label{sec:high-mass-model-selection}
 
 Repeating the model selection analysis discussed in Section
-\ref{sec:model-selection} for the sample including the higher-mass
-systems, we find that the model probabilities have changed with the
-inclusion of the extra five systems.  This is conclusive evidence (as
-if more were needed) that the higher-mass, wind-fed systems come from
-a different underlying population than the lower-mass, Roche-lobe
-overflow systems.  As before, we assume for this analysis that the
-model priors are equal.
+\ref{sec:lower-mass-model-selection} for the sample including the
+higher-mass systems, we find that the model probabilities have changed
+with the inclusion of the extra five systems.  As before, we assume
+for this analysis that the model priors are equal.
 
@@
 The ranking of models differs significantly from the lower-mass
-samples.  The model selection provides more strong evidence that the
-wind-fed and Roche-lobe systems come from different underlying
-populations.  The improvement of the exponential model relative to the
+samples.  The improvement of the exponential model relative to the
 lower-mass analysis is encouraging for theoretical calculations that
@@
     black hole mass calculated from the MCMC samples for the models in
-    Section \ref{sec:models}.  The minimum black hole mass is defined
-    as the 1\% mass quantile.  For the most favored models, the
-    power-law and Gaussian, the 90\% confidence limit on the minimum
-    black hole mass is 4.3 $\Msun$ and 2.9 $\Msun$, respectively.}
-\end{figure}
-
-We find that the best-fit model (the power-law) has a 90\% confidence
-limit on the minimum black hole mass of 4.3 $\Msun$.  This is
-significantly above the maximum theoretically-allowed neutron star
-mass, $\sim 3 \Msun$ \citep{Kalogera1996}\fixme{Vicky, do you have a
-  more recent reference than this?}, so we conclude that the
-lower-mass systems show strong evidence of a mass gap.
+    Section \ref{sec:models} applied to the lower-mass systems.  The
+    minimum black hole mass is defined as the 1\% mass quantile.  For
+    the most favored models, the power-law and Gaussian, the 90\%
+    confidence limit on the minimum black hole mass is 4.3 $\Msun$ and
+    2.9 $\Msun$, respectively.}
+\end{figure}
+
+We find that the best-fit model for the lower-mass systems (the
+power-law) has a 90\% confidence limit on the minimum black hole mass
+of 4.3 $\Msun$.  This is significantly above the maximum
+theoretically-allowed neutron star mass, $\sim 3 \Msun$
+\citep{Kalogera1996}\fixme{Vicky, do you have a more recent reference
+  than this?}, so we conclude that the lower-mass systems show strong
+evidence of a mass gap.
 
@@
     The minimum black hole mass is defined as the 1\% mass quantile.
-    For the most favored models, the power-law, exponential and
-    two-Gaussian, the 90\% confidence limit on the minimum black hole
-    mass is 5.2 $\Msun$, 4.5 $\Msun$, and 2.2 $\Msun$, respectively.}
+    For the most favored model, the exponential, the 90\% confidence
+    limit on the minimum black hole mass is 4.5 $\Msun$.}
 \end{figure}
@@
 \section{Conclusion}
+
+We have presented a Bayesian analysis of the mass distribution of
+stellar-mass black holes in X-ray binary systems.  We considered
+separately a sample of 15 lower-mass, Roche lobe-filling systems and a
+sample of 20 systems containing the 15 lower-mass systems and five
+higher-mass, wind-fed X-ray binaries.  We used MCMC methods to sample
+the posterior distributions of the parameters for five parametric
+models and five non-parametric (histogram) models for the mass
+distribution that are implied by these data.  For both sets of
+systems, we used reversible jump MCMCs (exploiting a new algorithm for
+efficient jump proposals in such calculations) to perform model
+selection on the suite of models.
+
+For the lower-mass systems, we found the limits on model parameters in
+Tables \ref{tab:low-mass-parametric} and
+\ref{tab:low-mass-non-parametric}.  The relative model probabilites
+from the model selection are given in Table \ref{tab:rj}.  The most
+favored model for the lower-mass systems is a power law.  The
+equivalent limits on the model parameters for the combined systems are
+given in Tables \ref{tab:high-mass-parametric} and
+\ref{tab:high-mass-non-parametric}.  Unlike the lower-mass systems,
+the most favored model for the combined sample is the exponential
+model.
+
+We found strong evidence for a mass gap between the most massive
+neutron stars, $\sim 3 \Msun$ \citep{Kalogera1996}, and the least
+massive black holes.  For the lower-mass systems, the most favored,
+power law model gives a black hole mass distribution whose 1\%
+quantile lies above $4.3 \Msun$ with 90\% confidence.  For the
+complete sample of systems, the most favored, exponential model gives
+a black hole mass distribution whose 1\% quantile lise above $4.5
+\Msun$ with 90\% confidence.  The existence of a mass gap was also
+observed in \citet{Bailyn1998} (which used different models in a study
+of 7 systems) and \citet{Ozel2010} (which did not contain a power law
+model, and applied the exponential model to a sample of 16 lower-mass
+systems, where it is strongly disfavored).  The existence of a mass
+gap poses a challenge to current theoretical models of compact object
+formation \citep{Fryer2001}, which predict a continuous mass
+distribution from neutron star masses to stellar mass black hole
+masses.
 
@@
 
+\appendix
+
+\section{Markov Chain Monte Carlo}
+
+MCMC methods produce a Markov chain (or sequence) of parameter
+samples, $\{ \vtheta_i \, | \, i = 1, \ldots \}$, such that a particular
+parameter set, $\vtheta$, appears in the sequence with a frequency
+equal to its probability according to a posterior, $p(\vtheta|d)$.  A
+Markov chain has the property that the transition probability from one
+element to the next, $p(\vtheta_i \to \vtheta_{i+1})$, depends only on
+the value of $\vtheta_i$, not on any previous values in the chain.
+
+One way to produce a sequence of MCMC samples is via the following
+algorithm, first proposed by \citet{Metropolis1953} and used widely in
+the physical sciences thereafter:
+\begin{enumerate}
+  \item Begin with the current sample, $\vtheta_i$.
+  \item Propose a new sample, $\vtheta_p$, by drawing randomly from a
+    ``jump proposal distribution'' with probability $Q(\vtheta_i \to
+    \vtheta_p)$.  Note that $Q(\vtheta_i \to \vtheta_p)$ can depend on
+    the current parameters, $\vtheta_i$, and any other ``constant''
+    data, but cannot examine the history of the chain beyond the most
+    recent point.  This is necessary to preserve the Markovian
+    property of the chain.
+  \item Compute the ``acceptance'' probability,
+    \begin{equation}
+      \label{eq:paccept}
+      p_{\textnormal{accept}} \equiv
+      \frac{p(\vtheta_p|d)}{p(\vtheta_i|d)} \frac{Q(\vtheta_p \to
+        \vtheta_i)}{Q(\vtheta_i \to \vtheta_p)}
+    \end{equation}
+  \item With probability $\min(1,p_{\textnormal{accept}})$ ``accept''
+    the propsed $\vtheta_p$, setting $\vtheta_{i+1} = \vtheta_p$;
+    otherwise set $\vtheta_{i+1} = \vtheta_i$.
+\end{enumerate}
+This algorithm is more likely to accept a proposed jump when it
+increases the posterior (the first factor in Equation
+\eqref{eq:paccept}) and when it is to a location in parameter space
+from which it is easy to return (the second factor in Equation
+\eqref{eq:paccept}); the combination of these influences in Equation
+\eqref{eq:paccept} ensures that the equilibrium distribution of the
+chain is $p(\vtheta|d)$.  As $i \to \infty$ the samples $\vtheta_i$
+are distributed according to $p(\vtheta|d)$.
+
+In practice the number of samples required before the chain
+appropriately samples $p(\vtheta|d)$ depends strongly on the jump
+proposal distribution; proposal distributions that often propose jumps
+toward or within regions of large $p(\vtheta|d)$ can be very
+efficient, while poor proposal distributions can require prohibitively
+large numbers of samples before convergence.
+
+There is no foolproof test for the convergence of a chain.  In this
+work, we test the convergence of our chains by comparing the
+statistics calculated from the entire chain to statistics calculated
+from only the first half of the chain; when the chain has converged,
+the two calculations agree.  This is a necessary, but not sufficient,
+condition for convergence.
+
+We begin the chain at an arbitrary point in parameter space; this is
+equivalent to taking a finite section of an infinite chain that begins
+with the chosen point.  Every point in parameter space occurs in an
+infinite chain, and no section of the chain is better than any other,
+so a sufficiently long, but finite, section of the infinite chain
+chosen in this manner can be representative of the statistics of the
+chain as a whole.  However, because consecutive samples in a chain are
+correlated with each other, the beginning of our finite chain has a
+``memory'' of the starting point; we discard enough points at the
+beginning of the finite chain that we can be confident it does not
+retain a memory of the arbitrary starting point.  The points discarded
+in this way are commonly called ``burn-in'' points.
+
+\section{Reversible-Jump MCMC}
+\label{sec:reversible-jump-mcmc}
+
+Consider the problem of model selection among a set of models, and the
+``super-model'' that encompasses all the models under consideration.
+The parameter space of the super-model consists of a discrete
+parameter that identifies the choice of model, $M_i$, and the
+continuous parameters appropriate for this model, $\vtheta_i$.  We
+denote a point in the super-model parameter space by $\{M_i,
+\vtheta_i\}$; each such point is a statement that, e.g., ``the
+underlying mass distribution for black holes in the galaxy is a
+Gaussian, with parameters $\mu$ and $\sigma$,'' or ``the underlying
+mass distribution for black holes in the galaxy is a triple-bin
+histogram with parameters $w_1$, $w_2$, $w_3$, and $w_4$,'' or ....
+To compare models, we are interested in the quantity (see Equation
+\eqref{eq:model-posterior-def})
+\begin{equation}
+  p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d).
+\end{equation}
+If we perform an MCMC in the super-model parameter space, then we
+obtain a chain of samples $\{M_i, \vtheta_i \, | \, i = 1, \ldots\}$
+distributed in parameter space with density $p(M_i,\vtheta_i|d)
+d\vtheta_i$ and we can estimate the integral as
+\begin{equation}
+    p(M_i|d) = \int d\vtheta_i p(M_i, \vtheta_i|d) \approx \frac{N_i}{N},
+\end{equation}
+where $N_i$ is the number of samples in the chain lying in the
+parameter space of model $M_i$ and $N$ is the total number of samples
+in the chain.  The fraction of samples lying in the parameter space of
+model $M_i$ gives the probability of that model relative to the other
+models under consideration.
+
+To perform the MCMC in the super-model parameter space, we must
+propose jumps not only between points in a particular model's
+parameter space, but also between the parameter spaces of different
+models.  For this MCMC to be efficient, proposed jumps into a model
+from another should favor regions with large posterior; when the
+posterior is highly-peaked in a small region of parameter space,
+proposed jumps outside this region are unlikely to be accepted, and
+the reversible-jump MCMC samples will require a very long chain to
+properly sample the ``super-model'' posterior.
+
+We can exploit the information we have from single-model MCMCs to
+generate efficient jump proposal distributions for our reversible jump
+MCMC.  We would like to propose jumps that roughly follow the
+distribution of samples in the single-model MCMCs.  We can do this by
+assigning a neighborhood to each point in the sample using an
+algorithm we will describe in the following paragraphs; the
+neighborhoods are non-overlapping, completely cover the region of
+parameter space with prior support, and contain only one point from
+the MCMC samples.  To propose a jump into model $M_i$, we choose a
+point uniformly from that single-model MCMC and then propose a jump
+drawn uniformly from that point's neighborhood.  This is equivalent to
+drawing from a piecewise-constant approximation to the single-model
+posterior, where each neighborhood contributes a constant fraction,
+$1/N_i$, to the cumulative jump probability.  In regions of high
+density the neighborhoods are smaller, and the jump probability
+density is correspondingly higher.  Because the neighborhoods cover
+the entire region of prior support, it is possible for the proposal to
+propose any point in parameter space with prior support (though points
+in regions of low single-model posterior are of course unlikely to be
+proposed).
+
+To assign a neighborhood to each point in a single-model MCMC we use a
+data structure called a kD-tree.  A kD-tree is a binary
+space-partitioning tree.  To construct a kD-tree, we begin with the
+set of points from a single-model MCMC and a box in parameter space
+bounding the region of prior support (which must necessarily enclose
+all the points).  The construction proceeds recursively: we choose a
+dimension%
+\footnote{Our algorithm chooses the dimension along which the
+  numerical extent of the points is largest.  Other choices are
+  possible; some algorithms cycle through the dimensions in order,
+  while others choose a random dimension for each subdivision.  Our
+  goal by picking the longest dimension is to produce neighborhoods
+  that are ``square,'' at least in the chosen parametrization.} %
+along which to divide the points, find the median point along that
+dimension and its nearest neighbor, and divide the box at the midpoint
+between these two points, producing two sub-boxes.  We then partition
+the points into those to the left (i.e.\ smaller coordinate along the
+given dimension) and right of the dividing line, and repeat this
+procedure for each subset and the corresponding bounding box, until we
+have only one point in each box.  An example of the neighborhoods that
+result from a two-dimensional kD-tree constructed around a Gaussian
+point distribution appears in Figure \ref{fig:kD-tree}.
+
+\begin{figure}
+  \begin{center}
+    \plotone{plots/kdtree}
+  \end{center}
+  \caption{\label{fig:kD-tree} The neighborhoods constructed from a
+    two-dimensional kD-tree built from a sample of points with a
+    Gaussian density distribution.  Each line on the figure
+    corresponds to a sub-dividing box boundary drawn between the
+    median of a subset of the sample points and its nearest
+    neighbor. The peak of the Gaussian lies in the center of the
+    figure; here the point density is highest and the neighborhoods
+    are smallest.  Near the edges the density is lower, and the
+    neighborhoods correspondingly larger.  The tree adapts itself to
+    the local density of points.  If these were single-model MCMC
+    samples, the corresponding jump proposal would first select one of
+    the boxes uniformly at random, and then choose a point uniformly
+    within the box to propose.  Since there are many more boxes near
+    the center (each box corresponds to one point), and these boxes
+    are smaller, the proposal will tend to concentrate its points
+    there, approximately tracking the distribution of single-model
+    MCMC samples.}
+\end{figure}
+
+Construction of a kD-tree is an $\order{N\log N}$ operation, where $N$
+is the number of points in the tree.  Median finding is $\order{n}$,
+where $n$ is the number of points from which the median is to be
+obtained.  At level $i$ in the tree, there are $2^i$ subsets of
+points, each of length $\order{N/2^i}$, so the total cost of the $2^i$
+median calculations is $\order{N}$ at each level.  There are
+$\order{\log N}$ levels in the tree, yielding a total construction
+cost for the tree of $\order{N \log N}$.
+
+To find the neighborhood of a point using the tree, we begin at the
+root of the tree, and examine the two sub-boxes at the next level
+down.  The point will be in one of them; following that branch, we
+have again two sub-boxes, one of which contains the point; following
+that branch....  Eventually, the search terminates at a leaf of the
+tree, containing the point in question.  The box at the leaf defines
+the neighborhood of the point in the jump proposal algorithm described
+above.  The total cost for this operation is proportional to the
+number of levels in the tree, which is $\order{\log N}$.
+
 \bibliography{paper}

### 2010-10-26 paper.tex
Plot edits (partial) for Vicky.

@@
   \begin{center}
-    \plottwo{plots/dist-parametric}{plots/dist-non-parametric}
+    \plotone{plots/dist}
   \end{center}
@@
     \ref{tab:sources} (the complete sample is analyzed in Section
-    \ref{sec:higher-mass}).  Error bars span the 10\% to 90\% range.
-    Note that these ``distributions of distributions'' are not
-    necessarily normalized, and need not be shaped like the underlying
-    model distributions.}
+    \ref{sec:higher-mass}).  Note that these ``distributions of
+    distributions'' are not necessarily normalized, and need not be
+    shaped like the underlying model distributions.}
 \end{figure}
@@
   \end{center}
-  \caption{\label{fig:power-law} A histogram of the marginalized
+  \caption{\label{fig:power-law} Histograms of the marginalized
     distribution for the three parameters $\Mmin$, $\Mmax$, and
@@
   \caption{\label{fig:exp-marginal} The distribution of scale masses,
-    $M_0$, and minimum masses, $\Mmin$, both measured in units of a
-    solar mass for the exponential underlying mass distribution
-    defined in Equation \eqref{eq:exp-def}.  The median scale mass is
-    $M_0 = 1.55$, and $0.78 \leq M_0 \leq 3.25$ with 90\%
-    confidence.}
+    $M_0$ (dashed histogram), and minimum masses, $\Mmin$ (solid
+    histogram), both measured in units of a solar mass for the
+    exponential underlying mass distribution defined in Equation
+    \eqref{eq:exp-def}.  The median scale mass is $M_0 = 1.55$, and
+    $0.78 \leq M_0 \leq 3.25$ with 90\% confidence.}
 \end{figure}
@@
   \caption{\label{fig:gaussian} Marginalized posterior distributions
-    for the mean $\mu$ and standard deviation $\sigma$ (both in solar
-    masses) for the Gaussian underlying mass distribution defined in
-    Equation \eqref{eq:gaussian-def}.  The peak of the Gaussian,
-    $\mu$, is constrained in $6.63 \leq \mu \leq 8.08$ with 90\%
-    confidence.}
+    for the mean, $\mu$ (solid histogram), and standard deviation,
+    $\sigma$ (dashed histogram), both in solar masses for the Gaussian
+    underlying mass distribution defined in Equation
+    \eqref{eq:gaussian-def}.  The peak of the Gaussian, $\mu$, is
+    constrained in $6.63 \leq \mu \leq 8.08$ with 90\% confidence.}
 \end{figure}
@@
   \caption{\label{fig:two-gaussian} The marginal distributions for the
-    five parameters of the two-Gaussian model.  We have $\alpha > 0.8$
-    with 62\% probability, favoring the first of the two Gaussians.
-    The distributions for $\mu_1$ and $\sigma_1$ are similar to those
-    of the single Gaussian model displayed in Figure
+    five parameters of the two-Gaussian model.  The top panel is
+    $\mu_1$ (solid histogram) and $\sigma_1$ (dashed histogram), the
+    middle panel is $\mu_2$ (solid histogram) and $\sigma_2$ (dashed
+    histogram), and the bottom panel is $\alpha$. We have $\alpha >
+    0.8$ with 62\% probability, favoring the first of the two
+    Gaussians.  The distributions for $\mu_1$ and $\sigma_1$ are
+    similar to those of the single Gaussian model displayed in Figure
     \ref{fig:gaussian}; the second Gaussian's parameter distributions
@@
   \caption{\label{fig:log-normal} Marginalized distributions of the
-    mean mass, $\langle M \rangle$, and standard deviation of the
-    mass, $\sigma_M$, for the log-normal model in Section
-    \ref{sec:log-normal}.  The distributions are similar to the
-    distributions of $\mu$ and $\sigma$ in the Gaussian model of
-    Section \ref{sec:gaussian}.}
+    mean mass, $\langle M \rangle$ (solid histogram), and standard
+    deviation of the mass, $\sigma_M$ (dashed histogram), for the
+    log-normal model in Section \ref{sec:log-normal}.  The
+    distributions are similar to the distributions of $\mu$ and
+    $\sigma$ in the Gaussian model of Section \ref{sec:gaussian}.}
 \end{figure}
@@
   \begin{center}
-    \plottwo{plots/dist-parametric-high}{plots/dist-non-parametric-high-mass}
+    \plotone{plots/dist-high}
   \end{center}
@@
     complete sample of 20 observations in Table \ref{tab:sources},
-    including the higher-mass, wind-fed systems.  Error bars span the
-    10\% to 90\% range.  Note that these ``distributions of
-    distributions'' are not necessarily normalized, and need not be
-    ``shaped'' like the underlying model distributions.  Compare to
-    Figure \ref{fig:dists}, which includes only the low-mass systems
-    in the analysis.  Including the higher-mass systems tends to widen
-    the distribution toward the high-mass end and, in models that
-    allow it, produce a second, higher-mass peak in addition to the
-    one in Figure \ref{fig:dists}. }
+    including the higher-mass, wind-fed systems.  Note that these
+    ``distributions of distributions'' are not necessarily normalized,
+    and need not be ``shaped'' like the underlying model
+    distributions.  Compare to Figure \ref{fig:dists}, which includes
+    only the low-mass systems in the analysis.  Including the
+    higher-mass systems tends to widen the distribution toward the
+    high-mass end and, in models that allow it, produce a second,
+    higher-mass peak in addition to the one in Figure
+    \ref{fig:dists}. }
 \end{figure}
@@
   \end{center}
-  \caption{\label{fig:power-law-high} A histogram of the marginalized
+  \caption{\label{fig:power-law-high} Histograms of the marginalized
     distribution for the three parameters $\Mmin$, $\Mmax$, and
@@
   \begin{center}
-    \plottwo{plots/mmin-parametric}{plots/mmin-non-parametric}
+    \plotone{plots/mmin}
   \end{center}
@@
 \begin{center}
-    \plottwo{plots/mmin-parametric-high}{plots/mmin-non-parametric-high}
+    \plotone{plots/mmin-high}
   \end{center}

### 2010-10-30 paper.tex
Vicky's changes.

@@
 \newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
+
+\newcommand{\changed}[1]{{\bf #1 }}
 
@@
   calculating their relative Bayesian evidence as supported by the
-  measurements.  The mass distribution of the low-mass systems is best
-  fit by a power-law, while the distribution of the complete sample is
-  best fit by the exponential model.  We examine the existence of a
-  ``gap'' between the most massive neutron stars and least massive
-  black holes, finding that the best model (the power law) fitted to
-  the low-mass systems only gives a minimum black-hole mass of 4.3
-  $\Msun$ (90\% confidence), while the best model (the exponential)
-  fitted to all 20 systems gives a minumum black-hole mass of 4.5
-  $\Msun$ (90\% confidence).  We conclude that our sample of black
-  hole masses provides strong evidence of a gap between the maximum
-  neutron star mass and minimum black hole mass.
+  measurements, \changed{taking into account the number of degrees of
+    freedom of each model.}  The mass distribution of the low-mass
+  systems is best fit by a power-law, while the distribution of the
+  complete sample is best fit by the exponential model.  \changed{This
+    difference clearly indicates that the character of the low mass
+    X-ray binary systems differs from the higher-mass systems.}  We
+  examine the existence of a ``gap'' between the most massive neutron
+  stars and least massive black holes, finding that the best model
+  (the power law) fitted to the low-mass systems only gives a minimum
+  black-hole mass of 4.3 $\Msun$ (90\% confidence), while the best
+  model (the exponential) fitted to all 20 systems gives a minumum
+  black-hole mass of 4.5 $\Msun$ (90\% confidence).  We conclude that
+  our sample of black hole masses provides strong evidence of a gap
+  between the maximum neutron star mass and minimum black hole mass.
+  \changed{Our results on the low-mass sample are in qualitative
+    agreement with those of \citet{Ozel2010}, although our broad model
+    selection analysis more reliably reveals the best-fit quantitative
+    description of the underlying mass distribution.  The results on
+    the complete sample of low- and higher-mass systems are in
+    qualitative agreement with \citet{Fryer2001} although the presence
+    of a mass gap remains theoretically unexplained.}
 \end{abstract}
@@
 The most massive stars probably end their lives with a supernova
-explosion, becoming stellar-mass black holes.  The mass distribution
-of such black holes can provide important clues to the end stages of
-evolution of these stars.  In addition, the mass distribution of
-stellar-mass black holes is an important input in calculations of
-rates of gravitational wave emission events from coalescing neutron
-star-black hole and black hole-black hole binaries in the LIGO
-gravitational wave observatory \citep{Abadie2010}.
+explosion or a quiet core collapse, becoming stellar-mass black holes.
+The mass distribution of such black holes can provide important clues
+to the end stages of evolution of these stars.  In addition, the mass
+distribution of stellar-mass black holes is an important input in
+calculations of rates of gravitational wave emission events from
+coalescing neutron star-black hole and black hole-black hole binaries
+in the LIGO gravitational wave observatory \citep{Abadie2010}.
 
@@
 can provide a measurement of the mass of the compact object in these
-systems.  There are now enough such measurements where the mass of the
-compact object is large enough to classify it as a black hole to
-attempt to statistically reconstruct the mass distribution of these
-stellar-mass black holes.
+systems.  \changed{The current sample of dynamically measured stellar
+  mass black holes includes 15 low-mass, Roche lobe overflowing
+  systems and 5 wind-fed higher-mass systems.  Hence, sophisticaed
+  statistical analyses are possible.}
 
 The first study of the mass distribution of stellar-mass black holes,
-in \citet{Bailyn1998}, examined a sample of seven transient X-ray
+in \citet{Bailyn1998}, examined a sample of seven low-mass X-ray
 binaries thought to contain a black hole, concluding in a Bayesian
@@
 masses.  \citet{Bailyn1998} found evidence of a ``gap'' between the
-least massive black hole and the upper limit for neutron star masses
-of 3 $\Msun$ \citep{Kalogera1996}.  Such a gap is puzzling in light of
-theoretical studies that predict a continuous distribution of compact
-object supernova remnant masses with a smooth transition from neutron
-stars to black holes \citep{Fryer2001}.
-
-A more recent study \citep{Ozel2010}, also in a Bayesian framework,
-largely confirmed these findings, examining 16 low-mass X-ray binary
-systems containing black holes and finding a strongly peaked
-distribution at $7.8 \pm 1.2 \, \Msun$.  \citet{Ozel2010} use two
-models for the mass function: a Gaussian and a decaying exponential
-with a minimum ``turn-on'' mass (motivated by the analytical model of
-the black-hole mass function in \citet{Fryer2001}).  \citet{Ozel2010}
-do not provide confidence limits for the minimum black hole mass,
-instead discussing only the maximum-posterior parameters for their
-models.  Nevertheless, it appears that their analysis confirms the
-existence of a mass gap.  \citet{Ozel2010} discuss possible selection
-effects that could lead to the appearance of a mass gap, but conclude
-these effects could not produce the observed gap, which they therefore
-claim is a real property of the black hole mass distribution.
+least massive black hole and a ``safe'' upper limit for neutron star
+masses of 3 $\Msun$ (e.g.\ \citet{Kalogera1996}).  Such a gap is
+puzzling in light of theoretical studies that predict a continuous
+distribution of compact object supernova remnant masses with a smooth
+transition from neutron stars to black holes \citep{Fryer2001}.
+\changed{(It should be noted that \citet{Fryer2001} considered binary
+  evolution effects only heuristically and put forward some possible
+  explanations for the gap from \citet{Bailyn1998} both in the context
+  of selection effects or in connection to the energetics of supernova
+  explosions.)}
+
+\changed{Towards the end of our anlysis work, we became aware of a
+  more recent study \citep{Ozel2010}, also in a Bayesian framework,
+  analyzing the low-mass X-ray binary sample.}  It largely confirmed
+these findings, examining 16 low-mass X-ray binary systems containing
+black holes and finding a strongly peaked distribution at $7.8 \pm 1.2
+\, \Msun$.  \citet{Ozel2010} use two models for the mass function: a
+Gaussian and a decaying exponential with a minimum ``turn-on'' mass
+(motivated by the analytical model of the black-hole mass function in
+\citet{Fryer2001}).  \changed{\citet{Ozel2010} do not provide
+  confidence limits for the minimum black hole mass, instead
+  discussing only the model parameters at the peak of their posterior
+  distributions.  They also do not perform any model selection
+  analysis; thus, they give the distribution of parameters within each
+  of their models, but cannot say which model is more likely to
+  correspond to the true distribution of black hole masses.}
+Nevertheless, it appears that their analysis confirms the existence of
+a mass gap.  \citet{Ozel2010} discuss possible selection effects that
+could lead to the appearance of a mass gap, but conclude these effects
+could not produce the observed gap, which they therefore claim is a
+real property of the black hole mass distribution.
 
@@
 
-We use a Bayesian Markov-Chain Monte Carlo analysis to explore various
-models for the black hole mass function for both samples.  We include
-both parameteric models, such as a Gaussian, and non-parameteric
-models where the mass function is represented by histograms with
-various numbers of bins.  (Our set of models includes those of
-\citet{Ozel2010} and \citet{Bailyn1998}.)  After computing posterior
-distributions for the model parameters, we use model selection
-techniques (including a new technique for efficient reversible-jump
-MCMC \citep{Farr2010}) to compare the evidence for the various models
-from both samples.
-
-We find, like \citet{Ozel2010} and \citet{Bailyn1998}, strong evidence
-for a mass gap among the best models for both systems, with the best
-model for the lower-mass systems giving a minimum black hole mass of
-4.3 $\Msun$ (at 90\% confidence), and the best model for the combined
-sample of lower- and high-mass systems giving a minimum black hole
-mass of 4.5 $\Msun$ (at 90\% confidence).  Among the models with lower
-evidence, most also have a mass gap.
-
-The theoretical model from \citet{Fryer2001}, a decaying exponential,
-considered for the low-mass systems in \citet{Ozel2010}, is strongly
-disfavored by our model selection.  We find that the lower-mass
-systems are best described by a power law, followed closely by a
-Gaussian (which is the other model from \citet{Ozel2010}).  We find
-that the theoretical model from \citet{Fryer2001} is, however, the
-preferred model for the combined sample of lower- and high-mass
-systems.  A model with two separate Gaussian peaks also has high
-evidence for the combined sample of systems.  While the failure of the
-theoretically-motivated model to explain the lower-mass, Roche
-lobe-filling systems suggests gaps in our understanding of the
-formation of these systems, it is encouraging for population synthesis
-studies \fixme{Vicky, do you have a couple of go-to papers for
-  pop. synth we could cite here?} that the model of \citet{Fryer2001}
-is a good fit to the full sample including both lower- and high-mass
-systems.
+We use a Bayesian Markov-Chain Monte Carlo (MCMC) analysis to
+quantitatively assess a wide range of models for the black hole mass
+function for both samples.  We include both parameteric models, such
+as a Gaussian, and non-parameteric models where the mass function is
+represented by histograms with various numbers of bins.  (Our set of
+models includes those of \citet{Ozel2010} and \citet{Bailyn1998}.)
+After computing posterior distributions for the model parameters, we
+use model selection techniques (including a new technique for
+efficient reversible-jump MCMC \citep{Farr2010}) to compare the
+evidence for the various models from both samples.
+
+In agreement with \citet{Ozel2010} and \citet{Bailyn1998}, we find
+strong evidence for a mass gap among the best models for both samples,
+with the best model for the low-mass systems giving a minimum black
+hole mass of 4.3 $\Msun$ (at 90\% confidence), and the best model for
+the combined sample of lower- and high-mass systems giving a minimum
+black hole mass of 4.5 $\Msun$ (at 90\% confidence).  Further, among
+the models with lower evidence, most also have a mass gap.
+
+\changed{We find that, for the low-mass X-ray binary sample, the
+  theoretical model from \citet{Fryer2001}, a decaying exponential is
+  strongly disfavored by our model selection.  We find that the
+  low-mass systems are best described by a power law, followed
+  closely by a Gaussian (which is the other model from
+  \citet{Ozel2010}).  On the other hand, we find that the theoretical
+  model from \citet{Fryer2001} is the preferred model for the combined
+  sample of low- and high-mass X-ray binaries.  A model with two
+  separate Gaussian peaks also has relatively high evidence for the
+  combined sample of systems.  The difference in best-fitting model
+  clearly indicates that the character of the low mass X-ray binary
+  systems differs from the higher-mass systems.}
 
@@
 black-hole mass distributions.  We will first analyze the 15 low-mass
-systems alone (Section \ref{sec:results-lower-mass}), and then the
+systems alone (Section \ref{sec:results-low-mass}), and then the
 complete sample of 20 systems (Section \ref{sec:higher-mass}).
@@
 
-  \caption{\label{fig:low-masses} The mass distributions implied by
-    Equation \eqref{eq:mass-function} and the assumed distributions on
-    observational parameters given in Table \ref{tab:sources} for the
-    lower-mass sources.  The significant asymmetry and long tails in
-    many of these distributions are the result of the non-linear
-    relationship (Equation \eqref{eq:mass-function}) between $M$, $f$,
-    $q$, and $i$.}
+  \caption{\label{fig:low-masses} The individual mass distributions
+    implied by Equation \eqref{eq:mass-function} and the assumed
+    distributions on observational parameters $f$, $q$, and $i$ given
+    in Table \ref{tab:sources} for the low-mass sources.  The
+    significant asymmetry and long tails in many of these
+    distributions are the result of the non-linear relationship
+    (Equation \eqref{eq:mass-function}) between $M$, $f$, $q$, and
+    $i$.}
 \end{figure}
@@
     observed data in Table \ref{tab:sources} using Equation
-    \eqref{eq:mass-function}.  The asymmetry and long tails in these
+    \eqref{eq:mass-function}.  (Similar to Figure
+    \ref{fig:low-masses}.)  The asymmetry and long tails in these
     distributions are the result of the non-linear relationship
@@
 various models for the underlying mass distribution from which the
-low-mass systems in Table \ref{tab:sources} were drawn.  Results of
-the analysis appear in Section \ref{sec:results}.
+low-mass sample and the complete sample of X-ray binary systems in
+Table \ref{tab:sources} were drawn.  The results of our analysis are
+presented in Section \ref{sec:results}.
 
@@
 \citep{Ozel2010,Bailyn1998} have considered models with only two
-parameters, evaluating $p(\vtheta|d)$ on a grid.  Many of our models
-for the underlying mass distribution have three or more parameters.
-Exploring the entirety of these parameter spaces with a grid rapidly
-becomes prohibitive as the number of parameters increases.  A more
-efficient way to explore the distribution $p(\vtheta | d)$ is to use a
-Markov Chain Monte Carlo (MCMC) method.  MCMC methods produce a chain
-(sequence) of parameter samples, $\{ \vtheta_i \, | \, i = 1, \ldots
-\}$, such that a particular parameter sample, $\vtheta$, appears in
-the sequence with a frequency equal to its posterior probability,
-$p(\vtheta|d)$.  In this way, regions of parameter space where
-$p(\vtheta|d)$ is large are sampled densely while regions where
-$p(\vtheta|d)$ is small are effectively ignored.
+parameters; for such models evaluating $p(\vtheta|d)$ on a grid may be
+a reliable method.  Many of our models for the underlying mass
+distribution have three or more parameters.  Exploring the entirety of
+these parameter spaces with a grid rapidly becomes prohibitive as the
+number of parameters increases.  A more efficient way to explore the
+distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
+(MCMC) method.  MCMC methods produce a chain (sequence) of parameter
+samples, $\{ \vtheta_i \, | \, i = 1, \ldots \}$, such that a
+particular parameter sample, $\vtheta$, appears in the sequence with a
+frequency equal to its posterior probability, $p(\vtheta|d)$.  In this
+way, regions of parameter space where $p(\vtheta|d)$ is large are
+sampled densely while regions where $p(\vtheta|d)$ is small are
+effectively ignored.
 
@@
 that the evidence for the histogram models (see Sections
-\ref{sec:model-selection}, \ref{sec:lower-mass-model-selection}, and
-\ref{sec:high-mass-model-selection}) from both the lower-mass and
+\ref{sec:model-selection}, \ref{sec:low-mass-model-selection}, and
+\ref{sec:high-mass-model-selection}) from both the low-mass and
 complete datasets is decreasing as the number of bins reaches five,
@@
 also discuss model selection results.  The results in Section
-\ref{sec:results-lower-mass} apply to the lower-mass sample of
+\ref{sec:results-low-mass} apply to the low-mass sample of
 systems, while those of Section \ref{sec:higher-mass} apply to the
@@
 
-\subsection{Lower-Mass Systems}
-\label{sec:results-lower-mass}
+\subsection{Low-Mass Systems}
+\label{sec:results-low-mass}
 
@@
   \caption{\label{fig:power-law} Histograms of the marginalized
-    distribution for the three parameters $\Mmin$, $\Mmax$, and
-    $\alpha$ from the power-law model.  The marginalized distribution
-    for $\alpha$ is broad, with $-11.8 < \alpha < 6.8$ enclosing 90\%
-    of the probability.  We have $p(\alpha < 0) = 0.6$; the median
-    value is $\alpha = -3.35$.  The broad distribution for $\alpha$
-    (and the other parameters) is due to correlations between the
-    parameters discussed in the main text; see Figure
-    \ref{fig:power-law-2D}.}
+    distribution for the three parameters $\Mmin$ (top, left), $\Mmax$
+    (top, right), and $\alpha$ (bottom) from the power-law model.  The
+    marginalized distribution for $\alpha$ is broad, with $-11.8 <
+    \alpha < 6.8$ enclosing 90\% of the probability.  We have
+    $p(\alpha < 0) = 0.6$; the median value is $\alpha = -3.35$.  The
+    broad distribution for $\alpha$ (and the other parameters) is due
+    to correlations between the parameters discussed in the main text;
+    see Figure \ref{fig:power-law-2D}.}
 \end{figure}
@@
 the need to match the relatively narrow range in mass of the
-lower-mass systems.  When $\alpha$ is negative, the resulting mass
+low-mass systems.  When $\alpha$ is negative, the resulting mass
 distribution slopes down; $\Mmin$ is constrained to be near the lowest
@@
 the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
-that the first Gaussian is centered around the peaks of the lower-mass
+that the first Gaussian is centered around the peaks of the low-mass
 distributions.  The second Gaussian's parameter distributions are much
@@
 4.7$ (see Sections \ref{sec:model-selection} and
-\ref{sec:lower-mass-model-selection} for discussion).
+\ref{sec:low-mass-model-selection} for discussion).
 
@@
     $4.7$ on this dataset (see Sections \ref{sec:model-selection} and
-    \ref{sec:lower-mass-model-selection} for discussion).}
+    \ref{sec:low-mass-model-selection} for discussion).}
 \end{figure}
@@
 histogram is the most probable of the histogram models for the
-low-mass data (see Section \ref{sec:lower-mass-model-selection} for
+low-mass data (see Section \ref{sec:low-mass-model-selection} for
 discussion).  This occurs because the extra fitting power does not
@@
 
-\subsubsection{Lower-Mass Model Selection}
-\label{sec:lower-mass-model-selection}
+\subsubsection{Low-Mass Model Selection}
+\label{sec:low-mass-model-selection}
 We have performed a suite of 500 independent reversible-jump MCMCs
-jumping between all the models (both parametric and non-parametric) in
-Section \ref{sec:models} using the single-model MCMC samples to
-construct an efficient jump proposal for each model as described above
-(see Appendix \ref{sec:reversible-jump-mcmc}).  The numbers of counts in
-each model are consistent across the MCMCs in the suite; Figure
-\ref{fig:rj} displays the average probability for each model across
-the suite, along with the 1-$\sigma$ errors on the average inferred
-from the standard deviation of the model counts across the suite.
-Table \ref{tab:rj} gives the numerical values of the average
-probability for each model across the suite of MCMCs.
+jumping between all the models (both parametric and non-parametric)
+described in Section \ref{sec:models} using the single-model MCMC
+samples to construct an efficient jump proposal for each model as
+described above (see Appendix \ref{sec:reversible-jump-mcmc}).  The
+numbers of counts in each model are consistent across the MCMCs in the
+suite; Figure \ref{fig:rj} displays the average probability for each
+model across the suite, along with the 1-$\sigma$ errors on the
+average inferred from the standard deviation of the model counts
+across the suite.  Table \ref{tab:rj} gives the numerical values of
+the average probability for each model across the suite of MCMCs.
 
@@
   \caption{\label{tab:rj} Relative probabilities of the various models
-    from Section \ref{sec:models} implied by the lower-mass data.  These probabilities have been 
+    from Section \ref{sec:models} implied by the low-mass data.  These probabilities have been 
     computed from reversible-jump
@@
 Section \ref{sec:power-law} and displayed in Figure
-\ref{fig:power-law-2D} for the lower-mass subset of the observations.
+\ref{fig:power-law-2D} for the low-mass subset of the observations.
 
@@
   \caption{\label{fig:power-law-high} Histograms of the marginalized
-    distribution for the three parameters $\Mmin$, $\Mmax$, and
-    $\alpha$ from the power-law model including the higher-mass
-    samples in the MCMC.  The distribution for $\Mmax$ is quite broad
-    because the best fit power laws slope downward ($\alpha < 0$),
-    making this parameter less relevant.  The range $-5.05 \leq \alpha
-    \leq -1.77$ encloses 90\% of the probability; the median value of
-    $\alpha$ is -3.23.  The presence of the higher-mass samples in the
-    analysis produces a distinctive tail, eliminating the correlations
-    discussed in Section \ref{sec:power-law} and displayed in Figure
-    \ref{fig:power-law-2D} for the lower-mass subset of the
-    observations. }
+    distribution for the three parameters $\Mmin$ (top, left), $\Mmax$
+    (top, right), and $\alpha$ (bottom) from the power-law model
+    including the higher-mass samples in the MCMC.  The distribution
+    for $\Mmax$ is quite broad because the best fit power laws slope
+    downward ($\alpha < 0$), making this parameter less relevant.  The
+    range $-5.05 \leq \alpha \leq -1.77$ encloses 90\% of the
+    probability; the median value of $\alpha$ is -3.23.  The presence
+    of the higher-mass samples in the analysis produces a distinctive
+    tail, eliminating the correlations discussed in Section
+    \ref{sec:power-law} and displayed in Figure \ref{fig:power-law-2D}
+    for the low-mass subset of the observations. }
 \end{figure}
@@
   \caption{\label{fig:exp-cutoff-high} The marginalized distributions
-    for the exponential parameters $\Mmin$ and $M_0$ (Section
-    \ref{sec:exponential}) from an analysis including the higher-mass
-    systems.  The distribution for the scale mass, $M_0$, has moved to
-    higher masses relative to Figure \ref{fig:exp-marginal} to fit the
-    tail of the mass distribution; we now have $2.8292 \leq M_0 \leq
-    7.9298$ with 90\% confidence, with median 4.7003.  The
-    distribution for $\Mmin$ is less affected, though it has broadened
-    somewhat toward low masses.}
+    for the exponential parameters $\Mmin$ (top) and $M_0$ (bottom)
+    defined in Section \ref{sec:exponential} from an analysis
+    including the higher-mass systems.  The distribution for the scale
+    mass, $M_0$, has moved to higher masses relative to Figure
+    \ref{fig:exp-marginal} to fit the tail of the mass distribution;
+    we now have $2.8292 \leq M_0 \leq 7.9298$ with 90\% confidence,
+    with median 4.7003.  The distribution for $\Mmin$ is less
+    affected, though it has broadened somewhat toward low masses.}
 \end{figure}
@@
     for the Gaussian parameters when the higher-mass objects are
-    included in the mass distribution.  The mean mass, $\mu$, and the
-    mass standard deviation, $\sigma$, are both increased relative to
-    Figure \ref{fig:gaussian} to account for the broader distribution
-    and higher-mass tail.  The peak of the underlying mass
-    distribution lies in the range $7.8660 \leq \mu \leq 10.9836$ with
-    90\% confidence; the median value is 9.2012.}
+    included in the mass distribution.  The mean mass, $\mu$ (solid
+    histogram), and the mass standard deviation, $\sigma$ (dashed
+    histogram), are both increased relative to Figure
+    \ref{fig:gaussian} to account for the broader distribution and
+    higher-mass tail.  The peak of the underlying mass distribution
+    lies in the range $7.8660 \leq \mu \leq 10.9836$ with 90\%
+    confidence; the median value is 9.2012.}
 \end{figure}
@@
 to Figure \ref{fig:two-gaussian}, there are two well-defined,
-separated peaks; the lower-mass peak reproduces the results from the
+separated peaks; the low-mass peak reproduces the results from the
 low-mass samples, while the higher-mass peak ($13.5534 \leq \mu_2 \leq
@@
     the analysis.  In stark contrast to Figure \ref{fig:two-gaussian},
-    there are two well-defined, separated peaks; the lower-mass peak
+    there are two well-defined, separated peaks; the low-mass peak
     reproduces the results from the low-mass samples, while the
@@
 Repeating the model selection analysis discussed in Section
-\ref{sec:lower-mass-model-selection} for the sample including the
+\ref{sec:low-mass-model-selection} for the sample including the
 higher-mass systems, we find that the model probabilities have changed
@@
 complete sample, with the two-Gaussian model the second-most favored.
-The ranking of models differs significantly from the lower-mass
+The ranking of models differs significantly from the low-mass
 samples.  The improvement of the exponential model relative to the
-lower-mass analysis is encouraging for theoretical calculations that
+low-mass analysis is encouraging for theoretical calculations that
 attempt to model the entire population of X-ray binaries with this
@@
 
-\section{The Minimum Mass of the Black Hole Mass Distribution}
+\section{The Minimum Black Hole Mass}
 \label{sec:minimum-mass}
 
-It is interesting to use our models for the underlying mass
-distribution of X-ray binary black holes to try to place constraints
-on the minimum black hole mass.  \citet{Bailyn1998} addressed this
-question in the context of a ``mass gap'' between the most massive
-neutron stars and the least massive black holes.  The more recent
-study of \citet{Ozel2010} also looked for a mass gap using a subset of
-the models and systems presented here.  Both works found that the
-minimum black hole mass is significantly above the maximum neutron
-star mass.
+It is interesting to use our models for the underlying black hole mass
+distribution in X-ray binaries to place constraints on the minimum
+black hole mass.  \citet{Bailyn1998} addressed this question in the
+context of a ``mass gap'' between the most massive neutron stars and
+the least massive black holes.  The more recent study of
+\citet{Ozel2010} also looked for a mass gap using a subset of the
+models and systems presented here.  Both works found that the minimum
+black hole mass is significantly above the maximum neutron star mass
+\citep{Kalogera1996} of $\sim 3 \Msun$ \changed{(though
+  \citet{Ozel2010} only state their evidence for a gap in terms of the
+  maximum-posterior parameters for their distributions)}.
 
 The distribution of the minimum black hole mass from the analysis of
-the lower-mass samples is displayed in Figure \ref{fig:min-mass}.  For
+the low-mass samples is displayed in Figure \ref{fig:min-mass}.  For
 all distributions, the minimum black hole mass is defined as the 1\%
@@
     black hole mass calculated from the MCMC samples for the models in
-    Section \ref{sec:models} applied to the lower-mass systems.  The
+    Section \ref{sec:models} applied to the low-mass systems.  The
     minimum black hole mass is defined as the 1\% mass quantile.  For
@@
     confidence limit on the minimum black hole mass is 4.3 $\Msun$ and
-    2.9 $\Msun$, respectively.}
+    2.9 $\Msun$, respectively.  In all plots, we indicate the 90\%
+    confidence limit on the minimum black hole mass with a vertical
+    line.}
 \end{figure}
 
-We find that the best-fit model for the lower-mass systems (the
+We find that the best-fit model for the low-mass systems (the
 power-law) has a 90\% confidence limit on the minimum black hole mass
 of 4.3 $\Msun$.  This is significantly above the maximum
-theoretically-allowed neutron star mass, $\sim 3 \Msun$
-\citep{Kalogera1996}\fixme{Vicky, do you have a more recent reference
-  than this?}, so we conclude that the lower-mass systems show strong
-evidence of a mass gap.
+theoretically-allowed neutron star mass, $\sim 3 \Msun$ (e.g.\
+\citet{Kalogera1996}).  Hence we conclude that the low-mass systems
+show strong evidence of a mass gap.
 
@@
     The minimum black hole mass is defined as the 1\% mass quantile.
-    For the most favored model, the exponential, the 90\% confidence
-    limit on the minimum black hole mass is 4.5 $\Msun$.}
+    For the two most favored models, the exponential and two-Gaussian,
+    the 90\% confidence limit on the minimum black hole mass is 4.5
+    $\Msun$ and 2.3 $\Msun$, respectively.  For every model, we
+    indicate the 90\% confidence limit on the minimum black hole mass
+    with a vertical line.}
 \end{figure}
 
-\section{Conclusion}
+\section{Summary and Conclusion}
 
@@
 stellar-mass black holes in X-ray binary systems.  We considered
-separately a sample of 15 lower-mass, Roche lobe-filling systems and a
-sample of 20 systems containing the 15 lower-mass systems and five
+separately a sample of 15 low-mass, Roche lobe-filling systems and a
+sample of 20 systems containing the 15 low-mass systems and five
 higher-mass, wind-fed X-ray binaries.  We used MCMC methods to sample
@@
 distribution that are implied by these data.  For both sets of
-systems, we used reversible jump MCMCs (exploiting a new algorithm for
+samples, we used reversible jump MCMCs (exploiting a new algorithm for
 efficient jump proposals in such calculations) to perform model
-selection on the suite of models.
-
-For the lower-mass systems, we found the limits on model parameters in
+selection on the suite of models.  \changed{The consideration of a
+  broad range of models and the model-selection analysis, along with
+  consideration of the full posterior distribution on the minimum
+  black hole mass significantly expand earlier analyses of black hole
+  mass measurements \citep{Bailyn1998,Ozel2010}.}
+
+For the low-mass systems, we found the limits on model parameters in
 Tables \ref{tab:low-mass-parametric} and
@@
 from the model selection are given in Table \ref{tab:rj}.  The most
-favored model for the lower-mass systems is a power law.  The
-equivalent limits on the model parameters for the combined systems are
-given in Tables \ref{tab:high-mass-parametric} and
-\ref{tab:high-mass-non-parametric}.  Unlike the lower-mass systems,
-the most favored model for the combined sample is the exponential
-model.
+favored model for the low-mass systems is a power law.  The equivalent
+limits on the model parameters for the combined systems are given in
+Tables \ref{tab:high-mass-parametric} and
+\ref{tab:high-mass-non-parametric}.  Unlike the low-mass systems, the
+most favored model for the combined sample is the exponential model.
+\changed{This difference clearly indicates that the character of the
+  low mass X-ray binary systems differs from the higher-mass systems.}
 
 We found strong evidence for a mass gap between the most massive
-neutron stars, $\sim 3 \Msun$ \citep{Kalogera1996}, and the least
-massive black holes.  For the lower-mass systems, the most favored,
-power law model gives a black hole mass distribution whose 1\%
-quantile lies above $4.3 \Msun$ with 90\% confidence.  For the
-complete sample of systems, the most favored, exponential model gives
-a black hole mass distribution whose 1\% quantile lise above $4.5
-\Msun$ with 90\% confidence.  The existence of a mass gap was also
-observed in \citet{Bailyn1998} (which used different models in a study
-of 7 systems) and \citet{Ozel2010} (which did not contain a power law
-model, and applied the exponential model to a sample of 16 lower-mass
-systems, where it is strongly disfavored).  The existence of a mass
-gap poses a challenge to current theoretical models of compact object
-formation \citep{Fryer2001}, which predict a continuous mass
-distribution from neutron star masses to stellar mass black hole
-masses.
+neutron stars and the least massive black holes.  For the low-mass
+systems, the most favored, power law model gives a black hole mass
+distribution whose 1\% quantile lies above $4.3 \Msun$ with 90\%
+confidence.  For the complete sample of systems, the most favored,
+exponential model gives a black hole mass distribution whose 1\%
+quantile lise above $4.5 \Msun$ with 90\% confidence.
+\changed{Although the study methodology was different, the existence
+  of a mass gap was pointed out first by \citet{Bailyn1998} and most
+  recently by \citet{Ozel2010} (who did not consider a power law
+  model, and applied the exponential model to the low-mass systems,
+  where it is strongly disfavored compared to the power-law model).}
+
+\changed{Theoretical expectations for the black hole mass distribution
+  have been examined in \citet{Fryer2001}.  They considered results of
+  supernova explosion and fallback simulations \citep{Fryer1999}
+  applied to single star populations; they also included a heuristic
+  treatment of the possible effects of binary evolution on the black
+  hole mass distribution.  It is interesting that we find the
+  most-favored model for the complete sample to be an exponential, as
+  discussed by \citet{Fryer1999}.  On the other hand, we find the
+  most-favored model for the low-mass sample to be a power law, with
+  the exponential model strongly disfavored for this sample. In
+  agreement with \citet{Bailyn1998} and \citet{Ozel2010}, we too
+  conclude that both the low-mass and complete samples require the
+  presence of a gap in the 3--4 or 4.5 $\Msun$ range.
+
+  \citet{Fryer1999} discussed two possible causes of such a gap: (1) a
+  step-like dependence of supernova energy on progenitor mass or (2)
+  selection biases.  Current simulations of core collapse in massive
+  stars may shed light on the dependence of supernova energy on
+  progenitor mass.  Selection biases can occur because the X-ray
+  binaries with very low-mass black holes systems are more likely to
+  be persistently Roche-lobe overflowing, preventing dynamical mass
+  measurements.  \citet{Ozel2010} conclude that the presence of such
+  biases is not enough to account for the gap, arguing that the number
+  (26) of observed persistent X-ray sources not known to be neutron
+  stars is insufficient to populate the 2--5 $\Msun$ region of any
+  black hole mass distribution that rises toward low masses.
+  Population synthesis models incorporating sophisticated treatment of
+  binary evolution and transient behavior (e.g.\
+  \citet{Fragos2008,Fragos2009}) could help settle this possibility.}
 

### 2010-10-31 paper.tex
All changes in near-final version.

@@
 
-\author{Andrew Cantrell \and Laura Kreidberg \and Charles Bailyn}
+\author{Andrew Cantrell \and Laura Kreidberg \and Charles D. Bailyn}
 
@@
   examine the existence of a ``gap'' between the most massive neutron
-  stars and least massive black holes, finding that the best model
-  (the power law) fitted to the low-mass systems only gives a minimum
-  black-hole mass of 4.3 $\Msun$ (90\% confidence), while the best
-  model (the exponential) fitted to all 20 systems gives a minumum
-  black-hole mass of 4.5 $\Msun$ (90\% confidence).  We conclude that
-  our sample of black hole masses provides strong evidence of a gap
-  between the maximum neutron star mass and minimum black hole mass.
-  \changed{Our results on the low-mass sample are in qualitative
-    agreement with those of \citet{Ozel2010}, although our broad model
-    selection analysis more reliably reveals the best-fit quantitative
-    description of the underlying mass distribution.  The results on
-    the complete sample of low- and higher-mass systems are in
-    qualitative agreement with \citet{Fryer2001} although the presence
-    of a mass gap remains theoretically unexplained.}
+  stars and least massive black holes.  \changed{Our analysis
+    generates distributions of lower bounds on the black hole mass
+    distribution; we use the 1\% quantile of the mass distribution,
+    $M_{1\%}$, for a lower bound because some of our distributions do
+    not have a sharp lower limit.  For the best model (the power law)
+    fitted to the low-mass systems this distribution of lower-bounds
+    has $M_{1\%} > 4.3$ $\Msun$ with 90\% confidence, while the best
+    model (the exponential) fitted to all 20 systems has $M_{1\%} >
+    4.5$ $\Msun$ with 90\% confidence.}  We conclude that our sample
+  of black hole masses provides strong evidence of a gap between the
+  maximum neutron star mass and \changed{the lower bound on} black
+  hole masses.  \changed{Our results on the low-mass sample are in
+    qualitative agreement with those of \citet{Ozel2010}, although our
+    broad model selection analysis more reliably reveals the best-fit
+    quantitative description of the underlying mass distribution.  The
+    results on the complete sample of low- and higher-mass systems are
+    in qualitative agreement with \citet{Fryer2001} although the
+    presence of a mass gap remains theoretically unexplained.}
 \end{abstract}
@@
 can provide a measurement of the mass of the compact object in these
-systems.  \changed{The current sample of dynamically measured stellar
-  mass black holes includes 15 low-mass, Roche lobe overflowing
-  systems and 5 wind-fed higher-mass systems.  Hence, sophisticaed
-  statistical analyses are possible.}
+systems.  \changed{The current sample of stellar mass black holes with
+  dynamically measured masses includes 15 low-mass, Roche lobe
+  overflowing systems and 5 wind-fed higher-mass systems.  Hence,
+  sophisticated statistical analyses of the black hole mass
+  distribution in these systems are possible.}
 
@@
 analysis that the mass function was strongly peaked around seven solar
-masses.  \citet{Bailyn1998} found evidence of a ``gap'' between the
-least massive black hole and a ``safe'' upper limit for neutron star
-masses of 3 $\Msun$ (e.g.\ \citet{Kalogera1996}).  Such a gap is
-puzzling in light of theoretical studies that predict a continuous
-distribution of compact object supernova remnant masses with a smooth
-transition from neutron stars to black holes \citep{Fryer2001}.
-\changed{(It should be noted that \citet{Fryer2001} considered binary
-  evolution effects only heuristically and put forward some possible
-  explanations for the gap from \citet{Bailyn1998} both in the context
-  of selection effects or in connection to the energetics of supernova
-  explosions.)}
-
-\changed{Towards the end of our anlysis work, we became aware of a
+masses.  \changed{(A similar analysis of the neutron star mass
+  distribution can be found in \citet{Finn1994}.)}  \citet{Bailyn1998}
+found evidence of a ``gap'' between the least massive black hole and a
+``safe'' upper limit for neutron star masses of 3 $\Msun$ (e.g.\
+\citet{Kalogera1996}).  Such a gap is puzzling in light of theoretical
+studies that predict a continuous distribution of compact object
+supernova remnant masses with a smooth transition from neutron stars
+to black holes \citep{Fryer2001}.  \changed{(But note that
+  \citet{Fryer2001} considered binary evolution effects only
+  heuristically and put forward some possible explanations for the gap
+  from \citet{Bailyn1998} both in the context of selection effects or
+  in connection to the energetics of supernova explosions.)}
+
+\changed{Towards the end of our analysis work, we became aware of a
   more recent study \citep{Ozel2010}, also in a Bayesian framework,
@@
 quantitatively assess a wide range of models for the black hole mass
-function for both samples.  We include both parameteric models, such
-as a Gaussian, and non-parameteric models where the mass function is
+function for both samples.  We include both parametric models, such
+as a Gaussian, and non-parametric models where the mass function is
 represented by histograms with various numbers of bins.  (Our set of
@@
 
-In agreement with \citet{Ozel2010} and \citet{Bailyn1998}, we find
-strong evidence for a mass gap among the best models for both samples,
-with the best model for the low-mass systems giving a minimum black
-hole mass of 4.3 $\Msun$ (at 90\% confidence), and the best model for
-the combined sample of lower- and high-mass systems giving a minimum
-black hole mass of 4.5 $\Msun$ (at 90\% confidence).  Further, among
-the models with lower evidence, most also have a mass gap.
+\changed{We define the ``minimum black hole mass'' to be the 1\%
+  quantile, $M_{1\%}$, in the black hole mass distribution (see
+  Section \ref{sec:minimum-mass}).  In agreement with \citet{Ozel2010}
+  and \citet{Bailyn1998}, we find strong evidence for a mass gap among
+  the best models for both samples.  Our analysis gives distributions
+  for $M_{1\%}$ implied by the data in the context of each of our
+  models for the black hole mass distribution.  In the context of the
+  best model for the low-mass systems (a power-law), the distribution
+  for $M_{1\%}$ has $M_{1\%} > 4.3$ $\Msun$ (at 90\% confidence); in
+  the context of the best model for the combined sample of lower- and
+  high-mass systems the distribution of $M_{1\%}$ has $M_{1\%} > 4.5$
+  $\Msun$ with 90\% confidence.  Further, in the context of models
+  with lower evidence, most also have a mass gap, with 90\% confidence
+  bounds on $M_{1\%}$ significantly above a ``safe'' maximum neutron
+  star mass of 3 $\Msun$ \citep{Kalogera1996}.}
 
@@
   systems differs from the higher-mass systems.}
+
+\changed{The structure of this paper is as follows.  In Section
+  \ref{sec:systems} we discuss the 15 systems that comprise the
+  low-mass X-ray binary black hole sample and the 5 additional
+  higher-mass, wind-fed systems that make up the complete sample.  In
+  Section \ref{sec:models} we discuss the Bayesian techniques we use
+  to analyze the black hole mass distribution, the techniques we use
+  for model selection, and the parametric and non-parametric models we
+  will for the black hole mass distribution.  In Section
+  \ref{sec:results} we discuss the results of our analysis and model
+  selection.  In Section \ref{sec:minimum-mass} we discuss the
+  distribution of the minimum black hole mass implied by the analysis
+  of Section \ref{sec:results}.  In Section \ref{sec:conclusion} we
+  summarize our results and comment on the significance of the
+  observed mass gap in the context of theoretical models.  Appendix
+  \ref{sec:mcmc} describes MCMC techniques in some detail.  Appendix
+  \ref{sec:reversible-jump-mcmc} explains our novel algorithm for
+  efficiently performing the reversible jump MCMC computations used in
+  the model comparisons of Section \ref{sec:results} (but see also
+  \citet{Farr2010}).}
 
@@
   \end{center}
-  \caption{\label{tab:sources} The source parameters for the 20 low
-    mass X-ray binaries used in this work.  The first 15 systems have
+  \caption{\label{tab:sources} The source parameters for the 20 X-ray binaries used in this work.  The first 15 systems have
     low-mass secondaries that feed the black hole via Roche lobe
@@
 distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
-(MCMC) method.  MCMC methods produce a chain (sequence) of parameter
-samples, $\{ \vtheta_i \, | \, i = 1, \ldots \}$, such that a
-particular parameter sample, $\vtheta$, appears in the sequence with a
-frequency equal to its posterior probability, $p(\vtheta|d)$.  In this
-way, regions of parameter space where $p(\vtheta|d)$ is large are
-sampled densely while regions where $p(\vtheta|d)$ is small are
-effectively ignored.
+(MCMC) method (see Appendix \ref{sec:mcmc}).  MCMC methods produce a
+chain (sequence) of parameter samples, $\{ \vtheta_i \, | \, i = 1,
+\ldots \}$, such that a particular parameter sample, $\vtheta$,
+appears in the sequence with a frequency equal to its posterior
+probability, $p(\vtheta|d)$.  In this way, regions of parameter space
+where $p(\vtheta|d)$ is large are sampled densely while regions where
+$p(\vtheta|d)$ is small are effectively ignored.
 
@@
 absence of any information about the distribution of parameters, it is
-best to choose a prior that is broad and uniformative to avoid biasing
+best to choose a prior that is broad and uninformative to avoid biasing
 the posterior as much as possible.
@@
 black-hole mass distribution may be well-represented by a decaying
-exponential with a minimum mass: \fixme{Vicky, do you want to say
-  anything more here?}
+exponential with a minimum mass: 
 \begin{equation}
@@
 \end{multline}
-The probability is a linear combination of two Gaussians with weight
-$\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose combined
-conditions on $\mu_i$ and $\sigma_i$ that ensure that most of the mass
-probability lies below $40 \Msun$ with the prior 
+The probability is a linear combination of two Gaussians with weights
+$\alpha$ and $1-\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose
+combined conditions on $\mu_i$ and $\sigma_i$ that ensure that most of
+the mass probability lies below $40 \Msun$ with the prior
 \begin{equation}
@@
 the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
-> 0.8$ with 62\% probability, clearly favoring the first Gaussian.
-The distributions for $\mu_1$ and $\sigma_1$ are similar to those of
-the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
-that the first Gaussian is centered around the peaks of the low-mass
-distributions.  The second Gaussian's parameter distributions are much
-broader.  The second Gaussian appears to be sampling the tail of the
-mass samples.  In spite of the ability to match a more complicated
-distribution, we find that this model is strongly disfavored relative
-to the single-Gaussian model for this dataset:
+> 0.8$ with 62\% probability, clearly favoring the Gaussian with
+smaller mean.  The distributions for $\mu_1$ and $\sigma_1$ are
+similar to those of the single Gaussian displayed in Figure
+\ref{fig:gaussian}, indicating that this Gaussian is centered around
+the peaks of the low-mass distributions.  The second Gaussian's
+parameter distributions are much broader.  The second Gaussian appears
+to be sampling the tail of the mass samples.  In spite of the extra
+degrees of freedom in this model, we find that this model is strongly
+disfavored relative to the single-Gaussian model for this dataset:
 $p(\textnormal{Gaussian}|d) / p(\textnormal{Two Gaussian}|d) \simeq
@@
 
-\subsubsection{Low-Mass Model Selection}
+\subsubsection{Model Selection for the Low-Mass Sample}
 \label{sec:low-mass-model-selection}
@@
     reversible-jump MCMC with the efficient jump proposal algorithm
-    described in Section \ref{sec:reversible-jump-mcmc}.  In
-    increasing order along the $x$-axis, the models are the power-law
-    of Section \ref{sec:power-law} (PL), the decaying exponential of
-    Section \ref{sec:exponential} (E), the single Gaussian of Section
-    \ref{sec:gaussian} (G), the double Gaussian of Section
-    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
-    five-bin histogram models of Section
+    described in Section \ref{sec:reversible-jump-mcmc}.  (See also
+    Table \ref{tab:rj}.)  In increasing order along the $x$-axis, the
+    models are the power-law of Section \ref{sec:power-law} (PL), the
+    decaying exponential of Section \ref{sec:exponential} (E), the
+    single Gaussian of Section \ref{sec:gaussian} (G), the double
+    Gaussian of Section \ref{sec:gaussian} (TG), and the one-, two-,
+    three-, four-, and five-bin histogram models of Section
     \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
@@
   \caption{\label{tab:rj} Relative probabilities of the various models
-    from Section \ref{sec:models} implied by the low-mass data.  These probabilities have been 
+    from Section \ref{sec:models} implied by the low-mass data.  (See also Figure \ref{fig:rj}.)  These probabilities have been 
     computed from reversible-jump
@@
 \citet{Linder2008} discuss these issues in the context of cosmological
-model selection, concluding with a similar warning against
-over-reliance on model selection probabilities.
+model selection, concluding with a warning against over-reliance on
+model selection probabilities.
 
@@
   \caption{\label{fig:power-law-high} Histograms of the marginalized
-    distribution for the three parameters $\Mmin$ (top, left), $\Mmax$
-    (top, right), and $\alpha$ (bottom) from the power-law model
-    including the higher-mass samples in the MCMC.  The distribution
-    for $\Mmax$ is quite broad because the best fit power laws slope
-    downward ($\alpha < 0$), making this parameter less relevant.  The
-    range $-5.05 \leq \alpha \leq -1.77$ encloses 90\% of the
-    probability; the median value of $\alpha$ is -3.23.  The presence
-    of the higher-mass samples in the analysis produces a distinctive
-    tail, eliminating the correlations discussed in Section
-    \ref{sec:power-law} and displayed in Figure \ref{fig:power-law-2D}
-    for the low-mass subset of the observations. }
+    distribution for the three parameters $\Mmin$ (top, dashed
+    histogram), $\Mmax$ (top, solid histogram), and $\alpha$ (bottom)
+    from the power-law model including the higher-mass samples in the
+    MCMC.  The distribution for $\Mmax$ is quite broad because the
+    best fit power laws slope downward ($\alpha < 0$), making this
+    parameter less relevant.  The range $-5.05 \leq \alpha \leq -1.77$
+    encloses 90\% of the probability; the median value of $\alpha$ is
+    -3.23.  The presence of the higher-mass samples in the analysis
+    produces a distinctive tail, eliminating the correlations
+    discussed in Section \ref{sec:power-law} and displayed in Figure
+    \ref{fig:power-law-2D} for the low-mass subset of the
+    observations. }
 \end{figure}
@@
 
-\subsubsection{High-Mass Model Selection}
+\subsubsection{Model Selection for the Complete Sample}
 \label{sec:high-mass-model-selection}
@@
 displayed in Figure \ref{fig:high-rj-evidence}; compare Figure
-\ref{fig:rj}.  The exponential model is the most favored model for the
-complete sample, with the two-Gaussian model the second-most favored.
-The ranking of models differs significantly from the low-mass
-samples.  The improvement of the exponential model relative to the
-low-mass analysis is encouraging for theoretical calculations that
-attempt to model the entire population of X-ray binaries with this
-mass model.  Note also that the increasing structure of the mass
+\ref{fig:rj}.  The relative model probabilities are given in Table
+\ref{tab:rj-high}.  The exponential model is the most favored model
+for the complete sample, with the two-Gaussian model the second-most
+favored.  The ranking of models differs significantly from the
+low-mass samples.  The improvement of the exponential model relative
+to the low-mass analysis is encouraging for theoretical calculations
+that attempt to model the entire population of X-ray binaries with
+this mass model.  Note also that the increasing structure of the mass
 distribution favors histogram models with three bins over those with
@@
     applied to all 20 systems in Table \ref{tab:sources} (i.e.\
-    including the higher-mass systems).  In increasing order along the
-    $x$-axis, the models are the power-law of Section
-    \ref{sec:power-law} (PL), the decaying exponential of Section
-    \ref{sec:exponential} (E), the single Gaussian of Section
-    \ref{sec:gaussian} (G), the double Gaussian of Section
-    \ref{sec:gaussian} (TG), and the one-, two-, three-, four-, and
-    five-bin histogram models of Section
+    including the higher-mass systems).  (See also Table
+    \ref{tab:rj-high}.)  In increasing order along the $x$-axis, the
+    models are the power-law of Section \ref{sec:power-law} (PL), the
+    decaying exponential of Section \ref{sec:exponential} (E), the
+    single Gaussian of Section \ref{sec:gaussian} (G), the double
+    Gaussian of Section \ref{sec:gaussian} (TG), and the one-, two-,
+    three-, four-, and five-bin histogram models of Section
     \ref{sec:non-parametric-models} (H1, H2, H3, H4, H5,
@@
 \end{figure}
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{|l|r|}
+      \hline
+      Model & Relative Evidence \\
+      \hline \hline
+      Exponential (Section \ref{sec:exponential}) & 0.346944 \\
+      \hline
+      Two Gaussian (Section \ref{sec:gaussian}) & 0.304923 \\
+      \hline
+      Power Law (Section \ref{sec:power-law}) & 0.120313 \\
+      \hline
+      Log Normal (Section \ref{sec:log-normal}) & 0.102536 \\
+      \hline
+      Histogram (3 Bin, Section \ref{sec:non-parametric-models}) &
+      0.0473464 \\
+      \hline
+      Histogram (4 Bin, Section \ref{sec:non-parametric-models}) &
+      0.0282086 \\
+      \hline 
+      Histogram (2 Bin, Section \ref{sec:non-parametric-models}) &
+      0.0210994 \\
+      \hline 
+      Histogram (5 Bin, Section \ref{sec:non-parametric-models}) &
+      0.0179703  \\
+      \hline      
+      Gaussian (Section \ref{sec:gaussian}) & 0.00901719 \\ 
+      \hline
+      Histogram (1 Bin, Section \ref{sec:non-parametric-models}) &
+      0.00164214 \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:rj-high} Relative probabilities of the various models
+    from Section \ref{sec:models} implied by the complete sample of systems.  (See also Figure \ref{fig:high-rj-evidence}.)  These probabilities have been 
+    computed from reversible-jump MCMC samples using the efficient jump proposal algorithm in Appendix \ref{sec:reversible-jump-mcmc}.}
+\end{table}
 
@@
 
-The distribution of the minimum black hole mass from the analysis of
-the low-mass samples is displayed in Figure \ref{fig:min-mass}.  For
-all distributions, the minimum black hole mass is defined as the 1\%
-mass quantile (i.e.\ the mass lying below 99\% of the mass
-distribution).  (A cutoff quantile is necessary in the case of those
-distributions that do not have a hard cutoff mass; even for those that
-do, like the power-law model, it can be useful to define a ``soft''
-cutoff in the event that the lower mass hard cutoff becomes an
-irrelevant parameter as discussed in Section \ref{sec:power-law}.)
+The distributions of the minimum black hole mass from the analysis of
+the low-mass samples is displayed in Figure \ref{fig:min-mass}.  The
+minimum black hole mass is defined as the 1\% mass quantile,
+$M_{1\%}$, of the black-hole mass distribution (i.e.\ the mass lying
+below 99\% of the mass distribution).  (A quantile-based definition is
+necessary in the case of those distributions that do not have a hard
+cutoff mass; even for those that do, like the power-law model, it can
+be useful to define a ``soft'' cutoff in the event that the lower mass
+hard cutoff becomes an irrelevant parameter as discussed in Section
+\ref{sec:power-law}.)  \changed{For each mass distribution parameter
+  sample from our MCMC, we can calculate the distribution's minimum
+  black hole mass; the collection of these minimum black hole masses
+  approximates the distribution of minimum black hole masses implied
+  by the data in the context of that distribution.  Figure
+  \ref{fig:min-mass} plots histograms of the minimum black hole mass
+  samples.}
 
@@
   \caption{\label{fig:min-mass} The distributions for the minimum
-    black hole mass calculated from the MCMC samples for the models in
-    Section \ref{sec:models} applied to the low-mass systems.  The
-    minimum black hole mass is defined as the 1\% mass quantile.  For
-    the most favored models, the power-law and Gaussian, the 90\%
-    confidence limit on the minimum black hole mass is 4.3 $\Msun$ and
-    2.9 $\Msun$, respectively.  In all plots, we indicate the 90\%
-    confidence limit on the minimum black hole mass with a vertical
-    line.}
+    black hole mass, $M_{1\%}$, calculated from the MCMC samples for
+    the models in Section \ref{sec:models} applied to the low-mass
+    systems.  For the most favored models, the power-law and Gaussian,
+    the 90\% confidence limit on the minimum black hole mass is 4.3
+    $\Msun$ and 2.9 $\Msun$, respectively.  In all plots, we indicate
+    the 90\% confidence bound (i.e.\ the 10\% quantile) on the minimum
+    black hole mass with a vertical line.}
 \end{figure}
 
-We find that the best-fit model for the low-mass systems (the
-power-law) has a 90\% confidence limit on the minimum black hole mass
-of 4.3 $\Msun$.  This is significantly above the maximum
+\changed{We find that the best-fit model for the low-mass systems (the
+  power-law) has $M_{1\%} > 4.3$ $\Msun$ in 90\% of the MCMC samples
+  (i.e.\ at 90\% confidence).  This is significantly above the maximum
   theoretically-allowed neutron star mass, $\sim 3 \Msun$ (e.g.\
@@
 
-The distribution of minimum black hole masses for the analysis of the
-complete sample (i.e.\ including the higher-mass systems) is shown in
-Figure \ref{fig:high-min-mass}.  For the most favored model, the
-exponential, the 90\% confidence limit on the minimum black hole mass
-is 4.5 $\Msun$.  We therefore conclude that there is strong evidence
-for a mass gap in the complete sample as well.
+  The distribution of minimum black hole masses for the analysis of
+  the complete sample (i.e.\ including the higher-mass systems) is
+  shown in Figure \ref{fig:high-min-mass}.  For the most favored
+  model, the exponential, we find that $M_{1\%} > 4.5$ $\Msun$ with
+  90\% confidence.  We therefore conclude that there is strong
+  evidence for a mass gap in the complete sample as well.}
 
@@
   \caption{\label{fig:high-min-mass} The distributions for the minimum
-    black hole mass calculated from the MCMC samples for the models in
-    Section \ref{sec:models} using the complete sample of systems.
-    The minimum black hole mass is defined as the 1\% mass quantile.
-    For the two most favored models, the exponential and two-Gaussian,
-    the 90\% confidence limit on the minimum black hole mass is 4.5
-    $\Msun$ and 2.3 $\Msun$, respectively.  For every model, we
-    indicate the 90\% confidence limit on the minimum black hole mass
-    with a vertical line.}
+    black hole mass, $M_{1\%}$, calculated from the MCMC samples for
+    the models in Section \ref{sec:models} using the complete sample
+    of systems.  For the two most favored models, the exponential and
+    two-Gaussian, the 90\% confidence limit on the minimum black hole
+    mass is 4.5 $\Msun$ and 2.3 $\Msun$, respectively.  For every
+    model, we indicate the 90\% confidence bound on the minimum black
+    hole mass with a vertical line.}
 \end{figure}
 
+\changed{Table \ref{tab:mmin-quants} gives the 10\%, 50\% (median),
+  and 90\% quantiles for the minimum black hole mass implied by the
+  low-mass sample; Table \ref{tab:mmin-quants-high} gives the same,
+  but for the complete sample of systems.}
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{|l|l|c|r|}
+      \hline
+      Model & 10\% & 50\% & 90\% \\
+      \hline \hline
+      Power Law (Section \ref{sec:power-law}) & 4.3 & 6.1 & 6.6 \\
+      \hline
+      Gaussian (Section \ref{sec:gaussian}) & 2.9 & 4.4 & 5.5 \\
+      \hline 
+      Log Normal (Section \ref{sec:log-normal}) & 3.9 & 4.9 & 5.8 \\
+      \hline 
+      Exponential (Section \ref{sec:exponential}) & 5.3 & 6.0 & 6.5 \\
+      \hline 
+      Two Gaussian (Section \ref{sec:gaussian}) & 2.4 & 4.2 & 5.5 \\
+      \hline 
+      Histogram (1 Bin, Section \ref{sec:non-parametric-models}) & 4.4 & 5.5 & 6.2 \\
+      \hline
+      Histogram (2 Bin, Section \ref{sec:non-parametric-models}) & 4.0 & 5.4 & 6.3 \\
+      \hline 
+      Histogram (3 Bin, Section \ref{sec:non-parametric-models}) & 3.2 & 5.2 & 6.3 \\
+      \hline
+      Histogram (4 Bin, Section \ref{sec:non-parametric-models}) & 2.4 & 4.7 & 6.0 \\
+      \hline
+      Histogram (5 Bin, Section \ref{sec:non-parametric-models}) & 1.9 & 4.4 & 6.0 \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:mmin-quants} The 10\%, 50\% (median), and 90\% quantiles for the minimum black hole mass (in units of $\Msun$) implied by the low-mass sample in the context of the various models for the black hole mass distribution.
+The models are listed in order of preference from model selection (Section \ref{sec:low-mass-model-selection}, Figure \ref{fig:rj}, and Table \ref{tab:rj}).}
+\end{table}
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{|l|l|c|r|}
+      \hline
+      Model & 10\% & 50\% & 90\% \\
+      \hline \hline
+      Exponential (Section \ref{sec:exponential}) & 4.5 & 5.4 & 6.1 \\
+      \hline
+      Two Gaussian (Section \ref{sec:gaussian}) & 2.3 & 4.3 & 5.5 \\
+      \hline
+      Power Law (Section \ref{sec:power-law}) & 5.1 & 5.9 & 6.4 \\
+      \hline
+      Histogram (3 Bin, Section \ref{sec:non-parametric-models}) & 4.0 & 5.5 & 6.3 \\
+      \hline 
+      Histogram (4 Bin, Section \ref{sec:non-parametric-models}) & 3.4 & 5.3 & 6.4 \\
+      \hline
+      Histogram (2 Bin, Section \ref{sec:non-parametric-models}) & 4.4 & 5.5 & 6.2 \\
+      \hline
+      Histogram (5 Bin, Section \ref{sec:non-parametric-models}) & 2.8 & 5.0 & 6.2 \\
+      \hline
+      Gaussian (Section \ref{sec:gaussian}) & -0.64 & 1.4 & 3.4 \\
+      \hline 
+      Histogram (1 Bin, Section \ref{sec:non-parametric-models}) & 2.9 & 4.4 & 5.5 \\
+      \hline
+    \end{tabular}
+  \end{center}
+  \caption{\label{tab:mmin-quants-high} The 10\%, 50\% (median), and 90\% quantiles for the distribution of minimum black hole masses (in units of $\Msun$) implied by the complete sample in the context of the various models for the black hole mass distribution.
+The models are listed in order of preference from model selection (Section \ref{sec:high-mass-model-selection}, Figure \ref{fig:high-rj-evidence}, and Table \ref{tab:rj-high}).
+}
+\end{table}
+
 \section{Summary and Conclusion}
+\label{sec:conclusion}
 
@@
 higher-mass, wind-fed X-ray binaries.  We used MCMC methods to sample
-the posterior distributions of the parameters for five parametric
-models and five non-parametric (histogram) models for the mass
-distribution that are implied by these data.  For both sets of
-samples, we used reversible jump MCMCs (exploiting a new algorithm for
-efficient jump proposals in such calculations) to perform model
-selection on the suite of models.  \changed{The consideration of a
-  broad range of models and the model-selection analysis, along with
-  consideration of the full posterior distribution on the minimum
-  black hole mass significantly expand earlier analyses of black hole
-  mass measurements \citep{Bailyn1998,Ozel2010}.}
+the posterior distributions of the parameters implied by the data for
+five parametric models and five non-parametric (histogram) models for
+the mass distribution.  For both sets of samples, we used reversible
+jump MCMCs (exploiting a new algorithm for efficient jump proposals in
+such calculations) to perform model selection on the suite of models.
+\changed{The consideration of a broad range of models and the
+  model-selection analysis, along with consideration of the full
+  posterior distribution on the minimum black hole mass significantly
+  expand earlier analyses of black hole mass measurements
+  \citep{Bailyn1998,Ozel2010}.}
 
@@
 Tables \ref{tab:low-mass-parametric} and
-\ref{tab:low-mass-non-parametric}.  The relative model probabilites
+\ref{tab:low-mass-non-parametric}.  The relative model probabilities
 from the model selection are given in Table \ref{tab:rj}.  The most
@@
   binary evolution and transient behavior (e.g.\
-  \citet{Fragos2008,Fragos2009}) could help settle this possibility.}
+  \citet{Fragos2008,Fragos2009}) could help shed light on this
+  possibility.}
 
@@
 
-This work was supported by grant \fixme{XXX}.  IM acknowledges support
-from the NSF AAPF under award AST-0901985.  Calculations for this work
-were performed on the Northwestern Fugu cluster, which was partially
-funded by NSF MRI grant \fixme{XXX}.  We thank Jonathan Gair for
-helpful discussions.
+AC, LK, and CB are supported by NSF grant NSF/AST-0707627.  WMF, NS,
+and VK are supported by \fixme{XXX}.  IM acknowledges support from the
+NSF AAPF under award AST-0901985.  Calculations for this work were
+performed on the Northwestern Fugu cluster, which was partially funded
+by NSF MRI grant \fixme{XXX}.  We thank Jonathan Gair for helpful
+discussions.
 
@@
 \section{Markov Chain Monte Carlo}
+\label{sec:mcmc}
 
@@
   \item With probability $\min(1,p_{\textnormal{accept}})$ ``accept''
-    the propsed $\vtheta_p$, setting $\vtheta_{i+1} = \vtheta_p$;
+    the proposed $\vtheta_p$, setting $\vtheta_{i+1} = \vtheta_p$;
     otherwise set $\vtheta_{i+1} = \vtheta_i$.

### 2010-11-05 paper.tex
Final changes before ApJ.

@@
   calculating their relative Bayesian evidence as supported by the
-  measurements, \changed{taking into account the number of degrees of
-    freedom of each model.}  The mass distribution of the low-mass
-  systems is best fit by a power-law, while the distribution of the
-  complete sample is best fit by the exponential model.  \changed{This
-    difference clearly indicates that the character of the low mass
-    X-ray binary systems differs from the higher-mass systems.}  We
-  examine the existence of a ``gap'' between the most massive neutron
-  stars and least massive black holes.  \changed{Our analysis
-    generates distributions of lower bounds on the black hole mass
-    distribution; we use the 1\% quantile of the mass distribution,
-    $M_{1\%}$, for a lower bound because some of our distributions do
-    not have a sharp lower limit.  For the best model (the power law)
-    fitted to the low-mass systems this distribution of lower-bounds
-    has $M_{1\%} > 4.3$ $\Msun$ with 90\% confidence, while the best
-    model (the exponential) fitted to all 20 systems has $M_{1\%} >
-    4.5$ $\Msun$ with 90\% confidence.}  We conclude that our sample
-  of black hole masses provides strong evidence of a gap between the
-  maximum neutron star mass and \changed{the lower bound on} black
-  hole masses.  \changed{Our results on the low-mass sample are in
-    qualitative agreement with those of \citet{Ozel2010}, although our
-    broad model selection analysis more reliably reveals the best-fit
+  measurements, taking into account the number of degrees of freedom
+  of each model.  The mass distribution of the low-mass systems is
+  best fit by a power-law, while the distribution of the combined
+  sample is best fit by the exponential model.  This difference
+  indicates that the low-mass subsample is not consistent with being
+  drawn from the distribution of the combined population.  We examine
+  the existence of a ``gap'' between the most massive neutron stars
+  and the least massive black holes by considering the value,
+  $M_{1\%}$, of the 1\% quantile from each black hole mass
+  distribution as the lower bound of black hole masses.  Our analysis
+  generates posterior distributions for $M_{1\%}$; the best model (the
+  power law) fitted to the low-mass systems has a distribution of
+  lower-bounds with $M_{1\%} > 4.3$ $\Msun$ with 90\% confidence,
+  while the best model (the exponential) fitted to all 20 systems has
+  $M_{1\%} > 4.5$ $\Msun$ with 90\% confidence.  We conclude that our
+  sample of black hole masses provides strong evidence of a gap
+  between the maximum neutron star mass and the lower bound on black
+  hole masses.  Our results on the low-mass sample are in qualitative
+  agreement with those of \citet{Ozel2010}, although our broad
+  model-selection analysis more reliably reveals the best-fit
   quantitative description of the underlying mass distribution.  The
-    results on the complete sample of low- and higher-mass systems are
+  results on the combined sample of low- and higher-mass systems are
   in qualitative agreement with \citet{Fryer2001} although the
-    presence of a mass gap remains theoretically unexplained.}
+  presence of a mass gap remains theoretically unexplained.
 \end{abstract}
@@
 
-Observations of low-mass X-ray binaries in both the optical and X-ray
-can provide a measurement of the mass of the compact object in these
-systems.  \changed{The current sample of stellar mass black holes with
-  dynamically measured masses includes 15 low-mass, Roche lobe
-  overflowing systems and 5 wind-fed higher-mass systems.  Hence,
-  sophisticated statistical analyses of the black hole mass
-  distribution in these systems are possible.}
+Observations of X-ray binaries in both the optical and X-ray bands can
+provide a measurement of the mass of the compact object in these
+systems.  The current sample of stellar mass black holes with
+dynamically measured masses includes 15 systems with low-mass, Roche
+lobe overflowing donors and 5 wind-fed systems with higher-mass
+donors.  Hence, sophisticated statistical analyses of the black hole
+mass distribution in these systems are possible.
 
@@
 analysis that the mass function was strongly peaked around seven solar
-masses.  \changed{(A similar analysis of the neutron star mass
-  distribution can be found in \citet{Finn1994}.)}  \citet{Bailyn1998}
-found evidence of a ``gap'' between the least massive black hole and a
-``safe'' upper limit for neutron star masses of 3 $\Msun$ (e.g.\
-\citet{Kalogera1996}).  Such a gap is puzzling in light of theoretical
-studies that predict a continuous distribution of compact object
-supernova remnant masses with a smooth transition from neutron stars
-to black holes \citep{Fryer2001}.  \changed{(But note that
+masses%
+\footnote{A similar analysis of the neutron star mass distribution can
+  be found in \citet{Finn1994}.}. %
+\citet{Bailyn1998} found evidence of a ``gap'' between the least
+massive black hole and a ``safe'' upper limit for neutron star masses
+of 3 $\Msun$ (e.g.\ \citet{Kalogera1996}).  Such a gap is puzzling in
+light of theoretical studies that predict a continuous distribution of
+compact object supernova remnant masses with a smooth transition from
+neutron stars to black holes \citep{Fryer2001}.  (We note that
 \citet{Fryer2001} considered binary evolution effects only
 heuristically and put forward some possible explanations for the gap
-  from \citet{Bailyn1998} both in the context of selection effects or
-  in connection to the energetics of supernova explosions.)}
-
-\changed{Towards the end of our analysis work, we became aware of a
-  more recent study \citep{Ozel2010}, also in a Bayesian framework,
-  analyzing the low-mass X-ray binary sample.}  It largely confirmed
-these findings, examining 16 low-mass X-ray binary systems containing
-black holes and finding a strongly peaked distribution at $7.8 \pm 1.2
-\, \Msun$.  \citet{Ozel2010} use two models for the mass function: a
-Gaussian and a decaying exponential with a minimum ``turn-on'' mass
-(motivated by the analytical model of the black-hole mass function in
-\citet{Fryer2001}).  \changed{\citet{Ozel2010} do not provide
-  confidence limits for the minimum black hole mass, instead
+from \citet{Bailyn1998} both in the context of selection effects or in
+connection to the energetics of supernova explosions.)
+
+Towards the end of our analysis work, we became aware of a more recent
+study \citep{Ozel2010}, also in a Bayesian framework, analyzing the
+low-mass X-ray binary sample.  Our results are largely consistent with
+those obtained by \citet{Ozel2010}, who examined 16 low-mass X-ray
+binary systems containing black holes and found a strongly peaked
+distribution at $7.8 \pm 1.2 \, \Msun$.  They used two models for the
+mass function: a Gaussian and a decaying exponential with a minimum
+``turn-on'' mass (motivated by the analytical model of the black-hole
+mass function in \citet{Fryer2001}).  We note that \citet{Ozel2010} do
+not provide confidence limits for the minimum black hole mass, instead
 discussing only the model parameters at the peak of their posterior
-  distributions.  They also do not perform any model selection
-  analysis; thus, they give the distribution of parameters within each
-  of their models, but cannot say which model is more likely to
-  correspond to the true distribution of black hole masses.}
-Nevertheless, it appears that their analysis confirms the existence of
-a mass gap.  \citet{Ozel2010} discuss possible selection effects that
-could lead to the appearance of a mass gap, but conclude these effects
-could not produce the observed gap, which they therefore claim is a
-real property of the black hole mass distribution.
-
-In this work, we use observations of 20 galactic X-ray binaries that
-contain black holes to constrain the mass function for stellar mass
-black holes.  We consider separately 15 low-mass X-ray binaries
-containing Roche-lobe-filling secondaries with masses smaller than the
-black hole primary, and a sample including these systems and five
-higher-mass, wind-fed systems where the secondary is heavier than the
-black hole primary.
+distributions.  They also do not perform any model selection analysis;
+thus, they give the distribution of parameters within each of their
+models, but cannot say which model is more likely to correspond to the
+true distribution of black hole masses.  Nevertheless, it appears that
+their analysis confirms the existence of a mass gap.  \citet{Ozel2010}
+discuss possible selection effects that could lead to the appearance
+of a mass gap, but conclude these effects could not produce the
+observed gap, which they therefore claim is a real property of the
+black hole mass distribution.
 
@@
 
-\changed{We define the ``minimum black hole mass'' to be the 1\%
-  quantile, $M_{1\%}$, in the black hole mass distribution (see
-  Section \ref{sec:minimum-mass}).  In agreement with \citet{Ozel2010}
-  and \citet{Bailyn1998}, we find strong evidence for a mass gap among
-  the best models for both samples.  Our analysis gives distributions
-  for $M_{1\%}$ implied by the data in the context of each of our
-  models for the black hole mass distribution.  In the context of the
-  best model for the low-mass systems (a power-law), the distribution
-  for $M_{1\%}$ has $M_{1\%} > 4.3$ $\Msun$ (at 90\% confidence); in
-  the context of the best model for the combined sample of lower- and
-  high-mass systems the distribution of $M_{1\%}$ has $M_{1\%} > 4.5$
-  $\Msun$ with 90\% confidence.  Further, in the context of models
-  with lower evidence, most also have a mass gap, with 90\% confidence
-  bounds on $M_{1\%}$ significantly above a ``safe'' maximum neutron
-  star mass of 3 $\Msun$ \citep{Kalogera1996}.}
-
-\changed{We find that, for the low-mass X-ray binary sample, the
-  theoretical model from \citet{Fryer2001}, a decaying exponential is
-  strongly disfavored by our model selection.  We find that the
-  low-mass systems are best described by a power law, followed
-  closely by a Gaussian (which is the other model from
-  \citet{Ozel2010}).  On the other hand, we find that the theoretical
-  model from \citet{Fryer2001} is the preferred model for the combined
-  sample of low- and high-mass X-ray binaries.  A model with two
-  separate Gaussian peaks also has relatively high evidence for the
-  combined sample of systems.  The difference in best-fitting model
-  clearly indicates that the character of the low mass X-ray binary
-  systems differs from the higher-mass systems.}
-
-\changed{The structure of this paper is as follows.  In Section
-  \ref{sec:systems} we discuss the 15 systems that comprise the
-  low-mass X-ray binary black hole sample and the 5 additional
-  higher-mass, wind-fed systems that make up the complete sample.  In
-  Section \ref{sec:models} we discuss the Bayesian techniques we use
-  to analyze the black hole mass distribution, the techniques we use
-  for model selection, and the parametric and non-parametric models we
-  will for the black hole mass distribution.  In Section
-  \ref{sec:results} we discuss the results of our analysis and model
-  selection.  In Section \ref{sec:minimum-mass} we discuss the
-  distribution of the minimum black hole mass implied by the analysis
-  of Section \ref{sec:results}.  In Section \ref{sec:conclusion} we
-  summarize our results and comment on the significance of the
-  observed mass gap in the context of theoretical models.  Appendix
-  \ref{sec:mcmc} describes MCMC techniques in some detail.  Appendix
-  \ref{sec:reversible-jump-mcmc} explains our novel algorithm for
-  efficiently performing the reversible jump MCMC computations used in
-  the model comparisons of Section \ref{sec:results} (but see also
-  \citet{Farr2010}).}
+We define the ``minimum black hole mass'' to be the 1\% quantile,
+$M_{1\%}$, in the black hole mass distribution (see Section
+\ref{sec:minimum-mass}).  In qualitative agreement with
+\citet{Ozel2010} and \citet{Bailyn1998}, we find strong evidence for a
+mass gap among the best models for both samples.  Our analysis gives
+distributions for $M_{1\%}$ implied by the data in the context of each
+of our models for the black hole mass distribution.  In the context of
+the best model for the low-mass systems (a power-law), the
+distribution for $M_{1\%}$ gives $M_{1\%} > 4.3$ $\Msun$ with 90\%
+confidence; in the context of the best model for the combined sample
+of lower- and high-mass systems the distribution of $M_{1\%}$ has
+$M_{1\%} > 4.5$ $\Msun$ with 90\% confidence.  Further, in the context
+of models with lower evidence, most also have a mass gap, with 90\%
+confidence bounds on $M_{1\%}$ significantly above a ``safe'' maximum
+neutron star mass of 3 $\Msun$ \citep{Kalogera1996}.
+
+We find that, for the low-mass X-ray binary sample, the theoretical
+model from \citet{Fryer2001}---a decaying exponential---is strongly
+disfavored by our model selection.  We find that the low-mass systems
+are best described by a power law, followed closely by a Gaussian
+(which is the second model considered by \citet{Ozel2010}).  On the
+other hand, we find that the theoretical model from \citet{Fryer2001}
+is the preferred model for the combined sample of low- and high-mass
+X-ray binaries.  A model with two separate Gaussian peaks also has
+relatively high evidence for the combined sample of systems.  The
+difference in best-fitting model indicates that the low-mass subsample
+is not consistent with being drawn from the distribution of the
+combined population.
+
+The structure of this paper is as follows.  In Section
+\ref{sec:systems} we discuss the 15 systems that comprise the low-mass
+X-ray binary black hole sample and the 5 additional higher-mass,
+wind-fed systems that make up the combined sample.  In Section
+\ref{sec:models} we discuss the Bayesian techniques we use to analyze
+the black hole mass distribution, the techniques we use for model
+selection, and the parametric and non-parametric models we will for
+the black hole mass distribution.  In Section \ref{sec:results} we
+discuss the results of our analysis and model selection.  In Section
+\ref{sec:minimum-mass} we discuss the distribution of the minimum
+black hole mass implied by the analysis of Section \ref{sec:results}.
+In Section \ref{sec:conclusion} we summarize our results and comment
+on the significance of the observed mass gap in the context of
+theoretical models.  Appendix \ref{sec:mcmc} describes MCMC techniques
+in some detail.  Appendix \ref{sec:reversible-jump-mcmc} explains our
+novel algorithm for efficiently performing the reversible jump MCMC
+computations used in the model comparisons of Section
+\ref{sec:results} (but see also \citet{Farr2010}).
 
@@
 systems alone (Section \ref{sec:results-low-mass}), and then the
-complete sample of 20 systems (Section \ref{sec:higher-mass}).
+combined sample of 20 systems (Section \ref{sec:higher-mass}).
 
@@
 various models for the underlying mass distribution from which the
-low-mass sample and the complete sample of X-ray binary systems in
+low-mass sample and the combined sample of X-ray binary systems in
 Table \ref{tab:sources} were drawn.  The results of our analysis are
@@
 \ref{sec:high-mass-model-selection}) from both the low-mass and
-complete datasets is decreasing as the number of bins reaches five,
+combined datasets is decreasing as the number of bins reaches five,
 indicating that increasing the number of bins beyond five would not
@@
 systems, while those of Section \ref{sec:higher-mass} apply to the
-complete sample of systems.
+combined sample of systems.
 
@@
     These distributions use only the 15 low-mass observations in Table
-    \ref{tab:sources} (the complete sample is analyzed in Section
+    \ref{tab:sources} (the combined sample is analyzed in Section
     \ref{sec:higher-mass}).  Note that these ``distributions of
@@
 
-\subsection{Complete Sample}
+\subsection{Combined Sample}
 \label{sec:higher-mass}
@@
     \ref{sec:non-parametric-models}.  These distributions use the
-    complete sample of 20 observations in Table \ref{tab:sources},
+    combined sample of 20 observations in Table \ref{tab:sources},
     including the higher-mass, wind-fed systems.  Note that these
@@
 
-\subsubsection{Model Selection for the Complete Sample}
+\subsubsection{Model Selection for the Combined Sample}
 \label{sec:high-mass-model-selection}
@@
 \ref{tab:rj-high}.  The exponential model is the most favored model
-for the complete sample, with the two-Gaussian model the second-most
+for the combined sample, with the two-Gaussian model the second-most
 favored.  The ranking of models differs significantly from the
@@
   \caption{\label{tab:rj-high} Relative probabilities of the various models
-    from Section \ref{sec:models} implied by the complete sample of systems.  (See also Figure \ref{fig:high-rj-evidence}.)  These probabilities have been 
+    from Section \ref{sec:models} implied by the combined sample of systems.  (See also Figure \ref{fig:high-rj-evidence}.)  These probabilities have been 
     computed from reversible-jump MCMC samples using the efficient jump proposal algorithm in Appendix \ref{sec:reversible-jump-mcmc}.}
@@
 distribution in X-ray binaries to place constraints on the minimum
-black hole mass.  \citet{Bailyn1998} addressed this question in the
-context of a ``mass gap'' between the most massive neutron stars and
-the least massive black holes.  The more recent study of
-\citet{Ozel2010} also looked for a mass gap using a subset of the
-models and systems presented here.  Both works found that the minimum
-black hole mass is significantly above the maximum neutron star mass
-\citep{Kalogera1996} of $\sim 3 \Msun$ \changed{(though
+black hole mass implied by the present sample.  \citet{Bailyn1998}
+addressed this question in the context of a ``mass gap'' between the
+most massive neutron stars and the least massive black holes.  The
+more recent study of \citet{Ozel2010} also looked for a mass gap using
+a subset of the models and systems presented here.  Both works found
+that the minimum black hole mass is significantly above the maximum
+neutron star mass \citep{Kalogera1996} of $\sim 3 \Msun$ (though
 \citet{Ozel2010} only state their evidence for a gap in terms of the
-  maximum-posterior parameters for their distributions)}.
+maximum-posterior parameters and not the full extent of their
+distributions).
 
 The distributions of the minimum black hole mass from the analysis of
-the low-mass samples is displayed in Figure \ref{fig:min-mass}.  The
+the low-mass samples are displayed in Figure \ref{fig:min-mass}.  The
 minimum black hole mass is defined as the 1\% mass quantile,
@@
 hard cutoff becomes an irrelevant parameter as discussed in Section
-\ref{sec:power-law}.)  \changed{For each mass distribution parameter
-  sample from our MCMC, we can calculate the distribution's minimum
-  black hole mass; the collection of these minimum black hole masses
-  approximates the distribution of minimum black hole masses implied
-  by the data in the context of that distribution.  Figure
-  \ref{fig:min-mass} plots histograms of the minimum black hole mass
-  samples.}
+\ref{sec:power-law}.)  For each mass distribution parameter sample
+from our MCMC, we can calculate the distribution's minimum black hole
+mass; the collection of these minimum black hole masses approximates
+the distribution of minimum black hole masses implied by the data in
+the context of that distribution.  Figure \ref{fig:min-mass} plots
+histograms of the minimum black hole mass samples.
 
@@
 
-\changed{We find that the best-fit model for the low-mass systems (the
+We find that the best-fit model for the low-mass systems (the
 power-law) has $M_{1\%} > 4.3$ $\Msun$ in 90\% of the MCMC samples
@@
 
-  The distribution of minimum black hole masses for the analysis of
-  the complete sample (i.e.\ including the higher-mass systems) is
-  shown in Figure \ref{fig:high-min-mass}.  For the most favored
-  model, the exponential, we find that $M_{1\%} > 4.5$ $\Msun$ with
-  90\% confidence.  We therefore conclude that there is strong
-  evidence for a mass gap in the complete sample as well.}
+The distribution of minimum black hole masses for the analysis of the
+combined sample (i.e.\ including the higher-mass systems) is shown in
+Figure \ref{fig:high-min-mass}.  For the most favored model, the
+exponential, we find that $M_{1\%} > 4.5$ $\Msun$ with 90\%
+confidence.  We therefore conclude that there is strong evidence for a
+mass gap in the combined sample as well.
 
@@
     black hole mass, $M_{1\%}$, calculated from the MCMC samples for
-    the models in Section \ref{sec:models} using the complete sample
+    the models in Section \ref{sec:models} using the combined sample
     of systems.  For the two most favored models, the exponential and
@@
 
-\changed{Table \ref{tab:mmin-quants} gives the 10\%, 50\% (median),
-  and 90\% quantiles for the minimum black hole mass implied by the
-  low-mass sample; Table \ref{tab:mmin-quants-high} gives the same,
-  but for the complete sample of systems.}
+Table \ref{tab:mmin-quants} gives the 10\%, 50\% (median), and 90\%
+quantiles for the minimum black hole mass implied by the low-mass
+sample; Table \ref{tab:mmin-quants-high} gives the same, but for the
+combined sample of systems.
 
@@
   \end{center}
-  \caption{\label{tab:mmin-quants-high} The 10\%, 50\% (median), and 90\% quantiles for the distribution of minimum black hole masses (in units of $\Msun$) implied by the complete sample in the context of the various models for the black hole mass distribution.
+  \caption{\label{tab:mmin-quants-high} The 10\%, 50\% (median), and 90\% quantiles for the distribution of minimum black hole masses (in units of $\Msun$) implied by the combined sample in the context of the various models for the black hole mass distribution.
 The models are listed in order of preference from model selection (Section \ref{sec:high-mass-model-selection}, Figure \ref{fig:high-rj-evidence}, and Table \ref{tab:rj-high}).
@@
 such calculations) to perform model selection on the suite of models.
-\changed{The consideration of a broad range of models and the
-  model-selection analysis, along with consideration of the full
-  posterior distribution on the minimum black hole mass significantly
-  expand earlier analyses of black hole mass measurements
-  \citep{Bailyn1998,Ozel2010}.}
+The consideration of a broad range of models and the model-selection
+analysis, along with consideration of the full posterior distribution
+on the minimum black hole mass, significantly expand earlier
+statistical analyses of black hole mass measurements
+\citep{Bailyn1998,Ozel2010}.
 
@@
 most favored model for the combined sample is the exponential model.
-\changed{This difference clearly indicates that the character of the
-  low mass X-ray binary systems differs from the higher-mass systems.}
+This difference indicates that the low-mass subsample is not
+consistent with being drawn from the distribution of the combined
+population.
 
@@
 distribution whose 1\% quantile lies above $4.3 \Msun$ with 90\%
-confidence.  For the complete sample of systems, the most favored,
+confidence.  For the combined sample of systems, the most favored,
 exponential model gives a black hole mass distribution whose 1\%
-quantile lise above $4.5 \Msun$ with 90\% confidence.
-\changed{Although the study methodology was different, the existence
-  of a mass gap was pointed out first by \citet{Bailyn1998} and most
-  recently by \citet{Ozel2010} (who did not consider a power law
-  model, and applied the exponential model to the low-mass systems,
-  where it is strongly disfavored compared to the power-law model).}
-
-\changed{Theoretical expectations for the black hole mass distribution
-  have been examined in \citet{Fryer2001}.  They considered results of
-  supernova explosion and fallback simulations \citep{Fryer1999}
-  applied to single star populations; they also included a heuristic
-  treatment of the possible effects of binary evolution on the black
-  hole mass distribution.  It is interesting that we find the
-  most-favored model for the complete sample to be an exponential, as
-  discussed by \citet{Fryer1999}.  On the other hand, we find the
-  most-favored model for the low-mass sample to be a power law, with
-  the exponential model strongly disfavored for this sample. In
-  agreement with \citet{Bailyn1998} and \citet{Ozel2010}, we too
-  conclude that both the low-mass and complete samples require the
-  presence of a gap in the 3--4 or 4.5 $\Msun$ range.
+quantile lise above $4.5 \Msun$ with 90\% confidence.  Although the
+study methodology was different, the existence of a mass gap was
+pointed out first by \citet{Bailyn1998} and most recently by
+\citet{Ozel2010} (who did not consider a power law model, and applied
+both Gaussian and exponential models to the low-mass systems, where
+the exponential is strongly disfavored compared to our power-law
+model).
+
+Theoretical expectations for the black hole mass distribution have
+been examined in \citet{Fryer2001}.  They considered results of
+supernova explosion and fallback simulations \citep{Fryer1999} applied
+to single star populations; they also included a heuristic treatment
+of the possible effects of binary evolution on the black hole mass
+distribution.  It is interesting that we find the most-favored model
+for the combined sample to be an exponential, as discussed by
+\citet{Fryer2001}.  On the other hand, we find the most-favored model
+for the low-mass sample to be a power law, with the exponential model
+strongly disfavored for this sample. In agreement with
+\citet{Bailyn1998} and \citet{Ozel2010}, we too conclude that both the
+low-mass and combined samples require the presence of a gap between 3
+and 4--4.5 $\Msun$.
 
@@
 progenitor mass.  Selection biases can occur because the X-ray
-  binaries with very low-mass black holes systems are more likely to
-  be persistently Roche-lobe overflowing, preventing dynamical mass
+binaries with very low-mass black holes systems are more likely to be
+persistently Roche-lobe overflowing, preventing dynamical mass
 measurements.  \citet{Ozel2010} conclude that the presence of such
@@
 (26) of observed persistent X-ray sources not known to be neutron
-  stars is insufficient to populate the 2--5 $\Msun$ region of any
-  black hole mass distribution that rises toward low masses.
-  Population synthesis models incorporating sophisticated treatment of
-  binary evolution and transient behavior (e.g.\
-  \citet{Fragos2008,Fragos2009}) could help shed light on this
-  possibility.}
+stars is insufficient to populate the 2--5 $\Msun$ region of any black
+hole mass distribution that rises toward low masses.  Population
+synthesis models incorporating sophisticated treatment of binary
+evolution and transient behavior (e.g.\ \citet{Fragos2008,Fragos2009})
+could help shed light on this possibility.
 
@@
 
-AC, LK, and CB are supported by NSF grant NSF/AST-0707627.  WMF, NS,
-and VK are supported by \fixme{XXX}.  IM acknowledges support from the
-NSF AAPF under award AST-0901985.  Calculations for this work were
-performed on the Northwestern Fugu cluster, which was partially funded
-by NSF MRI grant \fixme{XXX}.  We thank Jonathan Gair for helpful
-discussions.
+WMF, NS, and VK are supported by NSF grants CAREER AST-0449558 and
+AST–0908930.  AC, LK, and CB are supported by NSF grant
+NSF/AST-0707627.  .  IM acknowledges support from the NSF AAPF under
+award AST-0901985.  Calculations for this work were performed on the
+Northwestern Fugu cluster, which was partially funded by NSF MRI grant
+PHY-0619274.  We thank Jonathan Gair for helpful discussions.
 

### 2010-11-05 paper.tex
As submitted.
higher-mass --> high-mass

@@
   X-ray binary systems undergoing Roche lobe overflow and five
-  higher-mass, wind-fed X-ray binary systems.  Using Markov Chain
+  high-mass, wind-fed X-ray binary systems.  Using Markov Chain
   Monte Carlo calculations, we model the mass distribution both
@@
   quantitative description of the underlying mass distribution.  The
-  results on the combined sample of low- and higher-mass systems are
+  results on the combined sample of low- and high-mass systems are
   in qualitative agreement with \citet{Fryer2001} although the
@@
 dynamically measured masses includes 15 systems with low-mass, Roche
-lobe overflowing donors and 5 wind-fed systems with higher-mass
+lobe overflowing donors and 5 wind-fed systems with high-mass
 donors.  Hence, sophisticated statistical analyses of the black hole
@@
 \ref{sec:systems} we discuss the 15 systems that comprise the low-mass
-X-ray binary black hole sample and the 5 additional higher-mass,
+X-ray binary black hole sample and the 5 additional high-mass,
 wind-fed systems that make up the combined sample.  In Section
@@
 systems in which the central black hole appears to be fed by
-roche-lobe overflow from the secondary, and 5 higher-mass systems in
+roche-lobe overflow from the secondary, and 5 high-mass systems in
 which the black hole is fed via winds (these systems all have a
 secondary that appears to be more massive than the black hole).  The
-low- and higher-mass systems undoubtedly have different evolutionary
+low- and high-mass systems undoubtedly have different evolutionary
 tracks, and therefore it is reasonable that they would have different
@@
 systems alone (Section \ref{sec:results-low-mass}), and then the
-combined sample of 20 systems (Section \ref{sec:higher-mass}).
+combined sample of 20 systems (Section \ref{sec:high-mass}).
 
@@
   \begin{center}
-    \plotone{plots/all-masses} 
+    \plotone{all-masses} 
   \end{center}
@@
   \begin{center}
-    \plotone{plots/high-masses}
+    \plotone{high-masses}
   \end{center}
   \caption{\label{fig:high-masses} Mass distributions for the
-    wind-fed, higher-mass systems computed from the distributions on
+    wind-fed, high-mass systems computed from the distributions on
     observed data in Table \ref{tab:sources} using Equation
@@
 distribution, we may find evidence of one when exploring the combined
-low- and higher-mass samples.  To look for a second peak in the
+low- and high-mass samples.  To look for a second peak in the
 black-hole mass distribution, we use a two-Gaussian model:
@@
 \ref{sec:results-low-mass} apply to the low-mass sample of
-systems, while those of Section \ref{sec:higher-mass} apply to the
+systems, while those of Section \ref{sec:high-mass} apply to the
 combined sample of systems.
@@
   \begin{center}
-    \plotone{plots/dist}
+    \plotone{dist}
   \end{center}
@@
     \ref{tab:sources} (the combined sample is analyzed in Section
-    \ref{sec:higher-mass}).  Note that these ``distributions of
+    \ref{sec:high-mass}).  Note that these ``distributions of
     distributions'' are not necessarily normalized, and need not be
@@
   \begin{center}
-    \plotone{plots/power-law}
+    \plotone{power-law}
   \end{center}
@@
 correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
-When we include the higher-mass systems in the analysis, the long tail
+When we include the high-mass systems in the analysis, the long tail
 will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
@@
   \begin{center}
-    \plotone{plots/power-law-2D}
+    \plotone{power-law-2D}
   \end{center}
@@
   \begin{center}
-    \plotone{plots/exp-cutoff}
+    \plotone{exp-cutoff}
   \end{center}
@@
   \begin{center}
-    \plotone{plots/exp-cutoff-2d}
+    \plotone{exp-cutoff-2d}
   \end{center}
@@
   \begin{center}
-    \plotone{plots/gaussian}
+    \plotone{gaussian}
   \end{center}
@@
   \begin{center}
-    \plotone{plots/two-gaussian}
+    \plotone{two-gaussian}
   \end{center}
@@
   \begin{center}
-    \plotone{plots/log-normal}
+    \plotone{log-normal}
   \end{center}
@@
   \begin{center}
-    \plotone{plots/rj}
+    \plotone{rj}
   \end{center}
@@
 \subsection{Combined Sample}
-\label{sec:higher-mass}
+\label{sec:high-mass}
 
 This section repeats the analysis of the models from Section
-\ref{sec:models}, but including the higher-mass, wind-fed systems from
+\ref{sec:models}, but including the high-mass, wind-fed systems from
 Table \ref{tab:sources} (see also Figure \ref{fig:high-masses}) in the
@@
 \ref{sec:models} applied to this data set; compare to Figure
-\ref{fig:dists}.  The inclusion of the higher-mass, wind-fed systems
+\ref{fig:dists}.  The inclusion of the high-mass, wind-fed systems
 tends to widen the distribution toward the high-mass end and, in
-models that allow it, produce a second, higher-mass peak in addition
+models that allow it, produce a second, high-mass peak in addition
 to the one in Figure \ref{fig:dists}.
@@
   \begin{center}
-    \plotone{plots/dist-high}
+    \plotone{dist-high}
   \end{center}
@@
     combined sample of 20 observations in Table \ref{tab:sources},
-    including the higher-mass, wind-fed systems.  Note that these
+    including the high-mass, wind-fed systems.  Note that these
     ``distributions of distributions'' are not necessarily normalized,
@@
     only the low-mass systems in the analysis.  Including the
-    higher-mass systems tends to widen the distribution toward the
+    high-mass systems tends to widen the distribution toward the
     high-mass end and, in models that allow it, produce a second,
-    higher-mass peak in addition to the one in Figure
+    high-mass peak in addition to the one in Figure
     \ref{fig:dists}. }
@@
 $\alpha$ (Section \ref{sec:power-law}) from an analysis including the
-higher-mass systems.  The distribution for $\Mmax$ is quite broad
+high-mass systems.  The distribution for $\Mmax$ is quite broad
 because the best fit power laws slope downward ($\alpha < 0$), making
@@
 -1.77$ encloses 90\% of the probability; the median value of $\alpha$
-is -3.23.  The presence of the higher-mass samples in the analysis
+is -3.23.  The presence of the high-mass samples in the analysis
 produces a distinctive tail, eliminating the correlations discussed in
@@
   \begin{center}
-    \plotone{plots/power-law-high}
+    \plotone{power-law-high}
   \end{center}
@@
     histogram), $\Mmax$ (top, solid histogram), and $\alpha$ (bottom)
-    from the power-law model including the higher-mass samples in the
+    from the power-law model including the high-mass samples in the
     MCMC.  The distribution for $\Mmax$ is quite broad because the
@@
     encloses 90\% of the probability; the median value of $\alpha$ is
-    -3.23.  The presence of the higher-mass samples in the analysis
+    -3.23.  The presence of the high-mass samples in the analysis
     produces a distinctive tail, eliminating the correlations
@@
 (Section \ref{sec:exponential}) from an analysis including the
-higher-mass systems.  The distribution for the scale mass, $M_0$, has
+high-mass systems.  The distribution for the scale mass, $M_0$, has
 moved to higher masses relative to Figure \ref{fig:exp-marginal} to
@@
   \begin{center}
-    \plotone{plots/exp-cutoff-high}
+    \plotone{exp-cutoff-high}
   \end{center}
@@
     defined in Section \ref{sec:exponential} from an analysis
-    including the higher-mass systems.  The distribution for the scale
+    including the high-mass systems.  The distribution for the scale
     mass, $M_0$, has moved to higher masses relative to Figure
@@
 for the Gaussian parameters (Section \ref{sec:gaussian}) when the
-higher-mass objects are included in the mass distribution.  The mean
+high-mass objects are included in the mass distribution.  The mean
 mass, $\mu$, and the mass standard deviation, $\sigma$, are both
 increased relative to Figure \ref{fig:gaussian} to account for the
-broader distribution and higher-mass tail.
+broader distribution and high-mass tail.
 
@@
   \begin{center}
-    \plotone{plots/gaussian-high}
+    \plotone{gaussian-high}
   \end{center}
   \caption{\label{fig:gaussian-high} The marginalized distributions
-    for the Gaussian parameters when the higher-mass objects are
+    for the Gaussian parameters when the high-mass objects are
     included in the mass distribution.  The mean mass, $\mu$ (solid
@@
     \ref{fig:gaussian} to account for the broader distribution and
-    higher-mass tail.  The peak of the underlying mass distribution
+    high-mass tail.  The peak of the underlying mass distribution
     lies in the range $7.8660 \leq \mu \leq 10.9836$ with 90\%
@@
 The analysis of the two-Gaussian model shows the largest change when
-the higher-mass samples are included.  Figure
+the high-mass samples are included.  Figure
 \ref{fig:two-gaussian-high} shows the marginalized distributions for
 the two-Gaussian parameters (Section \ref{sec:gaussian}) when the
-higher-mass samples are included in the analysis.  In stark contrast
+high-mass samples are included in the analysis.  In stark contrast
 to Figure \ref{fig:two-gaussian}, there are two well-defined,
 separated peaks; the low-mass peak reproduces the results from the
-low-mass samples, while the higher-mass peak ($13.5534 \leq \mu_2 \leq
+low-mass samples, while the high-mass peak ($13.5534 \leq \mu_2 \leq
 27.9481$ with 90\% confidence; median 20.3839) matches the new
-higher-mass samples.  The peak in $\alpha$ near 0.8 is consistent with
+high-mass samples.  The peak in $\alpha$ near 0.8 is consistent with
 approximately 4/5 the total probability being concentrated in the 15
@@
   \begin{center}
-    \plotone{plots/two-gaussian-high}
+    \plotone{two-gaussian-high}
   \end{center}
@@
     distributions for the two-Gaussian parameters (Section
-    \ref{sec:gaussian}) when the higher-mass samples are included in
+    \ref{sec:gaussian}) when the high-mass samples are included in
     the analysis.  In stark contrast to Figure \ref{fig:two-gaussian},
@@
     reproduces the results from the low-mass samples, while the
-    higher-mass peak ($13.5534 \leq \mu_2 \leq 27.9481$ with 90\%
-    confidence; median 20.3839) matches the new higher-mass samples.
+    high-mass peak ($13.5534 \leq \mu_2 \leq 27.9481$ with 90\%
+    confidence; median 20.3839) matches the new high-mass samples.
     The peak in $\alpha$ near 0.8 is consistent with approximately 15
@@
 The marginalized distributions for the log-normal parameters (Section
-\ref{sec:log-normal}) when the higher-mass samples are included in the
+\ref{sec:log-normal}) when the high-mass samples are included in the
 analysis are displayed in Figure \ref{fig:log-normal-high}.  The
-changes when the higher-mass samples are included (compare to Figure
+changes when the high-mass samples are included (compare to Figure
 \ref{fig:log-normal}) are similar to the changes in the Gaussian
@@
   \begin{center}
-    \plotone{plots/log-normal-high}
+    \plotone{log-normal-high}
   \end{center}
@@
     for the log-normal parameters (Section \ref{sec:log-normal}) when
-    the higher-mass samples are included in the analysis.  The changes
-    when the higher-mass samples are included (compare to Figure
+    the high-mass samples are included in the analysis.  The changes
+    when the high-mass samples are included (compare to Figure
     \ref{fig:log-normal}) are similar to the changes in the Gaussian
@@
     discussed in Section \ref{sec:parametric-models} when the
-    higher-mass samples are included in the analysis (compare to Table
+    high-mass samples are included in the analysis (compare to Table
     \ref{tab:low-mass-parametric}).  We indicate the 5\%, 15\%, 50\% 
@@
 \ref{sec:non-parametric-models}) models also show evidence of a long
-tail from the inclusion of the higher-mass samples.  Table
+tail from the inclusion of the high-mass samples.  Table
 \ref{tab:high-mass-non-parametric} displays confidence limits on the
-histogram parameters for the analysis including the higher-mass
+histogram parameters for the analysis including the high-mass
 systems; compare to Table \ref{tab:low-mass-non-parametric}.
@@
     \ref{sec:non-parametric-models} in an 
-    analysis including the higher-mass, wind-fed systems.  
+    analysis including the high-mass, wind-fed systems.  
     The tails evident in Figure \ref{fig:high-mass-dists} are apparent
@@
 \ref{sec:low-mass-model-selection} for the sample including the
-higher-mass systems, we find that the model probabilities have changed
+high-mass systems, we find that the model probabilities have changed
 with the inclusion of the extra five systems.  As before, we assume
@@
   \begin{center}
-    \plotone{plots/rj-high}
+    \plotone{rj-high}
   \end{center}
@@
     applied to all 20 systems in Table \ref{tab:sources} (i.e.\
-    including the higher-mass systems).  (See also Table
+    including the high-mass systems).  (See also Table
     \ref{tab:rj-high}.)  In increasing order along the $x$-axis, the
@@
   \begin{center}
-    \plotone{plots/mmin}
+    \plotone{mmin}
   \end{center}
@@
 The distribution of minimum black hole masses for the analysis of the
-combined sample (i.e.\ including the higher-mass systems) is shown in
+combined sample (i.e.\ including the high-mass systems) is shown in
 Figure \ref{fig:high-min-mass}.  For the most favored model, the
@@
 \begin{center}
-    \plotone{plots/mmin-high}
+    \plotone{mmin-high}
   \end{center}
@@
 sample of 20 systems containing the 15 low-mass systems and five
-higher-mass, wind-fed X-ray binaries.  We used MCMC methods to sample
+high-mass, wind-fed X-ray binaries.  We used MCMC methods to sample
 the posterior distributions of the parameters implied by the data for
@@
   \begin{center}
-    \plotone{plots/kdtree}
+    \plotone{kdtree}
   \end{center}
